### 2017-07-29 README.Rmd
inital

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# saaabstracts

The goal of saaabstracts is to ...

### Citation

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Title of compendium_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### Installation

You can install saaabstracts from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("benmarwick/saaabstracts")
```

### Licenses

**Text and figures:**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code:** See the [DESCRIPTION](DESCRIPTION) file

$\textbf{Data:}$ [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2017-07-29 analysis/paper/paper.Rmd
inital

---
title: "Title Goes Here"
author:
  - author 1
  - author 2
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
bibliography: references.bib
csl: journal-of-archaeological-science.csl
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "figures"
)

library(saaabstracts)
library(tm)
library(topicmodels)
library(readxl)
library(Rmpfr)
library(ggplot2)
```

# Introduction

```{r eval=FALSE}
# read in the abstracts
general <- read_excel("../data/raw_data/General Abstracts.xlsx")
```

```{r eval=FALSE}
# prepare a mapping structure, first item is column with text to model, others are metadata
m <- list(content = "Abstract",  # content
          id = "Abstract Id",      # metadata
          title = "Title", 
          author_first = "First Name",
          author_last = "Last Name",
          geo = "Geographic Focus",
          keyword1 = "Keyword1", 
          keyword2 = "Keyword2", 
          keyword3 = "Keyword3", 
          org = "Affiliation")
```

```{r eval=FALSE}
# Create a corpus object from the abstracts
general_corpus <- tm::Corpus(tm::DataframeSource(general), 
                             readerControl = list(reader = tm::readTabular(mapping = m)))

# create a document term matrix
general_tm <- tm::DocumentTermMatrix(general_corpus, 
                                     control = list(stemming = TRUE, 
                                                    stopwords = TRUE,
                                                    minWordLength = 2, 
                                                    removeNumbers = TRUE, 
                                                    removePunctuation = TRUE))
```

```{r eval=FALSE}
# Use TF-IDF to wieght terms and remove rare ones
term_tfidf <- 
  tapply(general_tm$v/slam::row_sums(general_tm)[general_tm$i], 
         general_tm$j, mean) * log2(tm::nDocs(general_tm)/slam::col_sums(general_tm > 0))

summary(term_tfidf)
# Median =  0.08989
```

```{r eval=FALSE }
## Keeping the rows with tfidf >= median
general_reduced_dtm <- general_tm[,term_tfidf >= summary(term_tfidf)[3]]
summary(slam::col_sums(general_reduced_dtm))
# save
saveRDS(general_reduced_dtm, "../data/derived_data/general_reduced_dtm.rds")
```

```{r}
general_reduced_dtm <- readRDS("../data/derived_data/general_reduced_dtm.rds")
```

```{r}
# Determine k number of topics
harmonicMean <- function(logLikelihoods, precision = 2000L) {
  llMed <- median(logLikelihoods)
  as.double(llMed - log(Rmpfr::mean(exp(-Rmpfr::mpfr(logLikelihoods,
                                       prec = precision) + llMed))))
}
```

```{r}
# run a loop over the abstract with different numbers of topics to find which number of topcis is the best
seqk <- seq(2, 100, 1)
burnin <- 1000
iter <- 1000
keep <- 50
```

```{r eval=FALSE }
system.time(fitted_many <- 
              lapply(seqk, function(k) topicmodels::LDA(general_reduced_dtm, 
                                                        k = k,
                                                        method = "Gibbs",
                                                        control = list(burnin = burnin,
                                                                       iter = iter, 
                                                                       keep = keep) )))

saveRDS(fitted_many, "../data/derived_data/fitted_many.rds")
```

```{r}
fitted_many <- readRDS("../data/derived_data/fitted_many.rds")
```

```{r}
# extract logliks from each topic
logLiks_many <- lapply(fitted_many, function(L)  L@logLiks[-c(1:(burnin/keep))])

# compute harmonic means
hm_many <- sapply(logLiks_many, function(h) harmonicMean(h))

# to know the optimal number of topics:
optimal_number_of_topics <- seqk[which.max(hm_many)]

# plot it
library(ggplot2)
lda_plot <- ggplot(data.frame(seqk, 
                             hm_many), 
                  aes(x=seqk, 
                      y=hm_many)) + 
  geom_path(lwd=1.5)  +
  xlab('Number of Topics') +
  ylab('Harmonic Mean') +
  geom_vline(xintercept = seqk[which.max(hm_many)],
             colour = "red") +
  annotate("text",
           x = 55, 
           y = -150000, 
           label = paste("The optimal number of topics is", 
                         seqk[which.max(hm_many)])) +
  theme_bw() +
  ggtitle(expression(atop("Latent Dirichlet Allocation Analysis of SAA General Abstracts", atop(italic("How many distinct topics in the abstracts?"), ""))))
```

```{r}
# now run the model with the optimum number of topics
system.time(general_model <- topicmodels::LDA(general_reduced_dtm, 
                                              optimal_number_of_topics,
                                              method = "Gibbs", 
                                              control = list(iter=2000, 
                                                             seed = 0622)))

# explore the model
general_topics <- topicmodels::topics(general_model, 1)

## In this case I am returning the top 30 terms.
general_terms <- as.data.frame(topicmodels::terms(general_model, 30), 
                               stringsAsFactors = FALSE)
```

```{r}
library(tidytext)
tidy_topics <- tidy(general_model, matrix = "beta")

library(ggplot2)
library(dplyr)

tidy_topics_top_terms <- tidy_topics %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)
```

```{r}
tidy_topics_top_terms %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()
```

```{r}
tidy_documents <- tidy(general_model, matrix = "gamma")

library(tidyr)
tidy_documents_topics_gamma <-
  tidy_documents %>% 
    spread(topic, gamma)

abstract_classifications <- 
  tidy_documents %>%
  group_by(document) %>%
  top_n(1, gamma) %>%
  ungroup() %>% 
  arrange(desc(gamma))

abstract_classifications
```

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```

### 2017-07-29 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(saaabstracts)
+
+test_check("saaabstracts")

### 2017-09-14 analysis/paper/paper.Rmd
add 2018 general sessions

@@
   arrange(desc(gamma))
+```
 
-abstract_classifications
+```{r eval=FALSE}
+mydata <- tidy_documents_topics_gamma[,-1]
+
+# Determine number of clusters
+library(NbClust)
+nclust <- NbClust(mydata, method = "kmeans")
+
+library(fpc)
+pamk.best <- pamk(mydata)
+
+# Determine number of clusters
+wss <- (nrow(mydata)-1)*sum(apply(mydata,2,var))
+for (i in 2:15) wss[i] <- sum(kmeans(mydata, 
+  	centers=i)$withinss)
+plot(1:15, wss, type="b", xlab="Number of Clusters",
+  ylab="Within groups sum of squares")
+
+# K-Means Cluster Analysis
+fit <- kmeans(mydata, 5) # 5 cluster solution
+# get cluster means 
+aggregate(mydata,by=list(fit$cluster),FUN=mean)
+# append cluster assignment
+mydata <- data.frame(mydata, fit$cluster)
 ```
+
+## Wokring in a local Docker container
+
+To run this project in a local Docker container, I start a bash shell in the project directory:
+
+Do this one time only, build my Docker container from Dockerfile at top-level in my project:
+
+```
+docker build - < Dockerfile
+docker ps
+```
+
+Take note of the container ID, which you get after `docker ps`
+
+Then run my Docker container, access it via the web browser, link the project dir to the Docker drive. Make sure you're in a local directory that is sharable (C:\Users\yourname is usually good):
+
+```
+docker run -dp 8787:8787 -v ${pwd}:/home/rstudio/ -e ROOT=TRUE  <container ID>
+```
+
+Then go to `http://192.168.99.100:8787/` or `localhost:8787` in youor browser, log in (rstudio/rstudio), and start your RStudio project by double-clickin on the .Rproj file
+
+When you're done, in the shell, stop all docker containers with:
+
+```
+docker stop $(docker ps -a -q)
+```
+
+Or do `docker ps` and `docker stop <container ID>` just to stop one container.
 

### 2017-09-14 analysis/paper/saa2018_general_sessions.Rmd
make new Rmd for 2018 general sessions

--- 2017-07-29 analysis/paper/paper.Rmd
@@
                                                     minWordLength = 2, 
-                                                    removeNumbers = TRUE, 
+                      s                              removeNumbers = TRUE, 
                                                     removePunctuation = TRUE))

### 2017-09-14 analysis/paper/saa2018_general_sessions.Rmd
optimum number of topics for SAA2018 general abstracts

--- 2017-07-29 analysis/paper/paper.Rmd
@@
 # read in the abstracts
-general <- read_excel("../data/raw_data/General Abstracts.xlsx")
+general_2018 <- read_excel("../data/raw_data/General Abstracts_2018.xlsx")
 ```
@@
 # Create a corpus object from the abstracts
-general_corpus <- tm::Corpus(tm::DataframeSource(general), 
+general_2018_corpus <- tm::Corpus(tm::DataframeSource(general_2018), 
                              readerControl = list(reader = tm::readTabular(mapping = m)))
@@
 # create a document term matrix
-general_tm <- tm::DocumentTermMatrix(general_corpus, 
+general_2018_tm <- tm::DocumentTermMatrix(general_2018_corpus, 
                                      control = list(stemming = TRUE, 
@@
                                                     removePunctuation = TRUE))
+
 ```
@@
 # Use TF-IDF to wieght terms and remove rare ones
-term_tfidf <- 
-  tapply(general_tm$v/slam::row_sums(general_tm)[general_tm$i], 
-         general_tm$j, mean) * log2(tm::nDocs(general_tm)/slam::col_sums(general_tm > 0))
+term_2018_tfidf <- 
+  tapply(general_2018_tm$v/slam::row_sums(general_2018_tm)[general_2018_tm$i], 
+         general_2018_tm$j, mean) * log2(tm::nDocs(general_2018_tm)/slam::col_sums(general_2018_tm > 0))
 
-summary(term_tfidf)
-# Median =  0.08989
+summary(term_2018_tfidf)
+# Median =   0.08769
 ```
@@
 ## Keeping the rows with tfidf >= median
-general_reduced_dtm <- general_tm[,term_tfidf >= summary(term_tfidf)[3]]
-summary(slam::col_sums(general_reduced_dtm))
+general_2018_reduced_dtm <- general_2018_tm[,term_2018_tfidf >= summary(term_2018_tfidf)[3]]
+summary(slam::col_sums(general_2018_reduced_dtm))
 # save
-saveRDS(general_reduced_dtm, "../data/derived_data/general_reduced_dtm.rds")
+saveRDS(general_2018_reduced_dtm, "../data/derived_data/general_2018_reduced_dtm.rds")
 ```
@@
 ```{r}
-general_reduced_dtm <- readRDS("../data/derived_data/general_reduced_dtm.rds")
+general_2018_reduced_dtm <- readRDS("../data/derived_data/general_2018_reduced_dtm.rds")
+number_of_abstracts <- general_2018_reduced_dtm$nrow
 ```
@@
 ```{r eval=FALSE }
-system.time(fitted_many <- 
-              lapply(seqk, function(k) topicmodels::LDA(general_reduced_dtm, 
+system.time(fitted_many_2018 <- 
+              lapply(seqk, function(k) topicmodels::LDA(general_2018_reduced_dtm, 
                                                         k = k,
@@
 
-saveRDS(fitted_many, "../data/derived_data/fitted_many.rds")
+saveRDS(fitted_many_2018, "../data/derived_data/fitted_many_2018.rds")
 ```
@@
 ```{r}
-fitted_many <- readRDS("../data/derived_data/fitted_many.rds")
+fitted_many_2018 <- readRDS("../data/derived_data/fitted_many_2018.rds")
+
+# how many abstracts, if we just start from here?
+how_many_abstracts <- fitted_many_2018[[1]]@Dim[1]
 ```
@@
 # extract logliks from each topic
-logLiks_many <- lapply(fitted_many, function(L)  L@logLiks[-c(1:(burnin/keep))])
+logLiks_many_2018 <- lapply(fitted_many_2018, function(L)  L@logLiks[-c(1:(burnin/keep))])
 
 # compute harmonic means
-hm_many <- sapply(logLiks_many, function(h) harmonicMean(h))
+hm_many_2018 <- sapply(logLiks_many_2018, function(h) harmonicMean(h))
 
 # to know the optimal number of topics:
-optimal_number_of_topics <- seqk[which.max(hm_many)]
+optimal_number_of_topics_2018 <- seqk[which.max(hm_many_2018)]
 
@@
 lda_plot <- ggplot(data.frame(seqk, 
-                             hm_many), 
+                             hm_many_2018), 
                   aes(x=seqk, 
-                      y=hm_many)) + 
+                      y=hm_many_2018)) + 
   geom_path(lwd=1.5)  +
@@
   ylab('Harmonic Mean') +
-  geom_vline(xintercept = seqk[which.max(hm_many)],
+  geom_vline(xintercept = seqk[which.max(hm_many_2018)],
              colour = "red") +
@@
            label = paste("The optimal number of topics is", 
-                         seqk[which.max(hm_many)])) +
+                         seqk[which.max(hm_many_2018)])) +
   theme_bw() +
-  ggtitle(expression(atop("Latent Dirichlet Allocation Analysis of SAA General Abstracts", atop(italic("How many distinct topics in the abstracts?"), ""))))
+  ggtitle("Latent Dirichlet Allocation Analysis of SAA General Abstracts",
+          subtitle = paste0("How many distinct topics are present in the general abstracts (n = ", how_many_abstracts, ") for 2018?"))
 ```

### 2017-09-14 analysis/paper/saa2018_general_sessions.Rmd
keywords per topic plot

@@
 ```{r}
-# run a loop over the abstract with different numbers of topics to find which number of topcis is the best
+# run a loop over the abstract with different numbers of topics to find which number of topics is the best
 seqk <- seq(2, 100, 1)
@@
   ylab('Harmonic Mean') +
-  geom_vline(xintercept = seqk[which.max(hm_many_2018)],
+  geom_vline(xintercept = optimal_number_of_topics_2018,
              colour = "red") +
@@
            label = paste("The optimal number of topics is", 
-                         seqk[which.max(hm_many_2018)])) +
+                         optimal_number_of_topics_2018)) +
   theme_bw() +
@@
           subtitle = paste0("How many distinct topics are present in the general abstracts (n = ", how_many_abstracts, ") for 2018?"))
+
+ggsave("../figures/saa2018_general_abstracts_opti_topics.png")
 ```
@@
 # now run the model with the optimum number of topics
-system.time(general_model <- topicmodels::LDA(general_reduced_dtm, 
-                                              optimal_number_of_topics,
+system.time(general_model <- topicmodels::LDA(general_2018_reduced_dtm, 
+                                              optimal_number_of_topics_2018,
                                               method = "Gibbs", 
@@
 ```{r}
+# show top keywords for each topic
 tidy_topics_top_terms %>%
@@
   coord_flip()
+
+ggsave("../figures/keywords_for_saa2018_general_topics.png", h = 14, w = 14)
 ```

### 2017-09-15 analysis/paper/saa2018_general_sessions_hclust.Rmd
hclust on keywords looking good

---
title: "Title Goes Here"
author:
  - author 1
  - author 2
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
bibliography: references.bib
csl: journal-of-archaeological-science.csl
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "figures"
)

library(saaabstracts)
library(tm)
library(topicmodels)
library(readxl)
library(Rmpfr)
library(tidyverse)
```

# Introduction

```{r eval=TRUE}
# read in the abstracts
general_2018 <- read_excel("../data/raw_data/General Abstracts_2018.xlsx")

general_2018_posters <- general_2018 %>% filter(grepl("Poster", `Presentation Format`))
general_2018_papers  <- general_2018 %>% filter(grepl("Paper", `Presentation Format`))

# clean geo
clean_geo <- function(x){
  tmp <- stringr::str_trim(tolower(gsub("[[:punct:]]|\r\n| and ", " ", x)))
  tmp <- gsub(" ", "_", tmp)
  tmp <- gsub("_{2}", "_", tmp)
  tmp
}

general_2018_papers$geo <- 
  unlist(purrr::map(general_2018_papers$`Geographic Focus`, ~clean_geo(.x)))

general_2018_papers$kw1 <- 
  unlist(purrr::map(general_2018_papers$Keyword1, ~clean_geo(.x)))

general_2018_papers$kw2 <- 
  unlist(purrr::map(general_2018_papers$Keyword2, ~clean_geo(.x)))

general_2018_papers$kw3 <- 
  unlist(purrr::map(general_2018_papers$Keyword3, ~clean_geo(.x)))

general_2018_papers$text_to_model <- 
  with(general_2018_papers, paste(geo, kw1, kw2, kw3))
```

```{r}
# clean text to model
general_2018_papers$text_to_model  <- 
  stringr::str_trim(tolower(gsub("\\/|:|&", " ", general_2018_papers$text_to_model)))

general_2018_papers$text_to_model  <- 
  gsub(" and | archaeology | analysis |\r\n", " ", general_2018_papers$text_to_model)

```

```{r}
# Create a corpus object from the text to mode

general_2018_papers_corpus <- tm::Corpus(tm::VectorSource(general_2018_papers$text_to_model ))

ndocs <- length(general_2018_papers_corpus)
# ignore extremely rare words i.e. terms that appear in less then 1% of the documents
minTermFreq <- ndocs * 0.01
# ignore overly common words i.e. terms that appear in more than 50% of the documents
maxTermFreq <- ndocs * .5

# create a document term matrix
general_2018_papers_tm <- tm::DocumentTermMatrix(general_2018_papers_corpus, 
                                     control = list(stemming = TRUE, 
                                                    stopwords = TRUE,
                                                     wordLengths=c(4,30),  
                                                    removeNumbers = TRUE, 
                                                     bounds = list(global = c(minTermFreq, maxTermFreq)),
                                                    removePunctuation = TRUE))
```

```{r}
m  <- as.matrix(general_2018_papers_tm)
distMatrix <- dist(m, method="euclidean")

groups <- hclust(distMatrix,method="ward.D")
plot(groups, cex=0.9, hang=-1)
rect.hclust(groups, k=70)

# see what cluster each abstract is in:
cuts <- cutree(groups, k = 50)
```

```{r}
# put topics back onto abstract spreadsheet
# abstract_ids <- dimnames(general_2018_reduced_dtm)$Docs
# general_2018$`Abstract Id`

general_2018_papers_with_topics <- 
  bind_cols(general_2018_papers, data_frame(cuts))

write.csv(general_2018_papers_with_topics, "../data/derived_data/general_2018_papers_with_clusters.csv")
```

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```

### 2017-09-19 analysis/paper/paper.Rmd
what does word2vec do?

@@
 general_reduced_dtm <- readRDS("../data/derived_data/general_reduced_dtm.rds")
+```
+
+```{r}
+library(ldatuning)
+result <- FindTopicsNumber(
+  general_reduced_dtm,
+  topics = seq(from = 2, to = 100, by = 2),
+  metrics = c("Griffiths2004", "CaoJuan2009", "Arun2010", "Deveaud2014"),
+  method = "Gibbs",
+  control = list(seed = 77),
+  mc.cores = 2L,
+  verbose = TRUE
+)
+
+knitr::kable(result)
+FindTopicsNumber_plot(result)
 ```

### 2017-09-19 analysis/paper/saa2018_general_sessions_word2vec.Rmd
what does word2vec do?

--- 2017-09-15 analysis/paper/saa2018_general_sessions_hclust.Rmd
@@
 ```{r}
+library(wordVectors)
+library(magrittr)
+library(stringr)
+
+# Write abstracts out to text files
+lapply(1:length(general_2018_papers$Abstract),
+     function(i) write.table(tolower(str_trim(general_2018_papers$Abstract[i])),
+                  file = paste0("abstracts/", general_2018_papers$`Abstract Id`[i], ".txt"),
+                  quote = FALSE, 
+                  row.names = FALSE,
+                  col.names = FALSE))
+
+length(list.files("abstracts/")) #178
+
+prep_word2vec(origin = "abstracts/",
+              destination = "abstracts/abstracts.txt",
+              lowercase = TRUE,
+              bundle_ngrams = 2)
+
+model = train_word2vec("abstracts/abstracts.txt","abstracts/abstracts.bin",
+                       vectors=200,
+                       threads=4,
+                       window=12,
+                       iter=5,
+                       negative_samples=0)
+
+# model = read.vectors("cookbook_vectors.bin")
+
+plot(model,perplexity=50)
+
+set.seed(10)
+centers = 50
+clustering = kmeans(model,
+                    centers = centers,
+                    iter.max = 40)
+
+sapply(sample(1:centers,10),function(n) {
+  names(clustering$cluster[clustering$cluster==n][1:10])
+})
+
+```
+
+```{r}
+library(stringr)
+library(text2vec)
+
+itn <- map(general_2018_papers$Abstract, ~itoken(.x))
+
+it = itoken(general_2018_papers$Abstract, progressbar = FALSE)
+v = create_vocabulary(it) %>% prune_vocabulary(doc_proportion_max = 0.1, term_count_min = 5)
+vectorizer = vocab_vectorizer(v)
+
+# Jaccard similarity
+# they will be in the same space because we use same vectorizer
+# hash_vectorizer will also work fine
+dtm1 = create_dtm(it1, vectorizer)
+
+dtm2 = create_dtm(it2, vectorizer)
+
+# Once we have representation of documents in vector space we are almost done. One thing remains - call sim2():
+
+d1_d2_jac_sim = sim2(dtm1, dtm2, method = "jaccard", norm = "none")
+# Check result:
+
+dim(d1_d2_jac_sim)
+## [1] 300 200
+d1_d2_jac_sim[1:2, 1:5]
+```
+
+```{r}
+library(stringr)
+library(text2vec)
+data("movie_review")
+# select 500 rows for faster running times
+movie_review = movie_review[1:500, ]
+prep_fun = function(x) {
+  x %>% 
+    # make text lower case
+    str_to_lower %>% 
+    # remove non-alphanumeric symbols
+    str_replace_all("[^[:alnum:]]", " ") %>% 
+    # collapse multiple spaces
+    str_replace_all("\\s+", " ")
+}
+movie_review$review_clean = prep_fun(movie_review$review)
+
+doc_set_1 = movie_review[1:300, ]
+it1 = itoken(doc_set_1$review_clean, progressbar = FALSE)
+
+# specially take different number of docs in second set
+doc_set_2 = movie_review[301:500, ]
+it2 = itoken(doc_set_2$review_clean, progressbar = FALSE)
+
+it = itoken(movie_review$review_clean, progressbar = FALSE)
+v = create_vocabulary(it) %>% prune_vocabulary(doc_proportion_max = 0.1, term_count_min = 5)
+vectorizer = vocab_vectorizer(v)
+
+# Jaccard similarity
+# they will be in the same space because we use same vectorizer
+# hash_vectorizer will also work fine
+dtm1 = create_dtm(it1, vectorizer)
+dim(dtm1)
+## [1]  300 2338
+dtm2 = create_dtm(it2, vectorizer)
+dim(dtm2)
+## [1]  200 2338
+# Once we have representation of documents in vector space we are almost done. One thing remains - call sim2():
+
+d1_d2_jac_sim = sim2(dtm1, dtm2, method = "jaccard", norm = "none")
+# Check result:
+
+dim(d1_d2_jac_sim)
+## [1] 300 200
+d1_d2_jac_sim[1:2, 1:5]
+
+```
+
+```{r}
 m  <- as.matrix(general_2018_papers_tm)

### 2017-09-19 analysis/paper/saa2018_general_sessions_posters_hclust.Rmd
sent to AB, thinking about min and max cluster sizes...

---
title: "Title Goes Here"
author:
  - author 1
  - author 2
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
bibliography: references.bib
csl: journal-of-archaeological-science.csl
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "figures"
)

library(saaabstracts)
library(tm)
library(topicmodels)
library(readxl)
library(Rmpfr)
library(tidyverse)
```

# Introduction

```{r eval=TRUE}
# read in the abstracts
general_2018 <- read_excel("../data/raw_data/General Abstracts_2018.xlsx")

general_2018_posters <- general_2018 %>% filter(grepl("Poster", `Presentation Format`))
general_2018_papers  <- general_2018 %>% filter(grepl("Paper", `Presentation Format`))

# clean geo
clean_geo <- function(x){
  tmp <- stringr::str_trim(tolower(gsub("[[:punct:]]|\r\n| and ", " ", x)))
  tmp <- gsub("other", "", tmp)
  tmp <- gsub(" ", "_", tmp)
  tmp <- gsub("_{2}", "_", tmp)
  tmp
}

general_2018_posters$geo <- 
  unlist(purrr::map(general_2018_posters$`Geographic Focus`, ~clean_geo(.x)))

general_2018_posters$kw1 <- 
  unlist(purrr::map(general_2018_posters$Keyword1, ~clean_geo(.x)))

general_2018_posters$kw2 <- 
  unlist(purrr::map(general_2018_posters$Keyword2, ~clean_geo(.x)))

general_2018_posters$kw3 <- 
  unlist(purrr::map(general_2018_posters$Keyword3, ~clean_geo(.x)))

general_2018_posters$text_to_model <- 
  with(general_2018_posters, paste(geo, kw1, kw2, kw3))
```

```{r}
# clean text to model
general_2018_posters$text_to_model  <- 
  stringr::str_trim(tolower(gsub("\\/|:|&", " ", general_2018_posters$text_to_model)))

general_2018_posters$text_to_model  <- 
  gsub(" and | archaeology | analysis |\r\n", " ", general_2018_posters$text_to_model)

```

```{r}
# Create a corpus object from the text to mode

general_2018_posters_corpus <- tm::Corpus(tm::VectorSource(general_2018_posters$text_to_model ))

ndocs <- length(general_2018_posters_corpus)
# ignore extremely rare words i.e. terms that appear in less then 1% of the documents
minTermFreq <- ndocs * 0.01
# ignore overly common words i.e. terms that appear in more than 50% of the documents
maxTermFreq <- ndocs * 0.5

# create a document term matrix
general_2018_posters_tm <- tm::DocumentTermMatrix(general_2018_posters_corpus, 
                                     control = list(stemming = TRUE, 
                                                    stopwords = TRUE,
                                                    wordLengths=c(4,30),  
                                                    removeNumbers = TRUE, 
                                                    bounds = list(global = c(minTermFreq, maxTermFreq)),
                                                    removePunctuation = TRUE))
```

Distance matrix

```{r}
m  <- as.matrix(general_2018_posters_tm)
distMatrix <- dist(m, method="euclidean")
```

How many clusters?

```{r}
library(apcluster)
d.apclus <- apcluster(negDistMat(r=2), m)
k <- length(d.apclus@clusters)

windows()
heatmap(d.apclus)
```

Cluster analysis and dendrogram

```{r}
groups <- hclust(distMatrix,method="ward.D")
plot(groups, cex=0.9, hang=-1)
rect.hclust(groups, k)

# see what cluster each abstract is in:
cuts <- cutree(groups, k)
```

```{r}
# Make clustering with `sc_clustering`
library(scclust)
my_dist <- distances(m)
my_clustering <- sc_clustering(my_dist, 3)

# Refine clustering with `hierarchical_clustering`
my_refined_clustering <- hierarchical_clustering(my_dist,
                                                 size_constraint = 4,
                                                 existing_clustering = my_clustering)

cuts <- my_refined_clustering 
```

```{r}
hc <- groups

unnest <- function(x) { # from Vlo's answer
  if(is.null(names(x))) x
  else c(list(all=unname(unlist(x))), do.call(c, lapply(x, unnest)))
}

cuts <- hc$height + 1e-9

min_size <- 10
smallest <- 4
i <- 0

while(smallest < min_size & i <= length(cuts)){
  h_i <- cuts[i <- i+1]
  if(i > length(cuts)){
    warning("Couldn't find a cluster big enough.")
  }
  else  smallest <- 
           Reduce(min, 
                  lapply(X = unnest(cut(as.dendrogram(hc), h=h_i)$lower), 
                         FUN = attr, which = "members") ) # from lukeA's comment
}
h_i

cut(as.dendrogram(hc), h=h_i/2)
```

How many items in each cluster

```{r}
# histogram
freqs_per_cluster <- 
cuts %>%
  data_frame(cuts = .,
             freq = names(.)) %>% 
  group_by(cuts) %>% 
  tally()

hist(freqs_per_cluster$n)

median(freqs_per_cluster$n) # 10
```

```{r}
# put topics back onto abstract spreadsheet
# abstract_ids <- dimnames(general_2018_reduced_dtm)$Docs
# general_2018$`Abstract Id`

general_2018_posters_with_topics <- 
  bind_cols(general_2018_posters, data_frame(cuts))

write.csv(general_2018_posters_with_topics, "../data/derived_data/general_2018_posters_with_clusters.csv")
```

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```

### 2017-09-20 analysis/paper/saa2018_general_sessions_posters_hclust.Rmd
thinking about scheduling

@@
 ```{r}
-# Make clustering with `sc_clustering`
-library(scclust)
-my_dist <- distances(m)
-my_clustering <- sc_clustering(my_dist, 3)
+library(dynamicTreeCut)
+maxCoreScatter <- 0.99
+minGap <- (1 - maxCoreScatter) * 0.01
+dynamicCut <- cutreeDynamic(groups, 
+                           minClusterSize=4, 
+                           method="hybrid", 
+                           distM=as.matrix(dist(m, method="euclidean")), 
+                           deepSplit=4, 
+                           maxCoreScatter=maxCoreScatter, 
+                           maxAbsCoreScatter=NULL, 
+                           minAbsGap=NULL)
 
-# Refine clustering with `hierarchical_clustering`
-my_refined_clustering <- hierarchical_clustering(my_dist,
-                                                 size_constraint = 4,
-                                                 existing_clustering = my_clustering)
+hist(table(dynamicCut))
 
-cuts <- my_refined_clustering 
-```
+xfreqs_per_cluster <- 
+dynamicCut %>%
+  data_frame(dynamicCut = .,
+             freq = names(.)) %>% 
+  group_by(dynamicCut) %>% 
+  tally(sort=TRUE)
 
-```{r}
+hist(xfreqs_per_cluster$n)
+
+library(dendextend)
 hc <- groups
+dend <- as.dendrogram(hc)
+heights_per_k.dendrogram(dend)
 
@@
 
-cuts <- hc$height + 1e-9
+hc <- groups
+cuts2 <- hc$height + 1e-9
+# my experiment
+for(i in seq_along(cuts2)){
+  n <- attr(cut(as.dendrogram(groups), h=3)$lower[[i]], "members")
+  print(n)
+}
 
-min_size <- 10
-smallest <- 4
+my_list <- cut(as.dendrogram(groups), h=3)
+map(my_list, unnest(.x)$lower)
+
+map(cut(as.dendrogram(groups), h=3), ~attr(.x$lower, "members"))
+
+xcuts <- cutree(groups, k)
+
+xfreqs_per_cluster <- 
+xcuts %>%
+  data_frame(cuts = .,
+             freq = names(.)) %>% 
+  group_by(cuts) %>% 
+  tally(sort=TRUE)
+
+xmax <- max(xfreqs_per_cluster$n)
+xmin <- min(xfreqs_per_cluster$n)
+
+max_size <- 16
+min_size <- 4
+n <- 0
+
+while(xmax <= max_size &  xmin >= min_size & n <= length(xcuts)){
+  n <-  n + 1
+  ycuts <- cutree(groups, n)
+
+xfreqs_per_cluster <- 
+xcuts %>%
+  data_frame(cuts = .,
+             freq = names(.)) %>% 
+  group_by(cuts) %>% 
+  tally(sort=TRUE)
+
+xmax <- max(xfreqs_per_cluster$n)
+xmin <- min(xfreqs_per_cluster$n)
+
+  xcuts <- cutree(groups, n)
+
+}
+
+```
+
+```{r}
+
+hc <- groups
+
+hc$size
+
+unnest <- function(x) { # from Vlo's answer
+  if(is.null(names(x))) x
+  else c(list(all=unname(unlist(x))), do.call(c, lapply(x, unnest)))
+}
+
+cuts2 <- hc$height + 1e-9
+
+biggest <- 50
+smallest <- 0
+
+max_size <- 16
+min_size <- 4
+
 i <- 0
 
-while(smallest < min_size & i <= length(cuts)){
-  h_i <- cuts[i <- i+1]
-  if(i > length(cuts)){
+while(smallest < min_size & 
+      biggest > max_size &
+      i <= length(cuts2)){
+  h_i <- cuts2[i <- i+1]
+  if(i > length(cuts2)){
     warning("Couldn't find a cluster big enough.")
@@
                          FUN = attr, which = "members") ) # from lukeA's comment
+
+  biggest <- 
+           Reduce(max, 
+                  lapply(X = unnest(cut(as.dendrogram(hc), h=h_i)$upper), 
+                         FUN = attr, which = "members") ) # from lukeA's comment
+
+  print(paste0("biggest: ", biggest, ", smallest: ", smallest, " i: ", i, " h_i: ", h_i))
 }
@@
 
-cut(as.dendrogram(hc), h=h_i/2)
+cut(as.dendrogram(hc), h=h_i)
 ```
@@
   group_by(cuts) %>% 
-  tally()
+  tally(sort=TRUE)
 

### 2017-10-26 analysis/paper/paper.Rmd
get the knapsack algo working

--- 2017-09-14 analysis/paper/paper.Rmd
@@
 
+## Knapsack
+
+We have how many sessions?
+
+```{r}
+library(readxl)
+org_sessions <- read_excel("../data/raw_data/Organized Session Abstracts_2018_AB.xlsx")
+
+gen_sessions <- read_excel("../data/raw_data/01 - gen_paper_sessions_FINAL(oct5)_DG_including_moves v2.xlsx")
+
+# how many unique organised sessions? 
+n_org_sessions <- length(unique(org_sessions$`Session Id`))
+
+# how many unique general sessions? 
+n_gen_sessions <- length(unique(gen_sessions$SESSION))
+
+# total number of sessions
+total_number_of_sessions <-  n_org_sessions + n_gen_sessions
+```
+
+For sheduling into rooms, we don't include the posters, because they go in the ballroom.
+How many 4h slots do we have?
+I looked on the grid on 'Rooom&Time_DG.xlsx' and counted up the 4h slots, excluding poster sessions, because they are less than 4 h
+
+```{r}
+library(tidyverse)
+four_h_slots <- 
+  tribble(~time,       ~Th, ~Fr, ~Sa, ~Su,
+          "morning",    36,  31,  30,  28,
+          "afternoon",  35,  32,  30,  0,
+          "evening",    36,  0,    0,  0)
+
+four_h_slots_long <- 
+  four_h_slots %>% 
+  gather(times, slot, -time)
+
+# how many 4 h slots
+n_four_h_slots <-  sum(four_h_slots_long$slot)
+```
+
+So that's `r n_four_h_slots` 4 hour slots, where we consider each slot a knapsack.
+
+Now we need to pack the sessions into these slots.
+First we need to set the upper bounds for the session durations, according to the SAA rules, and then compute how much time each session is likely to take give the number of papers/speakers
+
+```{r}
+# combine organised and general sessions
+org_sessions$`Session Id` <- as.character(org_sessions$`Session Id`)
+gen_sessions$`Session Id` <- gen_sessions$SESSION
+gen_sessions$`Session Type` <- "Symposium"
+
+gen_sessions_to_join <- 
+gen_sessions %>% 
+  select(names(gen_sessions)[names(gen_sessions) %in% names(org_sessions)])
+
+all_nonposter_sessions <- 
+  bind_rows(gen_sessions_to_join, org_sessions)
+
+# how many papers per session
+papers_per_session <- 
+all_nonposter_sessions %>% 
+  group_by(`Session Id`) %>% 
+  tally(sort = TRUE)
+
+ggplot(papers_per_session,
+       aes(n)) +
+  geom_histogram() +
+  ylab("Papers per non-poster session") +
+  theme_bw()
+
+# what's the duration of each session, and he max allowed duration? 
+session_types <- unique(all_nonposter_sessions$`Session Type`)
+
+# this comes from 'Conflicts.docx'
+session_time_limits <- 
+  tribble(~`Session Type`,             ~max_time,
+          "Poster Symposium",           2,
+          "Symposium",                  4,
+          "Electronic Symposium",       2, 
+          "Forum" ,                     2, 
+          "Lightning Rounds",           2)
+
+# join time limits onto session data
+session_with_limits <- 
+  all_nonposter_sessions %>% 
+  filter(`Session Type` != "Poster Symposium") %>% 
+  left_join(session_time_limits)
+
+# now let's see which session exceed the time limits
+session_with_limits <- 
+session_with_limits %>% 
+  group_by(`Session Id`, `Session Type`, max_time ) %>% 
+  tally(sort = TRUE) %>% 
+  mutate(paper_time = n * 15/60) %>% 
+  mutate(paper_time = if_else(paper_time > max_time,  max_time, paper_time))
+
+```
+
+Now we have a known number of 4 hour sessions (knapsacks), and a known number of sessions of various lengths.
+We turn to http://www.sumsar.net/blog/2016/06/how-to-cut-your-planks-with-r/ to see how to arrange them most efficiently.
+Here is the original knapsack example that we'll try to adapt:
+
+```{r}
+# Here is Rasmus' example: http://www.sumsar.net/blog/2016/06/how-to-cut-your-planks-with-r/
+
+planks_we_have <- c(120, 137, 220, 420, 480)
+planks_we_want <- c(19, 19, 19, 19, 79, 79, 79, 103, 103,
+                    103, 135, 135, 135, 135, 160)
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(planks_we_want, planks_we_want + 1, planks_we_have)
+# Above I added +1 cm  to each length to compensate for the loss when sawing.
+solution$ksack
+
+# That is, cut plank length 1 from plank 1, plank length 2 from plank 4, etc.
+
+# Now pretty printing what to cut so that we don't make mistakes...
+assignment <- data.frame(
+  cut_this = planks_we_have[solution$ksack],
+  into_this = planks_we_want)
+t(assignment[order(assignment[,1]), ])
+```
+
+Now let's substitute with our data:
+
+```{r}
+# Convert the durations to minutes so we get integers, mknapsack doesn't work with decimals
+
+# into these
+sessions_we_have <- rep(4, n_four_h_slots) * 60 # so many 4 hour slots that we need to fill 
+# we want to fit these
+sessions_we_want <- session_with_limits$paper_time * 60 # the duration of each session
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_want, sessions_we_want, sessions_we_have)
+
+solution$ksack
+# That is, put session 1 into slot 1, session 2 into slot 2, etc.
+
+# Now pretty printing what to session to put in what slot so that we don't make mistakes...
+assignment <- data.frame(
+  fill_this_slot = sessions_we_have[solution$ksack],
+  with_this_session = sessions_we_want)
+t(assignment[order(assignment[,1]), ])
+
+# put the computed slot back onto the full details of the sessions
+session_with_limits$computed_slot <- solution$ksack
+```
+
+That seems to work great, assuming that we've got the time limits ok. Now we arbitrarily assign them to the schedule, then check for conflicts
+
+```{r}
+four_h_slots_long$cumsum <- cumsum(four_h_slots_long$slot)
+
+xx <- 
+four_h_slots_long %>% 
+  mutate(day_part = paste0(times, "-", time)) %>% 
+  select(-time, -times) %>% 
+  rename(end = cumsum) %>% 
+  filter( slot != 0)
+
+# get the number of the start slot to go with the number of the end slot
+for(i in 1:(nrow(xx))) {
+
+  xx$start[i] <- 
+    if_else(i == 1, 
+           1,
+            as.numeric(xx[ i-1, 2]) + 1)
+
+}
+
+intervals <- 
+xx %>% 
+  select(-slot) %>% 
+  select(day_part, start, end )
+
+elements <- session_with_limits$computed_slot
+
+library(fuzzyjoin)
+join_intervals_and_elements <- 
+fuzzy_left_join(data.frame(elements), 
+                intervals, 
+                by = c("elements" = "start", 
+                       "elements" = "end"), 
+                match_fun = list(`>=`, `<=`)) %>% 
+  distinct()
+
+# Which slot goes into which part of what day
+slot_day_part <- 
+join_intervals_and_elements %>% 
+  arrange(elements, day_part)
+
+# now we put the sessions on also
+session_slot_day_part <- 
+  session_with_limits %>% 
+  left_join(slot_day_part, by  = c("computed_slot" = "elements"))
+
+```
+
+sessions =
+
 ## Wokring in a local Docker container

### 2017-10-26 analysis/paper/paper.Rmd

@@
 
-sessions =
-
-## Wokring in a local Docker container
+## Working in a local Docker container
 

### 2017-10-26 analysis/paper/saa2018_general_sessions.Rmd

--- 2017-09-14 analysis/paper/saa2018_general_sessions.Rmd
@@
 
-```{r eval=FALSE}
+```{r eval=TRUE}
 # read in the abstracts
-general <- read_excel("../data/raw_data/General Abstracts.xlsx")
+general_2018 <- read_excel("../data/raw_data/General Abstracts_2018.xlsx")
+
+general_2018_posters <- general_2018 %>% filter(grepl("Poster", `Presentation Format`))
+general_2018_papers  <- general_2018 %>% filter(grepl("Paper", `Presentation Format`))
 ```
@@
 # Create a corpus object from the abstracts
-general_corpus <- tm::Corpus(tm::DataframeSource(general), 
+general_2018_papers_corpus <- tm::Corpus(tm::DataframeSource(general_2018_papers), 
                              readerControl = list(reader = tm::readTabular(mapping = m)))
@@
 # create a document term matrix
-general_tm <- tm::DocumentTermMatrix(general_corpus, 
+general_2018_papers_tm <- tm::DocumentTermMatrix(general_2018_papers_corpus, 
                                      control = list(stemming = TRUE, 
@@
                                                     minWordLength = 2, 
-                      s                              removeNumbers = TRUE, 
+                                                    removeNumbers = TRUE, 
                                                     removePunctuation = TRUE))
+
 ```
@@
 # Use TF-IDF to wieght terms and remove rare ones
-term_tfidf <- 
-  tapply(general_tm$v/slam::row_sums(general_tm)[general_tm$i], 
-         general_tm$j, mean) * log2(tm::nDocs(general_tm)/slam::col_sums(general_tm > 0))
+term_2018_papers_tfidf <- 
+  tapply(general_2018_papers_tm$v/slam::row_sums(general_2018_papers_tm)[general_2018_papers_tm$i], 
+         general_2018_papers_tm$j, mean) * log2(tm::nDocs(general_2018_papers_tm)/slam::col_sums(general_2018_papers_tm > 0))
 
-summary(term_tfidf)
-# Median =  0.08989
+summary(term_2018_papers_tfidf)
+# Median =   0.08120
 ```
@@
 ## Keeping the rows with tfidf >= median
-general_reduced_dtm <- general_tm[,term_tfidf >= summary(term_tfidf)[3]]
-summary(slam::col_sums(general_reduced_dtm))
+general_2018_papers_reduced_dtm <- general_2018_papers_tm[,term_2018_papers_tfidf >= summary(term_2018_papers_tfidf)[3]]
+summary(slam::col_sums(general_2018_papers_reduced_dtm))
 # save
-saveRDS(general_reduced_dtm, "../data/derived_data/general_reduced_dtm.rds")
+saveRDS(general_2018_papers_reduced_dtm, "../data/derived_data/general_2018_papers_reduced_dtmrds")
 ```
@@
 ```{r}
-general_reduced_dtm <- readRDS("../data/derived_data/general_reduced_dtm.rds")
+general_2018_papers_reduced_dtm <- readRDS("../data/derived_data/general_2018_papers_reduced_dtm.rds")
+number_of_paper_abstracts <- general_2018_papers_reduced_dtm$nrow
 ```
@@
 ```{r eval=FALSE }
-system.time(fitted_many <- 
-              lapply(seqk, function(k) topicmodels::LDA(general_reduced_dtm, 
+system.time(fitted_many_2018_papers <- 
+              lapply(seqk, function(k) topicmodels::LDA(general_2018_papers_reduced_dtm, 
                                                         k = k,
@@
 
-saveRDS(fitted_many, "../data/derived_data/fitted_many.rds")
+saveRDS(fitted_many_2018_papers, "../data/derived_data/fitted_many_2018_papers.rds")
+
+# rerun with k = 60 to force the abstracts into 60 topics
 ```
@@
 ```{r}
-fitted_many <- readRDS("../data/derived_data/fitted_many.rds")
+fitted_many_2018_papers <- readRDS("../data/derived_data/fitted_many_2018_papers.rds")
+
+# how many abstracts, if we just start from here?
+how_many_paper_abstracts <- fitted_many_2018_papers[[1]]@Dim[1]
 ```
@@
 # extract logliks from each topic
-logLiks_many <- lapply(fitted_many, function(L)  L@logLiks[-c(1:(burnin/keep))])
+logLiks_many_2018_papers <- lapply(fitted_many_2018_papers, function(L)  L@logLiks[-c(1:(burnin/keep))])
 
 # compute harmonic means
-hm_many <- sapply(logLiks_many, function(h) harmonicMean(h))
+hm_many_2018_papers <- sapply(logLiks_many_2018_papers, function(h) harmonicMean(h))
 
 # to know the optimal number of topics:
-optimal_number_of_topics <- seqk[which.max(hm_many)]
+optimal_number_of_topics_2018_papers <- seqk[which.max(hm_many_2018_papers)]
 
@@
 lda_plot <- ggplot(data.frame(seqk, 
-                             hm_many), 
+                             hm_many_2018_papers), 
                   aes(x=seqk, 
-                      y=hm_many)) + 
+                      y=hm_many_2018_papers)) + 
   geom_path(lwd=1.5)  +
@@
   ylab('Harmonic Mean') +
-  geom_vline(xintercept = seqk[which.max(hm_many)],
+  geom_vline(xintercept = optimal_number_of_topics_2018_papers,
              colour = "red") +
@@
            label = paste("The optimal number of topics is", 
-                         seqk[which.max(hm_many)])) +
+                         optimal_number_of_topics_2018_papers)) +
   theme_bw() +
-  ggtitle(expression(atop("Latent Dirichlet Allocation Analysis of SAA General Abstracts", atop(italic("How many distinct topics in the abstracts?"), ""))))
+  ggtitle("Latent Dirichlet Allocation Analysis of SAA General Abstracts",
+          subtitle = paste0("How many distinct topics are present in the general paper abstracts (n = ", how_many_paper_abstracts, ") for 2018?"))
+
+ggsave("../figures/saa2018_general_abstracts_papers_opti_topics.png")
 ```
@@
 # now run the model with the optimum number of topics
-system.time(general_model <- topicmodels::LDA(general_reduced_dtm, 
-                                              optimal_number_of_topics,
+system.time(general_model_papers <- topicmodels::LDA(general_2018_papers_reduced_dtm, 
+                                              optimal_number_of_topics_2018_papers,
                                               method = "Gibbs", 
@@
 # explore the model
-general_topics <- topicmodels::topics(general_model, 1)
+general_topics_papers <- topicmodels::topics(general_model_papers, 1)
 
 ## In this case I am returning the top 30 terms.
-general_terms <- as.data.frame(topicmodels::terms(general_model, 30), 
+general_terms_papers <- as.data.frame(topicmodels::terms(general_model_papers, 30), 
                                stringsAsFactors = FALSE)
@@
 library(tidytext)
-tidy_topics <- tidy(general_model, matrix = "beta")
+tidy_topics_papers <- tidy(general_model_papers, matrix = "beta")
 
@@
 
-tidy_topics_top_terms <- tidy_topics %>%
+tidy_topics_papers_top_terms <- tidy_topics_papers %>%
   group_by(topic) %>%
@@
 ```{r}
-tidy_topics_top_terms %>%
+# show top keywords for each topic
+tidy_topics_papers_top_terms %>%
   mutate(term = reorder(term, beta)) %>%
@@
   coord_flip()
+
+ggsave("../figures/keywords_for_saa2018_general_papers_topics.png", h = 14, w = 14)
 ```
@@
 ```{r}
-tidy_documents <- tidy(general_model, matrix = "gamma")
+tidy_documents_papers <- tidy(general_model_papers, matrix = "gamma")
 
 library(tidyr)
-tidy_documents_topics_gamma <-
-  tidy_documents %>% 
+tidy_documents_papers_topics_gamma <-
+  tidy_documents_papers %>% 
     spread(topic, gamma)
 
-abstract_classifications <- 
-  tidy_documents %>%
+abstract_classifications_papers <- 
+  tidy_documents_papers %>%
   group_by(document) %>%
@@
   ungroup() %>% 
-  arrange(desc(gamma))
+  arrange(desc(gamma)) %>% 
+  mutate(document = as.numeric(document))
 
-abstract_classifications
+abstract_classifications_papers
+```
+
+```{r}
+# put topics back onto abstract spreadsheet
+# abstract_ids <- dimnames(general_2018_reduced_dtm)$Docs
+# general_2018$`Abstract Id`
+
+general_2018_papers_with_topics <- 
+  general_2018_papers %>% 
+  left_join(abstract_classifications_papers, 
+            by = c("Abstract Id" = "document"))
+
+write.csv(general_2018_papers_with_topics, "../data/derived_data/general_2018_papers_with_topics.csv")
 ```

### 2017-10-26 analysis/paper/saa2018_general_sessions_hclust.Rmd

@@
   tmp <- stringr::str_trim(tolower(gsub("[[:punct:]]|\r\n| and ", " ", x)))
+  tmp <- gsub("other", "", tmp)
   tmp <- gsub(" ", "_", tmp)
@@
 # ignore overly common words i.e. terms that appear in more than 50% of the documents
-maxTermFreq <- ndocs * .5
+maxTermFreq <- ndocs * 0.5
 
@@
 
+Distance matrix
+
 ```{r}
@@
 distMatrix <- dist(m, method="euclidean")
+```
 
+How many clusters?
+
+```{r}
+library(apcluster)
+d.apclus <- apcluster(negDistMat(r=2), m)
+k <- length(d.apclus@clusters)
+
+windows()
+heatmap(d.apclus)
+```
+
+Cluster analysis and dendrogram
+
+```{r}
 groups <- hclust(distMatrix,method="ward.D")
 plot(groups, cex=0.9, hang=-1)
-rect.hclust(groups, k=70)
+rect.hclust(groups, k)
 
 # see what cluster each abstract is in:
-cuts <- cutree(groups, k = 50)
+cuts <- cutree(groups, k)
+```
+
+How many items in each cluster
+
+```{r}
+# histogram
+freqs_per_cluster <- 
+cuts %>%
+  data_frame(cuts = .,
+             freq = names(.)) %>% 
+  group_by(cuts) %>% 
+  tally()
+
+hist(freqs_per_cluster$n)
+
+median(freqs_per_cluster$n) # 10
 ```

### 2017-10-26 analysis/paper/saa2018_general_sessions_papers_hclust.Rmd

--- 2017-10-26 analysis/paper/saa2018_general_sessions_hclust.Rmd
@@
 # read in the abstracts
-general_2018 <- read_excel("../data/raw_data/General Abstracts_2018.xlsx")
+general_2018 <- ab# read_excel("../data/raw_data/General Abstracts_2018.xlsx")
 
@@
 
+library(stringi)
 general_2018_papers$text_to_model <- 
-  with(general_2018_papers, paste(geo, kw1, kw2, kw3))
+  paste(stri_dup(paste(general_2018_papers$geo, ""), 4),
+        stri_dup(paste(general_2018_papers$kw1, ""), 3),
+        general_2018_papers$kw2, 
+        general_2018_papers$kw3)
 ```
@@
   group_by(cuts) %>% 
-  tally()
+  tally(sort = TRUE)
 
@@
 write.csv(general_2018_papers_with_topics,  "../data/derived_data/general_2018_papers_with_clusters.csv")
+```
+
+Compare
+
+```{r}
+ab <- read_excel("../data/derived_data/Copy of general_2018_papers_with_clusters_AB.xlsx")
+
+compare <- left_join(general_2018_papers_with_topics, ab, by = "Abstract Id")
+
+compare_groups <- 
+compare %>% 
+  select(cuts.x, cuts.y)
+```
+
+reclassify using weighted keywords
+
+```{r}
+
 ```

### 2017-10-26 analysis/paper/saa2018_nonposter_sessions_knapsack.Rmd

--- 2017-10-26 analysis/paper/paper.Rmd
@@
 # Introduction
-
-```{r eval=FALSE}
-# read in the abstracts
-general <- read_excel("../data/raw_data/General Abstracts.xlsx")
-```
-
-```{r eval=FALSE}
-# prepare a mapping structure, first item is column with text to model, others are metadata
-m <- list(content = "Abstract",  # content
-          id = "Abstract Id",      # metadata
-          title = "Title", 
-          author_first = "First Name",
-          author_last = "Last Name",
-          geo = "Geographic Focus",
-          keyword1 = "Keyword1", 
-          keyword2 = "Keyword2", 
-          keyword3 = "Keyword3", 
-          org = "Affiliation")
-```
-
-```{r eval=FALSE}
-# Create a corpus object from the abstracts
-general_corpus <- tm::Corpus(tm::DataframeSource(general), 
-                             readerControl = list(reader = tm::readTabular(mapping = m)))
-
-# create a document term matrix
-general_tm <- tm::DocumentTermMatrix(general_corpus, 
-                                     control = list(stemming = TRUE, 
-                                                    stopwords = TRUE,
-                                                    minWordLength = 2, 
-                                                    removeNumbers = TRUE, 
-                                                    removePunctuation = TRUE))
-```
-
-```{r eval=FALSE}
-# Use TF-IDF to wieght terms and remove rare ones
-term_tfidf <- 
-  tapply(general_tm$v/slam::row_sums(general_tm)[general_tm$i], 
-         general_tm$j, mean) * log2(tm::nDocs(general_tm)/slam::col_sums(general_tm > 0))
-
-summary(term_tfidf)
-# Median =  0.08989
-```
-
-```{r eval=FALSE }
-## Keeping the rows with tfidf >= median
-general_reduced_dtm <- general_tm[,term_tfidf >= summary(term_tfidf)[3]]
-summary(slam::col_sums(general_reduced_dtm))
-# save
-saveRDS(general_reduced_dtm, "../data/derived_data/general_reduced_dtm.rds")
-```
-
-```{r}
-general_reduced_dtm <- readRDS("../data/derived_data/general_reduced_dtm.rds")
-```
-
-```{r}
-# Determine k number of topics
-harmonicMean <- function(logLikelihoods, precision = 2000L) {
-  llMed <- median(logLikelihoods)
-  as.double(llMed - log(Rmpfr::mean(exp(-Rmpfr::mpfr(logLikelihoods,
-                                       prec = precision) + llMed))))
-}
-```
-
-```{r}
-# run a loop over the abstract with different numbers of topics to find which number of topcis is the best
-seqk <- seq(2, 100, 1)
-burnin <- 1000
-iter <- 1000
-keep <- 50
-```
-
-```{r eval=FALSE }
-system.time(fitted_many <- 
-              lapply(seqk, function(k) topicmodels::LDA(general_reduced_dtm, 
-                                                        k = k,
-                                                        method = "Gibbs",
-                                                        control = list(burnin = burnin,
-                                                                       iter = iter, 
-                                                                       keep = keep) )))
-
-saveRDS(fitted_many, "../data/derived_data/fitted_many.rds")
-```
-
-```{r}
-fitted_many <- readRDS("../data/derived_data/fitted_many.rds")
-```
-
-```{r}
-# extract logliks from each topic
-logLiks_many <- lapply(fitted_many, function(L)  L@logLiks[-c(1:(burnin/keep))])
-
-# compute harmonic means
-hm_many <- sapply(logLiks_many, function(h) harmonicMean(h))
-
-# to know the optimal number of topics:
-optimal_number_of_topics <- seqk[which.max(hm_many)]
-
-# plot it
-library(ggplot2)
-lda_plot <- ggplot(data.frame(seqk, 
-                             hm_many), 
-                  aes(x=seqk, 
-                      y=hm_many)) + 
-  geom_path(lwd=1.5)  +
-  xlab('Number of Topics') +
-  ylab('Harmonic Mean') +
-  geom_vline(xintercept = seqk[which.max(hm_many)],
-             colour = "red") +
-  annotate("text",
-           x = 55, 
-           y = -150000, 
-           label = paste("The optimal number of topics is", 
-                         seqk[which.max(hm_many)])) +
-  theme_bw() +
-  ggtitle(expression(atop("Latent Dirichlet Allocation Analysis of SAA General Abstracts", atop(italic("How many distinct topics in the abstracts?"), ""))))
-```
-
-```{r}
-# now run the model with the optimum number of topics
-system.time(general_model <- topicmodels::LDA(general_reduced_dtm, 
-                                              optimal_number_of_topics,
-                                              method = "Gibbs", 
-                                              control = list(iter=2000, 
-                                                             seed = 0622)))
-
-# explore the model
-general_topics <- topicmodels::topics(general_model, 1)
-
-## In this case I am returning the top 30 terms.
-general_terms <- as.data.frame(topicmodels::terms(general_model, 30), 
-                               stringsAsFactors = FALSE)
-```
-
-```{r}
-library(tidytext)
-tidy_topics <- tidy(general_model, matrix = "beta")
-
-library(ggplot2)
-library(dplyr)
-
-tidy_topics_top_terms <- tidy_topics %>%
-  group_by(topic) %>%
-  top_n(10, beta) %>%
-  ungroup() %>%
-  arrange(topic, -beta)
-```
-
-```{r}
-tidy_topics_top_terms %>%
-  mutate(term = reorder(term, beta)) %>%
-  ggplot(aes(term, beta, fill = factor(topic))) +
-  geom_col(show.legend = FALSE) +
-  facet_wrap(~ topic, scales = "free") +
-  coord_flip()
-```
-
-```{r}
-tidy_documents <- tidy(general_model, matrix = "gamma")
-
-library(tidyr)
-tidy_documents_topics_gamma <-
-  tidy_documents %>% 
-    spread(topic, gamma)
-
-abstract_classifications <- 
-  tidy_documents %>%
-  group_by(document) %>%
-  top_n(1, gamma) %>%
-  ungroup() %>% 
-  arrange(desc(gamma))
-```
-
-```{r eval=FALSE}
-mydata <- tidy_documents_topics_gamma[,-1]
-
-# Determine number of clusters
-library(NbClust)
-nclust <- NbClust(mydata, method = "kmeans")
-
-library(fpc)
-pamk.best <- pamk(mydata)
-
-# Determine number of clusters
-wss <- (nrow(mydata)-1)*sum(apply(mydata,2,var))
-for (i in 2:15) wss[i] <- sum(kmeans(mydata, 
-  	centers=i)$withinss)
-plot(1:15, wss, type="b", xlab="Number of Clusters",
-  ylab="Within groups sum of squares")
-
-# K-Means Cluster Analysis
-fit <- kmeans(mydata, 5) # 5 cluster solution
-# get cluster means 
-aggregate(mydata,by=list(fit$cluster),FUN=mean)
-# append cluster assignment
-mydata <- data.frame(mydata, fit$cluster)
-```
 

### 2017-10-26 analysis/paper/scheduling.Rmd

---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}
# http://www.sumsar.net/blog/2016/06/how-to-cut-your-planks-with-r/
planks_we_have <- c(120, 137, 220, 420, 480)
planks_we_want <- c(19, 19, 19, 19, 79, 79, 79, 103, 103,
                    103, 135, 135, 135, 135, 160)

library(adagio)
# mknapsack calling signature is: mknapsack(values, weights, capacities)
solution <- mknapsack(planks_we_want, 
                      planks_we_want, 
                      planks_we_have)
# Above I added +1 cm  to each length to compensate for the loss when sawing.
solution$ksack

assignment <- data.frame(
  cut_this = planks_we_have[solution$ksack],
  into_this = planks_we_want)
t(assignment[order(assignment[,1]), ])

#-----------------------

# total hours per day in all rooms
hours_per_day_we_have <- 
  c(100, 110)

# hours for each session 
hours_per_session_we_want <- 
  c(11, 21, 31, 41, 51)

library(adagio)
# mknapsack calling signature is: mknapsack(values, weights, capacities)
solution <- mknapsack(hours_per_session_we_want, 
                      hours_per_session_we_want, 
                      hours_per_day_we_have)
# Above I added +1 cm  to each length to compensate for the loss when sawing.
solution$ksack

assignment <- data.frame(
  into_this = hours_per_day_we_have[solution$ksack],
  put_this = hours_per_session_we_want)
t(assignment[order(assignment[,1]), ])

#----------------------

# one day at a time
# rooms by hours
big <- c(4001, 4002)
big
sum(big)

small <- c(1.00, 2.00, 1.50, 3.0) * 1000
small
sum(small)

library(adagio)
# mknapsack calling signature is: mknapsack(values, weights, capacities)
solution <- mknapsack(small, 
                      small, 
                      big)

{
assignment <- data.frame(
  into_this = big[solution$ksack],
  put_this = small)
t(assignment[order(assignment[,1]), ])
}

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2017-10-27 analysis/paper/saa2018_nonposter_sessions_knapsack.Rmd
while loop seems to be working, we can assign sessions while checking for day-part conflicts

@@
 ---
-title: "Title Goes Here"
+title: "Notes on scheduling the SAA2019 meeting"
 author:
-  - author 1
-  - author 2
+  - Ben Marwick
 date: "`r format(Sys.time(), '%d %B, %Y')`"
 output:
-    bookdown::html_document2:
-    fig_caption: yes
-    reference_docx: templates/template.docx
-bibliography: references.bib
-csl: journal-of-archaeological-science.csl
-abstract: |
-  Text of abstract
-keywords: |
-  keyword 1; keyword 2; keyword 3
-highlights: |
-  These are the highlights. 
+  tint::tintHtml: 
+     self_contained: TRUE
+     df_print: paged
 ---
@@
 
-```{r, setup, echo = FALSE}
+```{r, setup, echo = FALSE, warning = FALSE, message = FALSE,}
 knitr::opts_chunk$set(
@@
 library(saaabstracts)
-library(tm)
-library(topicmodels)
 library(readxl)
-library(Rmpfr)
-library(ggplot2)
+library(tidyverse)
+library(knitr)
+library(kableExtra)
+options(knitr.table.format = "html")
+
+my_kable <- function(dt){
+kable(dt) %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
+                full_width = FALSE,
+                position = "left") %>% 
+    column_spec(seq(1, ncol(dt), 2),  background = "grey80")
+}
 ```
+## Introduction
 
-# Introduction
+In these notes I attempt to make a start on organising the conference sessions into time slots suitable for scheduling.
 
-## Knapsack
+We have the individual submissions already organised into sessions. The task here is to identify the slots available and fit the sessions into them. Here's what I've done:
 
-We have how many sessions?
+First, we computed a suitable duration for each session, given the number of participants and the SAA policy about the maximum duration of the various session types.
+
+Second, we assigned all non-poster sessions to the available four hour time slots. We tallied up the number of slots by looking at the 'Room & Time' spreadsheet.
+
+Third, we spread the slots across the parts of each day of the meeting (morning, afternoon, evening), since each day has slightly different room resources available.
+
+Fourth, we identified problems and schedule clashes, and what to do next.
+
+## How many sessions do we have?
 
@@
 
+poster_sessions <- 
+  org_sessions %>% 
+  filter(`Session Type` == "Poster Symposium")
+
+# organised sessions without posters
+org_sessions <- 
+   org_sessions %>% 
+  filter(`Session Type` != "Poster Symposium")
+
 # how many unique organised sessions? 
@@
 
+# how many unique poster sessions? 
+n_pos_sessions <- length(unique(poster_sessions$`Session Id`))
+
 # total number of sessions
-total_number_of_sessions <-  n_org_sessions + n_gen_sessions
+total_number_of_sessions <-  n_org_sessions + n_gen_sessions + n_pos_sessions
 ```
 
-For sheduling into rooms, we don't include the posters, because they go in the ballroom.
-How many 4h slots do we have?
-I looked on the grid on 'Rooom&Time_DG.xlsx' and counted up the 4h slots, excluding poster sessions, because they are less than 4 h
+We have this many sessions in each major category:
+
+```{r}
+# table
+tribble(       ~'Session type',             ~number,
+                      'Organised sessions',        n_org_sessions,
+                      'Poster sessions',           n_pos_sessions,
+                      'General sessions',          n_gen_sessions)
+```
+
+With a total of `r total_number_of_sessions` sessions altogether, and a total number of non-poster sessions `r total_number_of_sessions - n_pos_sessions`.
+
+For scheduling into rooms and time slots, we wont include the posters, because they go in their own schedule in the ballroom.
+
+## How many 4 hour slots do we have available?
+
+To find out how many 4 h slots we have available, I looked on the grid on 'Rooom&Time_DG.xlsx' and counted up the 4  slots.
+I did not count poster sessions here, because they are less than 4 h and they are outside of the room schedule.
+I rounded down a little:
+
+- Thur afternoon has the ethics bowl from 1-3 pm, so I removed one whole 4 h slot from that part   
+- Fri morning has some time reserved from 9am-12 pm with the note "Reserved for Business Meeting after 12:00pm", so removed one whole 4 h slot from that part also
+
+Here's what I got for session slots in each day and part of day:
 
@@
 four_h_slots <- 
-  tribble(~time,       ~Th, ~Fr, ~Sa, ~Su,
+  tribble(~part,       ~Th, ~Fr, ~Sa, ~Su,
           "morning",    36,  31,  30,  28,
@@
 
+(four_h_slots)
+```
+
+```{r}
 four_h_slots_long <- 
   four_h_slots %>% 
-  gather(times, slot, -time)
+  gather(times, slot, -part)
 
@@
 
-So that's `r n_four_h_slots` 4 hour slots, where we consider each slot a knapsack.
+So that's a total of `r n_four_h_slots` 4 hour slots, in which we have to fit all the organised and general sessions. This does not include the posters.
 
-Now we need to pack the sessions into these slots.
-First we need to set the upper bounds for the session durations, according to the SAA rules, and then compute how much time each session is likely to take give the number of papers/speakers
+## What are the durations of the sessions?
+
+We need to set the upper bounds for the session durations, according to the SAA rules, and then compute how much time each session is likely to take give the number of papers/speakers, if that is less than the allowed time.
+
+The plot below shows the distribution of papers across all non-poster session:
 
@@
 
+# get only columns that are also in the organised sessions data frame
 gen_sessions_to_join <- 
@@
 
-# how many papers per session
+# how many papers per session?
 papers_per_session <- 
@@
   geom_histogram() +
-  ylab("Papers per non-poster session") +
+  ylab("Papers per \nnon-poster session") +
   theme_bw()
+```
+
+Note that I treat chairs and discussants the same as presenters, even though only presenters are really supposed to be allocated program time.
+The reason for keeping them in is to keep the names so we can avoid schedule clashes where one person is scheduled to be in two rooms at once.
+
+Here are the SAA rules for the duration of each session type:
+
+```{r}
 
@@
 session_time_limits <- 
-  tribble(~`Session Type`,             ~max_time,
+  tribble(~`Session Type`,             ~saa_allowed_time,
           "Poster Symposium",           2,
@@
 
+(session_time_limits)
+```
+
+Here we take the SAA rules for the duration of each session type, and if a session has more than 16 papers, we simply assign the duration of that session the maximum allowed time according to the SAA rules.
+We also compute the 'anticipated duration' of each session by taking the number of papers and mulitplying by 15 minutes.
+This is most relevant for the 'symposium' types, because other session types don't always have paper presentations.
+But we apply the upper duration to all sessions for now to ensure they all follow the SAA rules about duration.
+
+```{r}
 # join time limits onto session data
-session_with_limits <- 
+session_with_limits_int <- 
   all_nonposter_sessions %>% 
@@
 # now let's see which session exceed the time limits
+min_duration <-  1.5 
+round_up_by <- 0.25
 session_with_limits <- 
+session_with_limits_int %>% 
+  group_by(`Session Id`, 
+           `Session Type`, 
+           saa_allowed_time ) %>% 
+  tally(sort = TRUE) %>% 
+  mutate(session_time_ideal = n * 15/60) %>% 
+  # round up to nearest 0.5 hour
+  mutate(session_time_allowed = ceiling(session_time_ideal / round_up_by) * round_up_by) %>% 
+  # set the SAA-allowed time as the max time
+  mutate(session_time_allowed = if_else(session_time_allowed >= saa_allowed_time,
+                                  saa_allowed_time, 
+                                  session_time_allowed)) %>% 
+  # set a min duration of 2 h
+    mutate(session_time_allowed = if_else(session_time_allowed <= min_duration,
+                                  min_duration, 
+                                  session_time_allowed)) 
+```
+
+The plot below shows how a bunch of sessions have had their durations change according the following rules:
+
+- round up each session duration to the nearest `r round_up_by` hour, for convienence, since it's awkward to have sessions starting and ending at odd times    
+- shorten sessions to fit into the SAA rules of 2 and 4 h, depending on the session type    
+- for very small sessions, set a minumum duration of `r min_duration`, because that's what most session organisers wil be expecting
+
+```{r fig.fullwidth = TRUE, fig.width = 10}
+
+# plot the effect of setting the SAA limits
 session_with_limits %>% 
-  group_by(`Session Id`, `Session Type`, max_time ) %>% 
-  tally(sort = TRUE) %>% 
-  mutate(paper_time = n * 15/60) %>% 
-  mutate(paper_time = if_else(paper_time > max_time,  max_time, paper_time))
+  ungroup() %>% 
+  select(`Session Id`, 
+         `Session Type`,
+         session_time_ideal, 
+         session_time_allowed) %>% 
+  gather(time_type, time,  
+         -`Session Id`, 
+         -`Session Type`) %>% 
+ggplot(aes(time)) +
+  geom_histogram() +
+  geom_vline(xintercept = c(2,4), 
+             colour = "red") +
+  facet_wrap(~   time_type + `Session Type`, ncol = 4) +
+  theme_bw() +
+  xlab("Duration in hours") +
+  ylab("Number of non-poster sessions") + 
+  ggtitle("Distribution of durations of non-poster sessions",
+          subtitle = paste0("The 'ideal' session times are computed by 15 min x N participants in the session.
+The 'allowed' times are mandated by the SAA, \nwith ideal times rounded up to the nearest ", round_up_by, " h, and a minimum duration of ", min_duration, "  h "))
 
@@
 
-Now we have a known number of 4 hour sessions (knapsacks), and a known number of sessions of various lengths.
-We turn to http://www.sumsar.net/blog/2016/06/how-to-cut-your-planks-with-r/ to see how to arrange them most efficiently.
-Here is the original knapsack example that we'll try to adapt:
+QUESTION!
+If there are a very small number of paricipants in a session, do we allocate the SAA-imposed limit to that session?
+Or do we allow smaller sessions to have shorter durations, for example schedule 1 hour for 4 people?
+That seems unfair, so I've set a minimum duration of `r min_duration` h for all sessions, since that's what most session chairs will be expecting.
+I've also rounded up the durations to the nearest `r round_up_by` min
 
-```{r}
-# Here is Rasmus' example: http://www.sumsar.net/blog/2016/06/how-to-cut-your-planks-with-r/
+## How to distribute the sessions in the 4 hour slots?
 
-planks_we_have <- c(120, 137, 220, 420, 480)
-planks_we_want <- c(19, 19, 19, 19, 79, 79, 79, 103, 103,
-                    103, 135, 135, 135, 135, 160)
+Now we have computed durations for all the sessions, we can try to find the optimum way to squeeze them into the slots available.
 
-library(adagio)
-# mknapsack calling signature is: mknapsack(values, weights, capacities)
-solution <- mknapsack(planks_we_want, planks_we_want + 1, planks_we_have)
-# Above I added +1 cm  to each length to compensate for the loss when sawing.
-solution$ksack
+This question is combinatorial optimization question, more specifically, it is type of multiple knapsack problem.
+Given a set of resources (i.e.
+sessions of various lengths), how do we allocate them to a give number of time slots (all the same size, four hours)?
+Once we solve that problem, then we will distribute these slots across the day-parts.
 
-# That is, cut plank length 1 from plank 1, plank length 2 from plank 4, etc.
-
-# Now pretty printing what to cut so that we don't make mistakes...
-assignment <- data.frame(
-  cut_this = planks_we_have[solution$ksack],
-  into_this = planks_we_want)
-t(assignment[order(assignment[,1]), ])
-```
-
-Now let's substitute with our data:
+To get started on this, we can use Rasmus Bååth's nice case study of the the multiple knapsack problem (<http://www.sumsar.net/blog/2016/06/how-to-cut-your-planks-with-r/>).
 
@@
 # we want to fit these
-sessions_we_want <- session_with_limits$paper_time * 60 # the duration of each session
+sessions_we_want <- session_with_limits$session_time_allowed * 60 # the duration of each session
+
+# add a buffer onto the sessions less than the full slot
+buffer <- 15
+sessions_we_want <- ifelse(sessions_we_want <= max(sessions_we_want) - buffer, 
+                           sessions_we_want + buffer,
+                           sessions_we_want)
 
@@
 solution <- mknapsack(sessions_we_want, sessions_we_want, sessions_we_have)
+```
 
-solution$ksack
-# That is, put session 1 into slot 1, session 2 into slot 2, etc.
+Let's put the result back onto the data frame of all the non-poster sessions so we can see which sessions can go in the same slot.
 
-# Now pretty printing what to session to put in what slot so that we don't make mistakes...
-assignment <- data.frame(
-  fill_this_slot = sessions_we_have[solution$ksack],
-  with_this_session = sessions_we_want)
-t(assignment[order(assignment[,1]), ])
+In the snippet below we see that our multiple knapsack function has allocated two sessions into one computed_slot.
+This looks good because Electronic Symposia and Lightning Rounds can only be 2 hours duration, so we should be able to fit two into one slot.
+So the multiple knapsack approach seems to be working.
 
+```{r}
 # put the computed slot back onto the full details of the sessions
 session_with_limits$computed_slot <- solution$ksack
+
+session_with_limits %>% 
+  arrange(session_time_allowed, computed_slot) %>% 
+  head 
 ```
 
-That seems to work great, assuming that we've got the time limits ok. Now we arbitrarily assign them to the schedule, then check for conflicts
+Now that we have assigned sessions to slots, we need to join this slot information onto the data frame of each individual paper.
+This is necessary so we can ensure that we don't put one person in two places at once when we distribute the slots across the day-parts on the schedule.
+
+```{r}
+all_nonposter_sessions_with_slots <- 
+all_nonposter_sessions %>% 
+  left_join(session_with_limits)
+```
+
+# How to avoid putting one person in two places at once?
+
+We make a unique ID out of the first name and last name to track individuals.
+
+```{r}
+all_nonposter_sessions_with_slots$uid <- 
+  with(all_nonposter_sessions_with_slots, paste0(`First Name`, 
+                                                 "_", 
+                                                 `Last Name`))
+```
+
+We assign slots to day-parts, then we join this day-part data back to the data frame of individual papers.
+
+The table below shows a snippet of how the slots available are distributed across the day-parts:
 
@@
 four_h_slots_long %>% 
-  mutate(day_part = paste0(times, "-", time)) %>% 
-  select(-time, -times) %>% 
+  mutate(day_part = paste0(times, "-", part)) %>% 
+  select(-part, -times) %>% 
   rename(end = cumsum) %>% 
-  filter( slot != 0)
+  filter(slot != 0)
 
@@
 for(i in 1:(nrow(xx))) {
-
   xx$start[i] <- 
@@
             as.numeric(xx[ i-1, 2]) + 1)
-
 }
@@
   select(-slot) %>% 
-  select(day_part, start, end )
+  select(day_part, start, end ) %>% 
+  rename(from_slot = start,
+         to_slot = end)
 
-elements <- session_with_limits$computed_slot
+(head(intervals))
+```
+
+Now we take the computed slot values for each session that we got from the multiple knapsack function and put the sessions into day-parts.
+This table shows a snippet of the computed slots assigned to day-parts:
+
+```{r}
+elements <- unique(session_with_limits$computed_slot)
 
@@
                 intervals, 
-                by = c("elements" = "start", 
-                       "elements" = "end"), 
+                by = c("elements" = "from_slot", 
+                       "elements" = "to_slot"), 
                 match_fun = list(`>=`, `<=`)) %>% 
@@
 
-# Which slot goes into which part of what day
-slot_day_part <- 
+computed_slots_in_day_parts <- 
 join_intervals_and_elements %>% 
-  arrange(elements, day_part)
+  arrange(elements, day_part) %>% 
+  rename(computed_slot = elements)
 
-# now we put the sessions on also
-session_slot_day_part <- 
+(head(computed_slots_in_day_parts))
+```
+
+Now we join the day-parts onto the individual sessions and then the individual abstracts by the computed slot values.
+
+And now we can plot one possible configuration of the sessions as a grid of sessions for each block of time on the schedule:
+
+```{r fig.fullwidth = TRUE, fig.width = 10, fig.height = 8}
+# sessions
+all_nonposter_sessions_with_day_parts <- 
 session_with_limits %>% 
-  left_join(slot_day_part, by  = c("computed_slot" = "elements"))
+  left_join(computed_slots_in_day_parts)
+
+# find out what sessions are fractions of 4 hour slots
+zzz <- 
+all_nonposter_sessions_with_day_parts %>% 
+  group_by(computed_slot ) %>% 
+  add_count() %>% 
+  ungroup() %>% 
+  mutate(fraction_of_slot = 1/nn) %>% 
+  arrange(computed_slot)
+
+zzzz <- 
+zzz %>% 
+  group_by(day_part) %>% 
+  mutate(day_slot =  computed_slot - from_slot + 1) %>% 
+  ungroup()
+
+# get session label string
+session_label_string <- 
+zzzz %>% 
+  group_by(computed_slot) %>% 
+  summarise(strings = toString(unique(`Session Id`)))
+
+z6 <- 
+  zzzz %>% 
+  left_join(session_label_string)
+
+ggplot(z6, 
+       aes(day_part, 
+           day_slot)) +
+  geom_tile(colour = "black", 
+           aes(fill = factor(session_time_allowed))) +
+  geom_text(aes(label = strings), 
+                  size = 3) +
+  scale_fill_viridis_d() +
+  scale_x_discrete(limits = xx$day_part) +
+  scale_y_continuous(breaks = 0:max(z6$day_slot)) +
+  guides(colour = FALSE) +
+  theme_bw() +
+  xlab("") +
+  ylab("Room-ish") +
+  ggtitle("Visual schedule for SAA2019 conference non-poster sessions",
+          subtitle = "The numbers on the plot are the Session Ids.
+\n We can think of each row as equivalent to a room at the conference venue.
+\n The colour of the cell indicates the duration of the individual session(s) that each cell represents.\n For example the purple cells with two session IDs means that each session is 2 h, for a total of 4 hours in that slot.")
+
+# can we output that as a spreadsheet?
+library(tidyr)
+z6_plot_as_spreadsheet <- 
+  z6 %>% 
+  select(day_part, day_slot, strings) %>% 
+  distinct(strings, .keep_all = TRUE) %>% 
+  spread(day_part, strings, fill = "") %>% 
+  select(one_of(xx$day_part))
+
+write_csv(z6_plot_as_spreadsheet,
+          "../data/derived_data/visual_schedule_plot_as_spreadsheet.csv")
+
+# papers
+all_nonposter_papers_with_day_parts <- 
+all_nonposter_sessions_with_slots %>% 
+  left_join(computed_slots_in_day_parts)
 
 ```
+
+Now let's identify problems where one person is in two different sessions in the same day-part (i.e. column in the plot above).
+
+In the data frame of individual papers, we check that the UID does not occur in more than one session ID for one day-part (it can occur twice on one session ID).
+
+This plot shows sessions where one person is currently scheduled to be in two places at one (the red cells indicate the clashing sessions):
+
+```{r fig.fullwidth = TRUE, fig.width = 10, fig.height = 8}
+# identify people with roles in more than one session, but only in the same day-part
+# we don't care about mulitple roles in one single session
+three_or_more <- 
+all_nonposter_papers_with_day_parts %>% 
+  group_by(day_part, uid) %>% 
+  add_count(sort = TRUE)  %>% 
+  filter(nn >= 3) %>% 
+  ungroup()  %>% 
+  distinct(`Session Id`, uid, .keep_all = TRUE) %>% 
+  group_by(uid) %>% 
+  filter(n()>1)
+
+# join clashing sessions with plotting data
+session_with_clash_flags <- 
+  z6 %>% 
+  left_join(three_or_more, c("Session Id", "Session Type", "saa_allowed_time", "session_time_ideal", "session_time_allowed", "computed_slot", "day_part", "from_slot", "to_slot")) %>% 
+  mutate(fill = if_else(is.na(nn.y), "no_clash", "clash"))
+
+# plot again 
+ggplot(session_with_clash_flags, 
+       aes(day_part, 
+           day_slot )) +
+  geom_tile(aes(fill = factor(fill)),
+            colour = "black") +
+  geom_text(aes(label = strings), 
+                  size = 3) +
+  scale_x_discrete(limits = xx$day_part) +
+  scale_y_continuous(breaks = 0:max(z6$day_slot)) +
+  scale_fill_manual(values = c("red", "green")) +
+  guides(fill=FALSE, 
+         colour = FALSE) +
+  theme_bw() +
+  xlab("") +
+  ylab("Room-ish") +
+  ggtitle("Visual schedule for SAA2019 conference non-poster sessions",
+          subtitle = "The numbers on the plot are the Session Ids.
+\n We can think of  each row as equivalent to a room at the conference venue.
+\n Red cells indicate that one person is currently scheduled to be in both sessions :(")
+
+```
+
+## How can we prevent one person scheduled in two simultaneous sessions?
+
+If we treat each day-part as one knapsack, and then loop over filling the knapsack to check if a name has already been added, and if it has, skip that day-part and move to the next, and check that next pay-part to see if the name is already there.
+
+We can compute the total capacity of each day-part in hours like this:
+
+```{r}
+# each slot is 4 hours long
+total_cap <- 
+  xx %>% 
+  mutate(total_capacity = slot * 4) %>% 
+  select(day_part, slot, total_capacity)
+total_cap
+```
+
+Now we will take session from the big table of sessions and add them into, for example, Th-morning until we reach the capacity of that day-part.
+But, we also need to loop over the sessions as we add more to the day-part to check to see if a person's name is already there or not.
+
+```{r echo = TRUE, eval = FALSE}
+
+# get all the sessions, but we're going to modify this, so make a copy
+z7 <- z6
+all_of_the_day_parts <- NULL
+total_cap$hours_of_sessions <- 0
+total_cap$n_sessions <- 0
+total_cap$c_sessions <- 0
+container_for_this_day_part_with_day_from_while <- 0
+i <- 1
+
+# start with each day-part...
+for(i in seq_len(nrow(total_cap))) {
+  target_capacity <- total_cap$total_capacity[i] - 4
+
+  # prepare a while statement that will take a session, 
+  # check to see if there is a name clash with sessions,
+  # skip adding it if there is clash, then check to see
+  # the culumative duration of the sessions in the day-part 
+  # has exceeeded the total duration of the day-part,
+  # and stop adding sessions if it has.
+
+# make sure some variables are empty
+cumulative_duration_of_sessions   <-  0
+container_for_this_day_part <- 0
+all_the_people_in_the_session <- 0
+j <- 1
+
+while(cumulative_duration_of_sessions <= target_capacity & 
+      j <= nrow(z7)) {
+
+    get_a_session <- z7[j, ]
+
+    # get the names of all the people in this session...
+    all_the_people_in_this_session <- 
+      left_join(get_a_session, 
+                all_nonposter_papers_with_day_parts, 
+                by = "Session Id")
+
+    # get names from what we already have added to this day-part
+     if (j == 1) {
+      # do nothing is this is the first incoming session 
+       distinct_names_for_each_session_so_far <- 0
+    } else {
+      # get distinct names for each session previously added
+      # to this day-part
+      # (don't care about same name multiple times in one session)
+      distinct_names_for_each_session_so_far <- 
+      container_for_this_day_part %>% 
+        group_by(uid) %>% 
+        distinct(uid, .keep_all = TRUE) %>% 
+        pull(uid)
+    }
+
+    # compare names already in here to names we are about to add
+    if (j == 1) {
+      # do nothing 
+      names_already_in <- 0
+      } else {
+      # check to see if incoming session has names that already exist in this day_part
+      names_already_in <- sum(all_the_people_in_this_session$uid %in% distinct_names_for_each_session_so_far)
+      }
+
+    # accumulate the incoming session onto the sessions
+    # we've already added to this day-part, if there are no
+    # name conflicts
+    if (j == 1) {
+       container_for_this_day_part <- 
+        all_the_people_in_this_session
+      } else {
+      if (names_already_in == 0) {
+      # if there are no name clashes, 
+      # add these names to previous names in this day-part
+
+      # bind rows of session that has no clashing  names with 
+      # sessions we already added to this day-part
+      container_for_this_day_part <- 
+        bind_rows(all_the_people_in_this_session,
+                  container_for_this_day_part)
+    }  else {
+
+        # do not add this session to the day-part
+      }
+    }
+
+    # compute the duration of session that we've added so far to 
+    # this day-part
+    cumulative_duration_of_sessions <- 
+      container_for_this_day_part %>% 
+      distinct(`Session Id`, .keep_all = TRUE) %>% 
+      pull(session_time_allowed.x) %>% 
+      sum
+
+    # store this result
+    total_cap$hours_of_sessions[i] <- cumulative_duration_of_sessions
+
+    # show a check of what's happening
+    if (j == 1){
+    progress_check <-    
+      data_frame(i = i, 
+                 j = j,
+               'cumulative duration' = cumulative_duration_of_sessions,
+               'sessions added' =  length(unique(container_for_this_day_part$`Session Id`)),
+               people =  nrow(container_for_this_day_part)
+                )
+    } else {
+      progress_check <- bind_rows(progress_check, 
+                                  data_frame(i = i,
+                                             j = j,
+               'cumulative duration' = cumulative_duration_of_sessions,
+               'sessions added' =  length(unique(container_for_this_day_part$`Session Id`)),
+               people =  nrow(container_for_this_day_part)
+                ))
+    }
+
+       # print(progress_check)
+
+       # move to next session
+       j <- j + 1
+    } 
+# end of the while loop
+
+ # add an identifier so we can see where we have assigned each session so far
+ container_for_this_day_part_with_day_from_while <- 
+   container_for_this_day_part %>% 
+   mutate(day_part_from_while  = as.character(total_cap$day_part[i]))
+
+ # bind rows to previous day-parts we might have done
+  if (i == 1) {
+
+  all_of_the_day_parts <- container_for_this_day_part_with_day_from_while
+
+  } else {
+
+    # check to see
+
+  all_of_the_day_parts <- bind_rows(container_for_this_day_part_with_day_from_while, 
+                                      all_of_the_day_parts)
+  }
+
+ # progress check
+ print(paste0("i = ", i))
+
+ print(
+ all_of_the_day_parts %>% 
+   group_by(day_part_from_while) %>% 
+   summarise(papers = n(),
+             sessions = length(unique(`Session Id`)))
+ )
+
+ total_cap$n_sessions[i] <- length(unique(container_for_this_day_part_with_day_from_while$`Session Id`))
+
+ total_cap$c_sessions[i] <-  sum(total_cap$n_sessions)
+
+print(total_cap)
+
+ # for the next day-part, we need to make sure
+ # we don't add the same session twice to different day-parts
+ # so let's update our big table of abstracts, and drop the session that 
+ # we've already got in the day-part
+
+ z7 <-  
+   z7 %>% 
+   filter(! `Session Id` %in% unique(container_for_this_day_part$`Session Id`))
+
+ if (nrow(z7) == 0) stop("Stopping the loop because all sessions have been assigned to day-parts")
+
+}
+```
+
+```{r}
+# check to see what we got from the loop...
+
+n_missing_papers <- 
+  nrow(all_nonposter_sessions_with_slots) - nrow(all_of_the_day_parts)
+
+n_missing_sessions <- 
+  length(unique(all_nonposter_sessions_with_slots$`Session Id`)) - length(unique(all_of_the_day_parts$`Session Id`))
+
+missing_sessions <- 
+  unique(all_nonposter_sessions_with_slots$`Session Id`)[!(unique(all_nonposter_sessions_with_slots$`Session Id`) %in% unique(all_of_the_day_parts$`Session Id`)) ]
+
+# have we succeeded with no name clashes?
+# in each day-part, do any names occur more than once?
+clash <- 
+all_of_the_day_parts %>% 
+  group_by(`Session Id`, day_part_from_while, uid) %>% 
+  tally() %>% 
+  arrange(`Session Id`, n)
+
+```
+
+## Summary
+
+Now we have a rough draft of the schedule of non-poster sessions.
+We used the multiple knapsack algorithm to distribute the sessions efficiently across the available sessions.
+This method identifies combinations of sessions that are <4 h and then combines these shorter sessions to fill the 4 h slots.
+
+We still have some things to do:
+
+- Resolve the clashes that currently have one person in two sessions at the same time. I'm thinking of how it can be done programatically, but nothing obvious has occured to me.     
+- Assign the sessions to specific times and rooms. Perhaps this is best be done manually?    
+- Consider all of the conflicts and special requests in 'Conflicts.docx'    
+- Repeat all of this for the posters. This could be done with some minor variations to the code used here.
+
+<!--
 
@@
 Or do `docker ps` and `docker stop <container ID>` just to stop one container. 
+-->
 

### 2017-11-01 analysis/paper/saa2018_nonposter_sessions_while_loop.Rmd
use while-loop to assign people to day-parts without clashes

--- 2017-10-26 analysis/paper/saa2018_nonposter_sessions_knapsack.Rmd
@@
   message = FALSE,
-  echo = FALSE,
+  echo = TRUE,
   comment = "#>",
@@
 library(saaabstracts)
-library(tm)
-library(topicmodels)
 library(readxl)
-library(Rmpfr)
-library(ggplot2)
+library(tidyverse)
+library(knitr)
+library(kableExtra)
+library(glue)
 ```
+## Introduction
 
-# Introduction
-
-## Knapsack
-
-We have how many sessions?
+## How many sessions do we have?
 
@@
 
+# how many unique poster sessions? 
+n_pos_sessions <- length(unique(poster_sessions$`Session Id`))
+
 # total number of sessions
-total_number_of_sessions <-  n_org_sessions + n_gen_sessions
-```
+total_number_of_sessions <-  n_org_sessions + n_gen_sessions + n_pos_sessions
 
-For sheduling into rooms, we don't include the posters, because they go in the ballroom.
-How many 4h slots do we have?
-I looked on the grid on 'Rooom&Time_DG.xlsx' and counted up the 4h slots, excluding poster sessions, because they are less than 4 h
-
-```{r}
-library(tidyverse)
-four_h_slots <- 
-  tribble(~time,       ~Th, ~Fr, ~Sa, ~Su,
-          "morning",    36,  31,  30,  28,
-          "afternoon",  35,  32,  30,  0,
-          "evening",    36,  0,    0,  0)
-
-four_h_slots_long <- 
-  four_h_slots %>% 
-  gather(times, slot, -time)
-
-# how many 4 h slots
-n_four_h_slots <-  sum(four_h_slots_long$slot)
-```
-
-So that's `r n_four_h_slots` 4 hour slots, where we consider each slot a knapsack.
-
-Now we need to pack the sessions into these slots.
-First we need to set the upper bounds for the session durations, according to the SAA rules, and then compute how much time each session is likely to take give the number of papers/speakers
-
-```{r}
 # combine organised and general sessions
@@
   bind_rows(gen_sessions_to_join, org_sessions)
-
-# how many papers per session
-papers_per_session <- 
-all_nonposter_sessions %>% 
-  group_by(`Session Id`) %>% 
-  tally(sort = TRUE)
-
-ggplot(papers_per_session,
-       aes(n)) +
-  geom_histogram() +
-  ylab("Papers per non-poster session") +
-  theme_bw()
-
-# what's the duration of each session, and he max allowed duration? 
-session_types <- unique(all_nonposter_sessions$`Session Type`)
-
-# this comes from 'Conflicts.docx'
-session_time_limits <- 
-  tribble(~`Session Type`,             ~max_time,
-          "Poster Symposium",           2,
-          "Symposium",                  4,
-          "Electronic Symposium",       2, 
-          "Forum" ,                     2, 
-          "Lightning Rounds",           2)
-
-# join time limits onto session data
-session_with_limits <- 
-  all_nonposter_sessions %>% 
-  filter(`Session Type` != "Poster Symposium") %>% 
-  left_join(session_time_limits)
-
-# now let's see which session exceed the time limits
-session_with_limits <- 
-session_with_limits %>% 
-  group_by(`Session Id`, `Session Type`, max_time ) %>% 
-  tally(sort = TRUE) %>% 
-  mutate(paper_time = n * 15/60) %>% 
-  mutate(paper_time = if_else(paper_time > max_time,  max_time, paper_time))
-
 ```
 
-Now we have a known number of 4 hour sessions (knapsacks), and a known number of sessions of various lengths.
-We turn to http://www.sumsar.net/blog/2016/06/how-to-cut-your-planks-with-r/ to see how to arrange them most efficiently.
-Here is the original knapsack example that we'll try to adapt:
+We have this many sessions in each major category:
 
 ```{r}
-# Here is Rasmus' example: http://www.sumsar.net/blog/2016/06/how-to-cut-your-planks-with-r/
-
-planks_we_have <- c(120, 137, 220, 420, 480)
-planks_we_want <- c(19, 19, 19, 19, 79, 79, 79, 103, 103,
-                    103, 135, 135, 135, 135, 160)
-
-library(adagio)
-# mknapsack calling signature is: mknapsack(values, weights, capacities)
-solution <- mknapsack(planks_we_want, planks_we_want + 1, planks_we_have)
-# Above I added +1 cm  to each length to compensate for the loss when sawing.
-solution$ksack
-
-# That is, cut plank length 1 from plank 1, plank length 2 from plank 4, etc.
-
-# Now pretty printing what to cut so that we don't make mistakes...
-assignment <- data.frame(
-  cut_this = planks_we_have[solution$ksack],
-  into_this = planks_we_want)
-t(assignment[order(assignment[,1]), ])
+# table
+tribble(       ~'Session type',             ~number,
+                      'Organised sessions',        n_org_sessions,
+                      'Poster sessions',           n_pos_sessions,
+                      'General sessions',          n_gen_sessions)
 ```
 
-Now let's substitute with our data:
+With a total of `r total_number_of_sessions` sessions altogether, and a total number of non-poster sessions `r total_number_of_sessions - n_pos_sessions`.
+
+For scheduling into rooms and time slots, we wont include the posters, because they go in their own schedule in the ballroom.
+
+To help with preventing schedule clashes, we make a unique ID out of the first name and last name to track specific individuals.
 
 ```{r}
-# Convert the durations to minutes so we get integers, mknapsack doesn't work with decimals
-
-# into these
-sessions_we_have <- rep(4, n_four_h_slots) * 60 # so many 4 hour slots that we need to fill 
-# we want to fit these
-sessions_we_want <- session_with_limits$paper_time * 60 # the duration of each session
-
-library(adagio)
-# mknapsack calling signature is: mknapsack(values, weights, capacities)
-solution <- mknapsack(sessions_we_want, sessions_we_want, sessions_we_have)
-
-solution$ksack
-# That is, put session 1 into slot 1, session 2 into slot 2, etc.
-
-# Now pretty printing what to session to put in what slot so that we don't make mistakes...
-assignment <- data.frame(
-  fill_this_slot = sessions_we_have[solution$ksack],
-  with_this_session = sessions_we_want)
-t(assignment[order(assignment[,1]), ])
-
-# put the computed slot back onto the full details of the sessions
-session_with_limits$computed_slot <- solution$ksack
+all_nonposter_sessions$uid <- 
+  with(all_nonposter_sessions, paste0(`First Name`, 
+                                                 "_", 
+                                                 `Last Name`))
 ```
 
-That seems to work great, assuming that we've got the time limits ok. Now we arbitrarily assign them to the schedule, then check for conflicts
+## How many 4 hour slots do we have available?
+
+To find out how many 4 h slots we have available, I looked on the grid on 'Rooom&Time_DG.xlsx' and counted up the 4  slots.
+I did not count poster sessions here, because they are less than 4 h and they are outside of the room schedule.
+I rounded down a little:
+
+- Thur afternoon has the ethics bowl from 1-3 pm, so I removed one whole 4 h slot from that part   
+- Fri morning has some time reserved from 9am-12 pm with the note "Reserved for Business Meeting after 12:00pm", so removed one whole 4 h slot from that part also
+
+Here's what I got for session slots in each day and part of day:
+
+```{r}
+library(tidyverse)
+four_h_slots <- 
+  tribble(~part,       ~Th, ~Fr, ~Sa, ~Su,
+          "morning",    36,  31,  30,  28,
+          "afternoon",  35,  32,  30,  0,
+          "evening",    36,  0,    0,  0)
+
+(four_h_slots)
+```
+
+```{r}
+four_h_slots_long <- 
+  four_h_slots %>% 
+  gather(times, slot, -part)
+
+# how many 4 h slots
+n_four_h_slots <-  sum(four_h_slots_long$slot)
+```
+
+So that's a total of `r n_four_h_slots` 4 hour slots, in which we have to fit all the organised and general sessions. This does not include the posters.
+
+## What are the durations of the sessions?
+
+We've received information from the SAA about the length of specific sessions, so let's use that:
+
+```{r}
+saa_info_on_session_duration <- 
+  read_excel("../data/raw_data/saa all organizedsession_timeslotABxls.xls", sheet = 2)
+
+# omit rows with no session id
+saa_info_on_session_duration <- 
+  saa_info_on_session_duration %>% 
+  filter(!is.na(`SESSION ID`)) %>% 
+  mutate(`Session Id` = as.character(`SESSION ID`))
+
+# join with our sessions 
+all_nonposter_sessions_with_saa_info <- 
+  all_nonposter_sessions %>% 
+  left_join(saa_info_on_session_duration)
+
+# papers in a session with saa info on duration 
+all_nonposter_sessions_have_saa_info <- 
+  all_nonposter_sessions_with_saa_info %>% 
+    filter(!is.na(TYPE))
+
+# papers in a session without saa infor on duration (general sessions?)
+all_nonposter_sessions_not_have_saa_info <- 
+  all_nonposter_sessions_with_saa_info %>% 
+  filter(is.na(TYPE))
+```
+
+## How can we prevent one person scheduled in two simultaneous sessions?
+
+Because we will have many parallel sessions, we need to take care that we don't put one person in two simultaneous sessions.
+
+We start by taking a day-part, assign a session to it, and then looping over that day-part to add another session, and another, and so on.
+Each time we add a session to that day-part, we check to see if the names in the incoming session already exist in the sessions we previously added to this day-part.
+If a name has already been added to that day-part, then we skip adding the session, and move to the next session.
+We keep doing this until the tally of hours of all the sessions in the day-part matches the total number of hours we have for that day-part.
+
+We can compute the total capacity in hours of each day-part like this:
 
@@
 four_h_slots_long %>% 
-  mutate(day_part = paste0(times, "-", time)) %>% 
-  select(-time, -times) %>% 
+  mutate(day_part = paste0(times, "-", part)) %>% 
+  select(-part, -times) %>% 
   rename(end = cumsum) %>% 
-  filter( slot != 0)
+  filter(slot != 0)
 
 # get the number of the start slot to go with the number of the end slot
+ xx$start <- 0
 for(i in 1:(nrow(xx))) {
-
   xx$start[i] <- 
@@
             as.numeric(xx[ i-1, 2]) + 1)
+}
+# each slot is 4 hours long
+total_cap <- 
+  xx %>% 
+  mutate(total_capacity = slot * 4) %>% 
+  select(day_part, slot, total_capacity)
+total_cap
+```
+
+Now we will takesession from the big table of sessions and add it into, for example, Th-morning. We'll keep adding more sessions until we reach the capacity of that day-part.
+
+But, we also need to loop over the sessions as we add more to the day-part to check to see if a person's name is already there or not.
+If the name is already there, we'll skip adding that session, and try the next session.
+
+```{r while-loop, echo = TRUE, eval = TRUE}
+
+# get all the sessions, but we're going to modify this, so make a copy
+all_nonposter_sessions_have_saa_info$session_time_allowed <- 
+  as.numeric(all_nonposter_sessions_have_saa_info$TIME)
+z7 <- all_nonposter_sessions_have_saa_info
+all_of_the_day_parts <- NULL
+total_cap$hours_of_sessions <- 0
+total_cap$n_sessions <- 0
+total_cap$c_sessions <- 0
+container_for_this_day_part_with_day_from_while <- 0
+i <- 1
+
+# start with each day-part...
+for(i in seq_len(nrow(total_cap))) {
+
+  # stop loop gracefully when we have assigned all people to day-parts
+  if (nrow(z7) == 0){ break }
+
+  # how many rows (people) to we have remaning to allocate to a day-part?
+  print(glue('number of people remaining to classify is {nrow(z7)}'))
+
+  target_capacity <- total_cap$total_capacity[i] - 4
+
+  # prepare a while statement that will take a session, 
+  # check to see if there is a name clash with sessions,
+  # skip adding it if there is clash, then check to see
+  # the culumative duration of the sessions in the day-part 
+  # has exceeeded the total duration of the day-part,
+  # and stop adding sessions if it has.
+
+# make sure some variables are empty
+cumulative_duration_of_sessions   <-  0
+container_for_this_day_part <- 0
+all_the_people_in_the_session <- 0
+j <- 1
+
+while(cumulative_duration_of_sessions <= target_capacity & 
+      j <= nrow(z7)) {
+
+    get_a_session <- z7[j, ]
+
+    # get the names of all the people in this session...
+    all_the_people_in_this_session <- 
+      all_nonposter_sessions_have_saa_info %>% 
+      filter(`Session Id` == get_a_session$`Session Id`)
+
+    # get names from what we already have added to this day-part
+     if (j == 1) {
+      # do nothing is this is the first incoming session 
+       distinct_names_for_each_session_so_far <- 0
+    } else {
+      # get distinct names for each session previously added
+      # to this day-part
+      # (don't care about same name multiple times in one session)
+      distinct_names_for_each_session_so_far <- 
+      container_for_this_day_part %>% 
+        group_by(uid) %>% 
+        distinct(uid, .keep_all = TRUE) %>% 
+        pull(uid)
+    }
+
+    # compare names already in here to names we are about to add
+    if (j == 1) {
+      # do nothing 
+      names_already_in <- 0
+      } else {
+      # check to see if incoming session has names that already exist in this day_part
+      names_already_in <- sum(all_the_people_in_this_session$uid %in% distinct_names_for_each_session_so_far)
+      }
+
+    # accumulate the incoming session onto the sessions
+    # we've already added to this day-part, if there are no
+    # name conflicts
+    if (j == 1) {
+       container_for_this_day_part <- 
+        all_the_people_in_this_session
+      } else {
+      if (names_already_in == 0) {
+      # if there are no name clashes, 
+      # add these names to previous names in this day-part
+
+      # bind rows of session that has no clashing  names with 
+      # sessions we already added to this day-part
+      container_for_this_day_part <- 
+        bind_rows(all_the_people_in_this_session,
+                  container_for_this_day_part)
+    }  else {
+
+        # do not add this session to the day-part
+      }
+    }
+
+    # compute the duration of session that we've added so far to 
+    # this day-part
+    cumulative_duration_of_sessions <- 
+      container_for_this_day_part %>% 
+      distinct(`Session Id`, .keep_all = TRUE) %>% 
+      pull(session_time_allowed) %>% 
+      sum
+
+    # store this result
+    total_cap$hours_of_sessions[i] <- cumulative_duration_of_sessions
+
+    # show a check of what's happening
+    if (j == 1){
+    progress_check <-    
+      data_frame(i = i, 
+                 j = j,
+               'cumulative duration' = cumulative_duration_of_sessions,
+               'sessions added' =  length(unique(container_for_this_day_part$`Session Id`)),
+               people =  nrow(container_for_this_day_part)
+                )
+    } else {
+      progress_check <- bind_rows(progress_check, 
+                                  data_frame(i = i,
+                                             j = j,
+               'cumulative duration' = cumulative_duration_of_sessions,
+               'sessions added' =  length(unique(container_for_this_day_part$`Session Id`)),
+               people =  nrow(container_for_this_day_part)
+                ))
+    }
+
+       # print(progress_check)
+
+       # move to next session
+       j <- j + 1
+    } 
+# end of the while loop
+
+ # add an identifier so we can see where we have assigned each session so far
+ container_for_this_day_part_with_day_from_while <- 
+   container_for_this_day_part %>% 
+   mutate(day_part_from_while  = as.character(total_cap$day_part[i]))
+
+ # bind rows to previous day-parts we might have done
+  if (i == 1) {
+
+  all_of_the_day_parts <- container_for_this_day_part_with_day_from_while
+
+  } else {
+
+    # check to see
+
+  all_of_the_day_parts <- bind_rows(container_for_this_day_part_with_day_from_while, 
+                                      all_of_the_day_parts)
+  }
+
+ # # progress check
+ print(glue('i = {i}'))
+ # 
+ # print(
+ # all_of_the_day_parts %>% 
+ #   group_by(day_part_from_while) %>% 
+ #   summarise(papers = n(),
+ #             sessions = length(unique(`Session Id`))))
+
+ total_cap$n_sessions[i] <- length(unique(container_for_this_day_part_with_day_from_while$`Session Id`))
+
+ total_cap$c_sessions[i] <-  sum(total_cap$n_sessions)
+
+ print(total_cap)
+
+ # for the next day-part, we need to make sure
+ # we don't add the same session twice to different day-parts
+ # so let's update our big table of abstracts, and drop the session that 
+ # we've already got in the day-part
+
+ z7 <-  
+   z7 %>% 
+   filter(! `Session Id` %in% unique(container_for_this_day_part$`Session Id`))
 
 }
+# end of for loop
 
-intervals <- 
-xx %>% 
-  select(-slot) %>% 
-  select(day_part, start, end )
+# out put the results
+write.csv(all_of_the_day_parts, 
+          "../data/derived_data/output_of_day_part_classification.csv")
+```
 
-elements <- session_with_limits$computed_slot
+What did we get from this process?
+The table below shows each day-part, the number of hours that we have available to fill (`total_capacity`) and the number of hours that we have filled so far (`hours_of_sessions`):
 
-library(fuzzyjoin)
-join_intervals_and_elements <- 
-fuzzy_left_join(data.frame(elements), 
-                intervals, 
-                by = c("elements" = "start", 
-                       "elements" = "end"), 
-                match_fun = list(`>=`, `<=`)) %>% 
-  distinct()
+```{r}
+total_cap %>% 
+  rename(`Number of sessions` = n_sessions,
+         `Cumulative number of sessions` = c_sessions)
+```
 
-# Which slot goes into which part of what day
-slot_day_part <- 
-join_intervals_and_elements %>% 
-  arrange(elements, day_part)
+Have we succeeded in avoiding placing one person in two sessions scheduled in the same day-part?
 
-# now we put the sessions on also
-session_slot_day_part <- 
-  session_with_limits %>% 
-  left_join(slot_day_part, by  = c("computed_slot" = "elements"))
+Let's see if there are any names that occur in two or more distinct sessions in the same day-part.
+
+```{r while-loop-testing, results = 'hide'}
+# check to see what we got from the loop...
+sessions_per_day <- 
+all_of_the_day_parts %>% 
+  group_by(day_part_from_while) %>% 
+  summarise(n_sessions = length(unique(`Session Id`)))
+
+hours_per_day <- 
+  all_of_the_day_parts %>% 
+  group_by(day_part_from_while, `Session Id`) %>% 
+  slice(1) %>% 
+  summarise(n_hours = sum(session_time_allowed)) %>% 
+  ungroup() %>% 
+  group_by(day_part_from_while) %>% 
+  summarise(n_hours = sum(n_hours))
+
+n_missing_papers <- 
+  nrow(all_nonposter_sessions) - nrow(all_of_the_day_parts)
+
+n_missing_sessions <- 
+  length(unique(all_nonposter_sessions$`Session Id`)) - length(unique(all_of_the_day_parts$`Session Id`))
+
+missing_sessions <- 
+  unique(all_nonposter_sessions$`Session Id`)[!(unique(all_nonposter_sessions$`Session Id`) %in% unique(all_of_the_day_parts$`Session Id`)) ]
+
+# have we succeeded with no name clashes?
+# in each day-part, do any names occur more than once?
+
+# how to get it? testing...
+ex <- 
+tribble(~'day_part', ~'sess', ~'auth',
+        "th-m",       1,        'al', 
+        "th-a",       2,        'be',
+        "th-a",       3,        'be',
+        "th-e",       4,        'ce',
+        "fr-a",       5,        'de',
+        "fr-a",       5,        'de')
+
+out <- 
+ex %>% 
+  group_by(day_part, auth) %>% 
+  filter(n() > 1) %>% 
+  group_by(sess) %>% 
+  filter(n() == 1)
+
+# deploy...
+clash <- 
+all_of_the_day_parts %>% 
+  # get authors that have multiple rows (roles or sessions) day-part
+  group_by(day_part_from_while, uid) %>% 
+  filter(n() > 1  ) %>% 
+  # get rows where 
+  group_by(`Session Id`) %>% 
+  filter(n() == 1) # get rows where there are names that occur in more than one session in each day-part
 
@@
 
-## Working in a local Docker container
+Currently we have `r nrow(clash)` schedule clashes, that is, people who are scheduled to be in two or more sessions that are scheduled in the same day-part.
+That's `r ifelse(nrow(clash) == 0, "great", "not ideal")`.
 
-To run this project in a local Docker container, I start a bash shell in the project directory:
+## Summary
 
-Do this one time only, build my Docker container from Dockerfile at top-level in my project:
+Now we've got all the organised sessions scheduled into day-parts with no clashes.
 
-```
-docker build - < Dockerfile
-docker ps
-```
+What do we stil need to do?
 
-Take note of the container ID, which you get after `docker ps`
-
-Then run my Docker container, access it via the web browser, link the project dir to the Docker drive. Make sure you're in a local directory that is sharable (C:\Users\yourname is usually good):
-
-```
-docker run -dp 8787:8787 -v ${pwd}:/home/rstudio/ -e ROOT=TRUE  <container ID>
-```
-
-Then go to `http://192.168.99.100:8787/` or `localhost:8787` in youor browser, log in (rstudio/rstudio), and start your RStudio project by double-clickin on the .Rproj file
-
-When you're done, in the shell, stop all docker containers with:
-
-```
-docker stop $(docker ps -a -q)
-```
-
-Or do `docker ps` and `docker stop <container ID>` just to stop one container.
+- Put the organised sessions into room slots (though not specific rooms).
+For example, if Thurs morning has 36 room slots, we need to distribute the sessions assigned to Thurs monrning accross these slots.
+This is where we can use the multiple knapsack algorithm, so that we can put two 2 hour sessions into one 4 hour slot for a room.
+- Put the general sessions and posters into day-parts and rooms with no clashes    
+- Deal with the special requests for certain people to be scheduled at certain times.
 

### 2017-11-03 analysis/paper/saa2018_nonposter_sessions_while_loop.Rmd
all papers allocated, no conflicts, plotted

@@
   message = FALSE,
-  echo = TRUE,
+  echo = FALSE,
   comment = "#>",
@@
 library(knitr)
-library(kableExtra)
 library(glue)
+
 ```
@@
 gen_sessions <- read_excel("../data/raw_data/01 - gen_paper_sessions_FINAL(oct5)_DG_including_moves v2.xlsx")
+
+gen_sessions <- 
+gen_sessions %>% 
+  mutate(`Session Id` = SESSION) %>% 
+  mutate(`Session Id` =  ifelse(`Session Id` == "35B",
+          max(as.numeric(`Session Id`), na.rm = TRUE) + 1, 
+          as.numeric(`Session Id`) ))
 
@@
 # how many unique general sessions? 
-n_gen_sessions <- length(unique(gen_sessions$SESSION))
+n_gen_sessions <- length(unique(gen_sessions$`Session Id`))
 
@@
 # combine organised and general sessions
-org_sessions$`Session Id` <- as.character(org_sessions$`Session Id`)
-gen_sessions$`Session Id` <- gen_sessions$SESSION
+org_sessions$`Session Id` <- as.numeric(org_sessions$`Session Id`)
 gen_sessions$`Session Type` <- "Symposium"
@@
 
-We have this many sessions in each major category:
+We have this many sessions in each major category, according to my tally of the individual papers:
 
@@
 ```{r}
-saa_info_on_session_duration <- 
-  read_excel("../data/raw_data/saa all organizedsession_timeslotABxls.xls", sheet = 2)
+saa_info_on_organised_session_duration <- 
+  read_excel("../data/raw_data/saa all organizedsession_timeslotABv3.xls", sheet = 2) %>% 
+  mutate(`Session Id` = `SESSION ID`)
 
-# omit rows with no session id
-saa_info_on_session_duration <- 
-  saa_info_on_session_duration %>% 
-  filter(!is.na(`SESSION ID`)) %>% 
-  mutate(`Session Id` = as.character(`SESSION ID`))
+computed_general_session_duration <- 
+  gen_sessions %>% 
+  group_by(`Session Id`) %>% 
+  summarise(n_papers = n()) %>% 
+  mutate(TIME = n_papers * 0.25) %>%  # 15 mins per paper
+  mutate(`Session Id` =  ifelse(`Session Id` == "35B",
+          max(as.numeric(`Session Id`), na.rm = TRUE) + 1, 
+          as.numeric(`Session Id`) ))
 
-# join with our sessions 
-all_nonposter_sessions_with_saa_info <- 
+# combine general and organised session durations
+org_and_gen_session_durations <- 
+  saa_info_on_organised_session_duration %>% 
+  filter(!is.na(`Session Id`)) %>% 
+  mutate(n_papers = `N papers`,
+         TIME = as.numeric(TIME)) %>% 
+  select(`Session Id`, TIME, n_papers) %>% 
+  rbind(computed_general_session_duration) %>% 
+  filter(!is.na(TIME)) %>% 
+  # add 15 m min  to each session for a changeover time
+  mutate(TIME = if_else(TIME == 4.0, 
+                        TIME,
+                        TIME + 0.25))
+
+head(org_and_gen_session_durations)
+
+# join the session duration onto each paper
+
+all_nonposter_sessions_with_session_durations <- 
   all_nonposter_sessions %>% 
-  left_join(saa_info_on_session_duration)
+  mutate(`Session Id` = as.numeric(`Session Id`)) %>% 
+  left_join(org_and_gen_session_durations)
 
-# papers in a session with saa info on duration 
-all_nonposter_sessions_have_saa_info <- 
-  all_nonposter_sessions_with_saa_info %>% 
-    filter(!is.na(TYPE))
-
-# papers in a session without saa infor on duration (general sessions?)
-all_nonposter_sessions_not_have_saa_info <- 
-  all_nonposter_sessions_with_saa_info %>% 
-  filter(is.na(TYPE))
+head(all_nonposter_sessions_with_session_durations)
 ```
@@
   select(day_part, slot, total_capacity)
+
+# View
 total_cap
@@
 
-Now we will takesession from the big table of sessions and add it into, for example, Th-morning. We'll keep adding more sessions until we reach the capacity of that day-part.
+Now we will take session from the big table of sessions and add it into, for example, Th-morning. We'll keep adding more sessions until we reach the capacity of that day-part.
 
@@
 
-```{r while-loop, echo = TRUE, eval = TRUE}
+```{r while-loop, echo = TRUE, eval = TRUE cache = TRUE}
 
-# get all the sessions, but we're going to modify this, so make a copy
-all_nonposter_sessions_have_saa_info$session_time_allowed <- 
-  as.numeric(all_nonposter_sessions_have_saa_info$TIME)
-z7 <- all_nonposter_sessions_have_saa_info
+# prepare some variables... get all the sessions, 
+# but we're going to modify this, so make a copy
+org_and_gen_session_durations$session_time_allowed <- 
+  as.numeric(org_and_gen_session_durations$TIME)
+all_nonposter_sessions_with_session_durations$session_time_allowed <- 
+  all_nonposter_sessions_with_session_durations$TIME
+z7 <- all_nonposter_sessions_with_session_durations
 all_of_the_day_parts <- NULL
@@
 
-  target_capacity <- total_cap$total_capacity[i] - 4
+  # add some buffer time to nudge it back so we don't overfill a day-part
+  # maybe if this buffer is bigger, the knapsack will be faster?
+  # A big enough buffer should avoid the 0 group in the knapsack results
+  # which indicates that the knapsack fn can't fit all the sessions in the
+  # day-part
+  buffer <- 30
+  target_capacity <- total_cap$total_capacity[i] - buffer
 
@@
   # skip adding it if there is clash, then check to see
-  # the culumative duration of the sessions in the day-part 
-  # has exceeeded the total duration of the day-part,
+  # the cumulative duration of the sessions in the day-part 
+  # has exceeded the total duration of the day-part,
   # and stop adding sessions if it has.
@@
     all_the_people_in_this_session <- 
-      all_nonposter_sessions_have_saa_info %>% 
+      all_nonposter_sessions_with_session_durations %>% 
       filter(`Session Id` == get_a_session$`Session Id`)
@@
           "../data/derived_data/output_of_day_part_classification.csv")
+
+write.csv(total_cap, "../data/derived_data/total_cap.csv")
+```
+
+```{r}
+# read them in, if we need to
+all_of_the_day_parts <- readr::read_csv("../data/derived_data/output_of_day_part_classification.csv")
+
+total_cap <- readr::read_csv("../data/derived_data/total_cap.csv")
 ```
@@
 
-```{r while-loop-testing, results = 'hide'}
+```{r while-loop-testing, results = 'hide', echo = TRUE}
 # check to see what we got from the loop...
@@
 n_missing_papers <- 
-  nrow(all_nonposter_sessions) - nrow(all_of_the_day_parts)
+  nrow(all_nonposter_sessions_with_session_durations) - nrow(all_of_the_day_parts)
 
 n_missing_sessions <- 
-  length(unique(all_nonposter_sessions$`Session Id`)) - length(unique(all_of_the_day_parts$`Session Id`))
+  length(unique(all_nonposter_sessions_with_session_durations$`Session Id`)) - length(unique(all_of_the_day_parts$`Session Id`))
 
 missing_sessions <- 
-  unique(all_nonposter_sessions$`Session Id`)[!(unique(all_nonposter_sessions$`Session Id`) %in% unique(all_of_the_day_parts$`Session Id`)) ]
+  unique(all_nonposter_sessions_with_session_durations$`Session Id`)[!(unique(all_nonposter_sessions_with_session_durations$`Session Id`) %in% unique(all_of_the_day_parts$`Session Id`)) ]
 
@@
 
+## Organise the sessions in each day-part into 4 hour slots
+
+Now that we have allocated all the organised sessions into day-parts, we need to group them into 4 hour slots so we can assign them to specific rooms.
+
+We can use the mulitple knapsack algorithm for this. For example, Thurs morning has 36 slots of 4 hours each.
+
+We have a series of functions, one for each day-part, then we run them, and write the output to CSV.
+
+```{r}
+# thu-morning
+
+thu_mo <- function(){
+
+# get the slots available for this day-part
+thu_morning_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Th-morning') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+thu_morning_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Th-morning') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, thu_morning_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- thu_morning_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+thu_morning_sessions$computed_slot <- solution$ksack
+
+write.csv(thu_morning_sessions,
+          "../data/derived_data/thu_morning_sessions.csv")
+}
+
+# Here's thursday afternoon:
+
+# thu-afternoon
+thu_af <- function(){
+
+# get the slots available for this day-part
+thu_afternoon_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Th-afternoon') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+thu_afternoon_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Th-afternoon') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, thu_afternoon_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- thu_afternoon_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+thu_afternoon_sessions$computed_slot <- solution$ksack
+
+write.csv(thu_afternoon_sessions, 
+          "../data/derived_data/thu_afternoon_sessions.csv")
+}
+
+# Here's thursday evening:
+
+# thu-evening
+
+thu_ev <- function(){
+
+# get the slots available for this day-part
+thu_evening_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Th-evening') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+thu_evening_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Th-evening') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, thu_evening_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- thu_evening_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+thu_evening_sessions$computed_slot <- solution$ksack
+
+write.csv(thu_evening_sessions, 
+          "../data/derived_data/thu_evening_sessions.csv")
+}
+
+# Here's friday morning:
+
+# fri-morning
+
+fri_mo <- function(){
+
+# get the slots available for this day-part
+fri_morning_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Fr-morning') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+fri_morning_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Fr-morning') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, fri_morning_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- fri_morning_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+fri_morning_sessions$computed_slot <- solution$ksack
+
+write.csv(fri_morning_sessions, 
+          "../data/derived_data/fri_morning_sessions.csv")
+}
+
+# Here's friday afternoon:
+
+# fri-afternoon
+
+fri_af <- function(){
+
+# get the slots available for this day-part
+fri_afternoon_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Fr-afternoon') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+fri_afternoon_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Fr-afternoon') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, fri_afternoon_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- fri_afternoon_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+fri_afternoon_sessions$computed_slot <- solution$ksack
+
+write.csv(fri_afternoon_sessions, 
+          "../data/derived_data/fri_afternoon_sessions.csv")
+}
+
+# Here's Saturday morning:
+
+# sat-morning
+sat_mo <- function(){
+
+# get the slots available for this day-part
+sat_morning_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Sa-morning') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+sat_morning_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Sa-morning') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, sat_morning_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- sat_morning_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+sat_morning_sessions$computed_slot <- solution$ksack
+
+write.csv(sat_morning_sessions, 
+          "../data/derived_data/sat_morning_sessions.csv")
+}
+
+# Here's saturday afternoon:
+
+# sat-afternoon
+
+sat_af <- function(){
+
+# get the slots available for this day-part
+sat_afternoon_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Sa-afternoon') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+sat_afternoon_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Sa-afternoon') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, sat_afternoon_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- sat_afternoon_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+sat_afternoon_sessions$computed_slot <- solution$ksack
+
+write.csv(sat_afternoon_sessions, 
+          "../data/derived_data/sat_afternoon_sessions.csv")
+}
+
+# Here's sunday morning:
+
+# sun-morning
+sun_mo <- function(){
+
+# get the slots available for this day-part
+sun_morning_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Su-morning') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+sun_morning_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Su-morning') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, sun_morning_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- sun_morning_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+sun_morning_sessions$computed_slot <- solution$ksack
+
+write.csv(sun_morning_sessions, 
+          "../data/derived_data/sun_morning_sessions.csv")
+}
+
+```
+
+```{r}
+# run all of these
+thu_mo()
+thu_af()
+thu_ev()
+fri_mo()
+fri_af()
+sat_mo()
+sat_af()
+sun_mo()
+
+# why do these sometime have a 0 group? Because the fn can't fit it all in! So we need to add more buffer to each day-part in the while loop
+```
+
+## Resolving conflicts with requested times
+
+Now let's get all the names of the people in each day-part to check for the special requests that a few people have made:
+
+```{r}
+
+# read in the day-part data from the while-loop
+sessions_in_day_parts <- list.files("../data/derived_data", pattern = "_sessions.csv", full.names = TRUE)
+sessions_in_day_parts_data <- 
+  map(sessions_in_day_parts, 
+       ~read_csv(.x))
+
+# for each day-part, join with the all names of the people in each session
+sessions_in_day_parts_data_with_all_names <- 
+  map(sessions_in_day_parts_data,
+      ~.x %>% 
+        left_join(all_of_the_day_parts, by = 'Session Id')
+      )
+
+# get day-parts to use in file names
+day_parts_of_list <- 
+  map_chr(sessions_in_day_parts_data_with_all_names, 
+    ~unique(.x$day_part_from_while.y))
+
+# write these out to CSV
+walk2(sessions_in_day_parts_data_with_all_names, 
+     day_parts_of_list,
+     ~write_csv(.x,  path = glue("../data/derived_data/{.y}-all-names.csv")))
+
+# any computer slots with 0?
+sessions_in_day_parts_data_with_all_names_df <- 
+bind_rows(sessions_in_day_parts_data_with_all_names)
+
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(computed_slot == 0)
+```
+
+Sat morning is a focus of many of the special requests, let's have a look at that
+
+```{r}
+Sa_morning_names <- sessions_in_day_parts_data_with_all_names[[which(day_parts_of_list == "Sa-morning")]]
+
+# read in the sheet on conflicts
+conflicts <- read_excel("../data/raw_data/Conflicts.xlsx")
+# make a unique name variable
+conflicts$uid <- glue('{conflicts$PRENOM}_{conflicts$NOM}')
+# get the ones for Sa-morning
+Sa_morning_conflicts <- 
+  conflicts %>% 
+  filter(grepl("Sa", CONFLICTS))
+
+# check to see if we have a problem
+Sa_morning_conflict_result <- Sa_morning_names$uid.y[Sa_morning_names$uid.y %in% Sa_morning_conflicts$uid]
+
+# double-check
+double_check <- 
+Sa_morning_names[Sa_morning_names$`Last Name.y` %in% Sa_morning_conflicts$NOM, ]
+# it's ok
+```
+
+We have `r length(Sa_morning_conflict_result)` conflicts on Saturday morning, that's `r ifelse(length(Sa_morning_conflict_result) == 0, "great", "not ideal")`
+
+What about conflicts in other day-parts?
+
+```{r}
+other_conflicts <- 
+  conflicts %>% 
+  filter(!grepl("Sa", CONFLICTS))
+```
+
+There are just `r nrow(other_conflicts)`, so let's do it one by one.
+
+1. Deborah Nichols doesn't want to be in Thurs AM, Fri AM. When do we have her?
+
+```{r}
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(uid.y == "Deborah_Nichols") %>% 
+  select(`Session Id`, uid.y, day_part_from_while.y )
+```
+
+She's ok, great!
+
+2. Leah McCurdy and session 3109, but what is the conflict exactly? Not on Sat morning from 9-11 am.
+
+```{r}
+Sa_morning_names %>% 
+  filter(uid.y == "Leah_McCurdy")
+
+# when is she?
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(uid.y == "Leah_McCurdy") %>% 
+  select(day_part_from_while.y)
+```
+
+She's ok, oon Fr-morning
+
+3. Session 2929 should not be on "Thurs PM, Thurs EVE, Fri PM, SAT AM, SAT PM". When do we have it?
+
+```{r}
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(`Session Id` == 2929) %>% 
+  select(`Session Id`, uid.y, day_part_from_while.y )
+```
+
+This is a problem, we put this session in Th-afternoon.
+
+Can we tack it onto Su-morning, right at the end? According to the results of our while-loop, we have many hours spare on Su-morning:
+
+```{r}
+total_cap
+```
+
+Let's double-check what we've got after the multiple knapsack work:
+
+```{r}
+day_parts_filled_hours <- 
+sessions_in_day_parts_data_with_all_names_df %>% 
+  group_by(`Session Id`) %>% 
+  slice(1) %>% 
+  group_by(day_part_from_while.y) %>% 
+  summarise(hours_in_day_part = sum(session_time_allowed.y ))
+```
+
+It's the same. So how many spare hours do we have on Su-morning?
+
+```{r}
+Su_morning_hours_available <- 
+total_cap %>% 
+  filter(day_part == "Su-morning") %>% 
+  select(total_capacity) %>% 
+  pull()
+
+Su_morning_hours_filled <- 
+  day_parts_filled_hours %>% 
+  filter(day_part_from_while.y == "Su-morning") %>% 
+  select(hours_in_day_part) %>% 
+  pull()
+
+Su_morning_spare_hours <- 
+  Su_morning_hours_available - Su_morning_hours_filled
+
+```
+
+We have `r Su_morning_spare_hours` hours unused on Sunday morning. Heaps of space. So let's move Session 2929 to Sunday morning:
+
+```{r}
+sessions_in_day_parts_data_with_all_names_df <- 
+sessions_in_day_parts_data_with_all_names_df %>% 
+  mutate(day_part_from_while.y = if_else(`Session Id` == 2929, 
+                                         "Su-morning", 
+                                         day_part_from_while.y))
+
+# check it
+check_session_2929_to_sunday_morning <- 
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(`Session Id` == 2929) %>% 
+  select(day_part_from_while.y) %>% 
+  unique()
+
+```
+
 ## Summary
 
-Now we've got all the organised sessions scheduled into day-parts with no clashes.
+Now we've got all the organised sessions scheduled into slots and day-parts with no clashes, and no conflicts with requested times.
+
+```{r}
+# this is our final result
+write_csv(sessions_in_day_parts_data_with_all_names_df,
+          "../data/derived_data/sessions_in_day_parts_data_with_all_names_df.csv")
+
+# can we make a visual schedule?
+x <- 
+sessions_in_day_parts_data_with_all_names_df %>% 
+  group_by(day_part_from_while.y, computed_slot) %>% 
+  summarise(sessions_for_plot = toString(unique(`Session Id`)))
+
+# put day-slots in order
+x$day_part_from_while.y <- factor(x$day_part_from_while.y, 
+                                  levels = total_cap$day_part)
+
+ggplot(x,
+       aes(day_part_from_while.y,
+           computed_slot)) +
+ geom_tile(colour = "black",
+           fill = "white") +
+  geom_text(aes(label = sessions_for_plot), 
+                  size = 3) +
+    scale_y_continuous(breaks = 0:max(x$computed_slot)) +
+  theme_bw() +
+  xlab("") +
+  ylab("Room-ish") +
+  ggtitle("Visual schedule for SAA2019 conference non-poster sessions",
+          subtitle = "The numbers on the plot are the Session Ids. \n We can think of  each row as equivalent to a room at the conference venue. ")
+
+ggsave("../figures/visual_schedule_for_saa2019_conference_non_poster_sessions.png")
+
+```
 
@@
 ```{r}
-# what commit is this file at? You may need to change the path value
-# if your Rmd is not in analysis/paper/
+# what commit is this file at? 
 git2r::repository("../..")

### 2017-11-03 analysis/paper/saa2018_nonposter_sessions_while_loop.Rmd
making a start on the posters

@@
 
-- Put the organised sessions into room slots (though not specific rooms).
-For example, if Thurs morning has 36 room slots, we need to distribute the sessions assigned to Thurs monrning accross these slots.
-This is where we can use the multiple knapsack algorithm, so that we can put two 2 hour sessions into one 4 hour slot for a room.
-- Put the general sessions and posters into day-parts and rooms with no clashes    
-- Deal with the special requests for certain people to be scheduled at certain times.
+- Posters
 

### 2017-11-03 analysis/paper/saa2018_posters.Rmd
making a start on the posters

---
title: "Notes on scheduling the SAA2019 meeting"
author:
  - Ben Marwick
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  tint::tintHtml: 
     self_contained: TRUE
     df_print: paged
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE, warning = FALSE, message = FALSE,}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "figures"
)

library(saaabstracts)
library(readxl)
library(tidyverse)
library(knitr)
library(glue)

```
## Introduction

## How many sessions do we have?

```{r}
library(readxl)

gen_poster_session_posters <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx")

org_poster_sessions_only <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", sheet = "ORGANIZED_POSTER_SESSION_SIZES")

after_hours_poster_sessions_only <- 
  read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", 
             sheet = "AFTER_HOURS") %>% 
  slice(1:9)

# In here we have org and after hours posters
org_and_after_hours_posters_individual <- 
  read_excel("../data/raw_data/Organized Session Abstracts_2018_AB.xlsx") %>% 
  filter(`Session Type` == "Poster Symposium")

# how many unique general sessions? 
n_gen_poster_sessions <- length(unique(gen_poster_sessions$cuts))

# how many unique organised sessions? 
n_org_poster_sessions <- nrow((org_poster_sessions))

# how many unique after hours sessions? 
n_ah_poster_sessions <- nrow(after_hours_poster_sessions_only)

# total number of sessions
total_number_of_sessions <-  n_gen_poster_sessions + n_org_poster_sessions + n_ah_poster_sessions

# combine organised and general sessions
gen_poster_session_posters$session <- gen_poster_session_posters$cuts

org_poster_session_posters <- 
  org_and_after_hours_posters_individual %>% 
  mutate(session = `Session Id`) %>% 
  filter(session %in% org_poster_sessions_only$ID)

common_cols <- intersect(names(gen_poster_session_posters),  names(org_poster_session_posters))

gen_poster_session_posters <- 
  gen_poster_session_posters %>% 
  select(common_cols)

org_poster_session_posters <- 
  org_poster_session_posters %>% 
  select(common_cols)

gen_and_org_poster_session_posters <- 
  bind_rows(gen_poster_session_posters, 
            org_poster_session_posters)

```

We have information about how many posters can go into each session

```{r}
# poster session durations
poster_session_durations <- 
  read_excel("../data/raw_data/saa all organizedsession_timeslotABv3.xls", 
             sheet = "Posters1")
```

What do we stil need to do?

- Posters

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
git2r::repository("../..")
```

### 2017-11-03 analysis/paper/saa2018_nonposter_sessions_while_loop.Rmd
csv for visual sched of nonposter sessions

@@
 
+# make a spreadsheet that looks like this plot
+
+library(tidyr)
+visual_schedule_plot_as_spreadsheet <- 
+  x %>% 
+  select(day_part_from_while.y, 
+         sessions_for_plot, 
+         computed_slot) %>% 
+  distinct(sessions_for_plot, 
+           .keep_all = TRUE) %>% 
+  spread(day_part_from_while.y, 
+         sessions_for_plot, 
+         fill = "")
+
+write_csv(visual_schedule_plot_as_spreadsheet,
+          "../data/derived_data/visual_schedule_plot_as_spreadsheet.csv")
+
 ```
@@
 What do we stil need to do?
+
+- Many sessions of the same topic are put in the same day-part, which is undesirable.
+We may want to consider changing the while-loop to fill each day-part with one session per day-part, rather than filling a day-part, then moving to the next day-part, etc.
 

### 2017-11-03 analysis/paper/saa2018_posters.Rmd
posters need to avoid clash with papers

@@
              sheet = "Posters1")
+# add day-part to compare with papers
+poster_session_durations <- 
+  poster_session_durations %>% 
+  mutate(start = as.character(start)) %>% 
+  mutate(end = as.character(end)) %>% 
+  mutate(day_part = case_when(
+    start == "1899-12-31 08:00:00" ~ "morning",
+    start == "1899-12-31 10:30:00" ~ "morning",
+    start == "1899-12-31 14:00:00" ~ "afternoon",
+    start == "1899-12-31 17:00:00" ~ "evening"
+  ))
+```
+
+We want to avoid people giving a poster and a talk at the same time, so we need to check with the paper schedule:
+
+```{r}
+sessions_in_day_parts_data_with_all_names_df <- 
+  read_csv("../data/derived_data/sessions_in_day_parts_data_with_all_names_df.csv")
 ```

### 2017-11-04 analysis/paper/saa2018_posters.Rmd
got the while loop going for posters

--- 2017-11-03 analysis/paper/saa2018_posters.Rmd
@@
              sheet = "AFTER_HOURS") %>% 
-  slice(1:9)
+  slice(1:9) %>% 
+  mutate(`Session Id` = as.numeric(gsub("\\D", "", ID)))
 
@@
 # how many unique general sessions? 
-n_gen_poster_sessions <- length(unique(gen_poster_sessions$cuts))
+n_gen_poster_sessions <- length(unique(gen_poster_session_posters$cuts))
 
 # how many unique organised sessions? 
-n_org_poster_sessions <- nrow((org_poster_sessions))
+n_org_poster_sessions <- nrow((org_poster_sessions_only))
 
@@
 
+# check the number of gen and org poster sessions:
+gen_and_org_poster_sessions_only <- 
+  gen_and_org_poster_session_posters %>% 
+  group_by(session) %>% 
+  slice(1)
+
 ```
@@
 # poster session durations
+library(stringi)
 poster_session_durations <- 
@@
              sheet = "Posters1")
+
+# add day-part to compare with papers
+poster_session_durations <- 
+  poster_session_durations %>% 
+  mutate(start = as.character(start)) %>% 
+  mutate(end = as.character(end)) %>% 
+  mutate(day_part = case_when(
+    start == "1899-12-31 08:00:00" ~ "morning",
+    start == "1899-12-31 10:30:00" ~ "morning",
+    start == "1899-12-31 14:00:00" ~ "afternoon",
+    start == "1899-12-31 17:00:00" ~ "evening"
+  )) %>% 
+  mutate(day_part = glue('{stri_trans_totitle(substr(day, 1, 2))}-{day_part}'))
+
+poster_session_durations
+```
+
+Thu-evening is the 'after hours' session, so that's already got the after-hours poster sessions in there.
+We'll just omit it from the following, and we don't need to worry about clashes because it's after all the papers.
+100 posters are permitted in the after-hours session.
+
+```{r}
+# exclude Thurs evening from now on
+poster_session_durations <- 
+  poster_session_durations %>% 
+  filter(day_part != "Th-evening")
+
+# put out the after hours sessions and posters in a CSV
+after_hours_poster_sessions_all_posters <- 
+  org_and_after_hours_posters_individual %>% 
+  filter(`Session Id` %in% after_hours_poster_sessions_only$`Session Id`) %>% 
+  mutate(poster_session_time = "Th-evening")
+
+write_csv(after_hours_poster_sessions_all_posters,
+          "../data/derived_data/after_hours_poster_sessions_all_posters.csv")
+
+# Make sure that we don't have the Thu-evening posters in the remaining posterings
+gen_and_org_poster_sessions_only$session[gen_and_org_poster_sessions_only$session %in% after_hours_poster_sessions_all_posters$`Session Id`] 
+```
+
+We have `r nrow(after_hours_poster_sessions_all_posters)` posters in the after hours session. That's `r ifelse(nrow(after_hours_poster_sessions_all_posters) <= 100, "great", "not ideal")`.
+
+We want to avoid people giving a poster and a talk at the same time, so we need to check with the paper schedule:
+
+```{r}
+sessions_in_day_parts_data_with_all_names_df <- 
+  read_csv("../data/derived_data/sessions_in_day_parts_data_with_all_names_df.csv")
+```
+
+Now we need to put the poster sessions into the remaining slots while checking to see if one of the poster authors is already scheduled to give a paper.
+
+```{r}
+# prepare sessions we have
+  poster_session_durations_summary <- 
+    poster_session_durations %>% 
+    group_by(day_part) %>% 
+    summarise(`N max` = sum(`N max`))
+
+  # make a unique value for names
+  gen_and_org_poster_sessions_only <- 
+  gen_and_org_poster_sessions_only %>% 
+    mutate(uid = glue('{`First Name`}_{`Last Name`}'))
+
+  gen_and_org_poster_session_posters <- 
+    gen_and_org_poster_session_posters %>% 
+    mutate(uid = glue('{`First Name`}_{`Last Name`}'))
+
+  # we're going to change these
+  gen_and_org_poster_sessions_only_tmp <- gen_and_org_poster_sessions_only
+  container_for_this_day_part <- 0
+  poster_session_durations_summary$numbers_of_posters <- 0
+  poster_session_durations_summary$n_posters <- 0
+  poster_session_durations_summary$n_sessions <- 0
+  poster_session_durations_summary$c_sessions <- 0
+  i <- 1
+
+# loop over each day-slot available for poster sessions
+for(i in seq_len(nrow(poster_session_durations_summary))) {
+
+    # stop loop gracefully when we have assigned all people to day-parts
+  if (nrow(gen_and_org_poster_sessions_only_tmp) == 0){ break }
+
+    # how many rows (posters) to we have remaning to allocate to a day-part?
+  print(glue('number of poster sessions remaining to classify is {nrow(gen_and_org_poster_sessions_only_tmp)}'))
+
+  # need a buffer
+  buffer <- 10
+  target_capacity <- poster_session_durations_summary$`N max`[i] - buffer
+
+  # prepare variables for the next round, just to be sure
+  cumulative_number_of_posters <- 0
+  container_for_this_day_part <- 0
+  this_session_posters <- 0
+  j <- 1
+
+  while(cumulative_number_of_posters <= target_capacity & j <= nrow(gen_and_org_poster_sessions_only_tmp)){
+
+  # get a session 
+  this_session <- gen_and_org_poster_sessions_only[j, ]
+
+  # get names for all the people in this poster session
+  this_session_posters <- 
+    gen_and_org_poster_session_posters %>% 
+      filter(session == this_session$session)
+
+#-------------------------------------------------------------------  
+  # check to see if these names are already in this day-part
+  # or in the paper sessions
+  if (j == 1){
+
+    # do nothing is this is the first incoming session 
+    names_already_in <-  0
+    names_already_in2 <- 0
+
+  } else {
+      # get distinct names for each session previously added
+      # to this day-part
+      # (don't care about same name multiple times in one session)
+      distinct_names_for_each_session_so_far <- 
+      container_for_this_day_part %>% 
+        group_by(uid) %>% 
+        distinct(uid, .keep_all = TRUE) %>% 
+        pull(uid)
+
+      # check to see if incoming session has 
+      # names that already exist in this day_part
+      names_already_in <- sum(this_session_posters$uid %in% distinct_names_for_each_session_so_far)
+
+      # check to see if papers for this day-part already have the name in it
+      names_papers_in_this_day_part <- 
+        sessions_in_day_parts_data_with_all_names_df %>% 
+        filter(day_part_from_while.y == poster_session_durations_summary[i,]$day_part) %>% 
+        pull(uid.y)
+
+      names_already_in2 <- sum(this_session_posters$uid %in% names_papers_in_this_day_part)
+
+      #  what if we remove the condition that there shoudl not be a clash with papers?
+      names_already_in2 <- 0
+
+  }
+
+#-------------------------------------------------------------------
+    # if there are no name conflicts, then
+    # accumulate the incoming session onto the sessions
+    # we've already added to this day-part
+
+    if (j == 1) {
+
+       container_for_this_day_part <- 
+        this_session_posters
+
+      } else {
+      if (names_already_in == 0 & names_already_in2 == 0) {
+
+      # if there are no name clashes, 
+      # add these names to previous names in this day-part
+      # bind rows of session that has no clashing  names with 
+      # sessions we already added to this day-part
+      container_for_this_day_part <- 
+        bind_rows(this_session_posters,
+                  container_for_this_day_part)
+
+    }  else {
+
+      # do not add this session to the day-part 
+      # because at least one of the names is already in there
+      }
+      }
+
+#-------------------------------------------------------------------  
+  # compute the number of posters we've added to this day-part
+  # by counting the abstract Id, because this excludes the session chairs
+    cumulative_number_of_posters <- 
+      length(!is.na(container_for_this_day_part$`Abstract Id`))
+
+   # store this result
+    poster_session_durations_summary$numbers_of_posters[i] <- cumulative_number_of_posters
+
+    print(glue('cumulative number of posters: {cumulative_number_of_posters}, j = {j}, i = {i}'))
+
+  # go to next session
+  j <- j + 1
+  }
+  # end of while loop
+
+   # add an identifier so we can see where we have assigned each session so far
+ container_for_this_day_part_with_session_from_while <- 
+   container_for_this_day_part %>% 
+   mutate(day_part_from_while  = as.character(poster_session_durations_summary$day_part[i]))
+
+ # bind rows to previous day-parts we might have done
+  if (i == 1) {
+
+  all_of_the_day_parts <- container_for_this_day_part_with_session_from_while
+
+  } else {
+
+    # check to see
+
+  all_of_the_day_parts <- bind_rows(container_for_this_day_part_with_session_from_while, 
+                                      all_of_the_day_parts)
+  }
+
+  poster_session_durations_summary$n_posters[i] <- length(unique(container_for_this_day_part_with_session_from_while$`Abstract Id`))
+
+ poster_session_durations_summary$n_sessions[i] <- length(unique(container_for_this_day_part_with_session_from_while$session))
+
+ poster_session_durations_summary$c_sessions[i] <-  sum(poster_session_durations_summary$n_sessions)
+
+ print(poster_session_durations_summary %>% select(`N max`, 
+                                                   day_part,
+                                                   n_posters,
+                                                   n_sessions, 
+                                                   c_sessions))
+
+ # for the next day-part, we need to make sure
+ # we don't add the same session twice to different day-parts
+ # so let's update our big table of posters, and drop the session that 
+ # we've already got in the day-part
+
+gen_and_org_poster_sessions_only_tmp <-  
+   gen_and_org_poster_sessions_only_tmp %>% 
+   filter(! session %in% unique(container_for_this_day_part$session))
+
+}
+# end of for loop
+
+# out put the results
+write.csv(all_of_the_day_parts, 
+          "../data/derived_data/posters_output_of_session_classification.csv")
+
+write.csv(poster_session_durations_summary, "../data/derived_data/poster_session_durations.csv")
 ```

### 2017-11-04 analysis/paper/saa2018_posters.Rmd
got the while loop going for posters

@@
 
+after_hours_poster_sessions_all_posters <- 
+  org_and_after_hours_posters_individual %>% 
+  filter(`Session Id` %in% after_hours_poster_sessions_only$`Session Id`) %>% 
+  mutate(poster_session_time = "Th-evening")
+
 ```
 
-We have information about how many posters can go into each session
+We have `r nrow(n_ah_poster_sessions)` sessions of after-hours posters, for a total of `r nrow(after_hours_poster_sessions_all_posters)` posters.
+
+We have `r nrow(gen_and_org_poster_sessions_only)` sessions of general and organised session posters, for a total of `r nrow(gen_and_org_poster_session_posters)` posters.
+
+We have from the SAA information about how many posters can go into each session
 
@@
 
+There are spaces for a total of `r sum(poster_session_durations$'N max')` posters.
+We have received `r nrow(gen_and_org_poster_session_posters)` poster submissions.
+So we have `r nrow(gen_and_org_poster_session_posters) - sum(poster_session_durations$'N max')` more posters than available spaces.
+
 Thu-evening is the 'after hours' session, so that's already got the after-hours poster sessions in there.
@@
 # put out the after hours sessions and posters in a CSV
-after_hours_poster_sessions_all_posters <- 
-  org_and_after_hours_posters_individual %>% 
-  filter(`Session Id` %in% after_hours_poster_sessions_only$`Session Id`) %>% 
-  mutate(poster_session_time = "Th-evening")
-
 write_csv(after_hours_poster_sessions_all_posters,
@@
   poster_session_durations_summary$c_sessions <- 0
+  poster_session_durations_summary$c_posters <- 0
+  all_of_the_day_parts <- 0
   i <- 1
@@
   # need a buffer
-  buffer <- 10
+  buffer <- 3
   target_capacity <- poster_session_durations_summary$`N max`[i] - buffer
@@
       names_already_in2 <- sum(this_session_posters$uid %in% names_papers_in_this_day_part)
-
-      #  what if we remove the condition that there shoudl not be a clash with papers?
-      names_already_in2 <- 0
 
@@
  poster_session_durations_summary$c_sessions[i] <-  sum(poster_session_durations_summary$n_sessions)
+ poster_session_durations_summary$c_posters[i] <-  sum(poster_session_durations_summary$n_posters)
 
@@
                                                    n_sessions, 
-                                                   c_sessions))
+                                                   c_sessions,
+                                                   c_posters))
 
@@
 
+Seems that we have too many posters for the available space. Let's double-check:
+
+```{r}
+# duplicates?
+all_of_the_day_parts %>% 
+  distinct()
+
+posters_with_abstract_ids <- 
+all_of_the_day_parts %>% 
+  filter(!is.na(`Abstract Id`)) %>% 
+  group_by(day_part_from_while) %>% 
+  tally() %>% 
+  summarise(sum(n)) %>% 
+  pull()
+```
+
+We have `r posters_with_abstract_ids` posters with Abstract IDs (and so not chair/co-chairs).
+Compare to `r nrow(gen_and_org_poster_session_posters)` posters that we started with, and space available for `r sum(poster_session_durations$'N max')` posters.
+
 What do we stil need to do?

### 2017-11-04 analysis/paper/saa2018_posters.Rmd
we have too many posters

@@
 # In here we have org and after hours posters
+
+all_abstracts  <-  read_excel("../data/raw_data/Organized Session Abstracts_2018_AB.xlsx")
+
 org_and_after_hours_posters_individual <- 
-  read_excel("../data/raw_data/Organized Session Abstracts_2018_AB.xlsx") %>% 
-  filter(`Session Type` == "Poster Symposium")
+all_abstracts %>% 
+  filter(`Session Type` == "Poster Symposium") %>% 
+  filter(Role == "Presenter")
+
+gen_and_org_poster_sessions_only <- 
+all_abstracts %>% 
+  filter(`Session Type` == "Poster Symposium") %>% 
+  filter(Role == "Chair (Primary)")
 
@@
 
+# why is each poster in each day-part?
+all_of_the_day_parts %>% 
+ arrange(uid) %>% View
+
+# how many abstracts
 posters_with_abstract_ids <- 

### 2017-11-04 analysis/paper/saa2018_posters.Rmd
posters: fill all day-parts will all posters, no overfill

@@
 
+We can combine general and organised sessions, for the purpose of scheduling. BUt we have to get the after hours sessions and put them aside.
+
 ```{r}
@@
 
-gen_poster_session_posters <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx")
+# Here we get all the general session posters
+# session number is cuts column
+gen_poster_session_posters_individual <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx")
+gen_poster_session_posters_individual$`Session Id` <- gen_poster_session_posters_individual$cuts
+# 422 posters in the general session
 
-org_poster_sessions_only <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", sheet = "ORGANIZED_POSTER_SESSION_SIZES")
+# grab the two general sessions from the 
+# ah posters from the other sheet (very inconvenient!)
+ah_poster_session_posters_individual <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", sheet = "AFTER_HOURS", skip = 11)
+ah_poster_session_posters_individual$`Session Id` <- ah_poster_session_posters_individual$cuts
 
-after_hours_poster_sessions_only <- 
-  read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", 
-             sheet = "AFTER_HOURS") %>% 
-  slice(1:9) %>% 
-  mutate(`Session Id` = as.numeric(gsub("\\D", "", ID)))
+ah_poster_session_posters_individual <- 
+ah_poster_session_posters_individual[, names(ah_poster_session_posters_individual) %in% names(gen_poster_session_posters_individual)  ]
+
+gen_poster_session_posters_individual <- 
+  rbind(gen_poster_session_posters_individual,
+        ah_poster_session_posters_individual)
 
 # In here we have org and after hours posters
+# but not general posters
 
@@
   filter(Role == "Presenter")
+# 196 posters in org + ah sessions
 
-gen_and_org_poster_sessions_only <- 
-all_abstracts %>% 
-  filter(`Session Type` == "Poster Symposium") %>% 
-  filter(Role == "Chair (Primary)")
+# combine general, organised and after hours individual posters
+common_names <- 
+names(gen_poster_session_posters_individual)[names(gen_poster_session_posters_individual) %in% names(org_and_after_hours_posters_individual)]
 
-# how many unique general sessions? 
-n_gen_poster_sessions <- length(unique(gen_poster_session_posters$cuts))
+gen_poster_session_posters_individual <- 
+  gen_poster_session_posters_individual %>% 
+  select(common_names)
 
-# how many unique organised sessions? 
-n_org_poster_sessions <- nrow((org_poster_sessions_only))
+org_and_after_hours_posters_individual <- 
+  org_and_after_hours_posters_individual %>% 
+    select(common_names)
 
-# how many unique after hours sessions? 
-n_ah_poster_sessions <- nrow(after_hours_poster_sessions_only)
+all_posters_individual <- 
+  bind_rows(gen_poster_session_posters_individual,
+            org_and_after_hours_posters_individual) %>% 
+  filter(!is.na(Title))
+```
 
-# total number of sessions
-total_number_of_sessions <-  n_gen_poster_sessions + n_org_poster_sessions + n_ah_poster_sessions
+We have `r nrow(all_posters_individual)` posters altogether, with `r nrow(gen_poster_session_posters_individual)` in the general session, and `r nrow(org_and_after_hours_posters_individual)` in the organised sessions (including after hours posters).
 
-# combine organised and general sessions
-gen_poster_session_posters$session <- gen_poster_session_posters$cuts
+Let's remove the after hours sessions from all the posters, since we know they will be assigned to Th-evening:
 
-org_poster_session_posters <- 
-  org_and_after_hours_posters_individual %>% 
-  mutate(session = `Session Id`) %>% 
-  filter(session %in% org_poster_sessions_only$ID)
+```{r}
+# how many ah sessions?
+after_hours_poster_sessions_only <- 
+  read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", 
+             sheet = "AFTER_HOURS") %>% 
+  slice(1:9) %>% 
+  mutate(`Session Id` = as.numeric(gsub("\\D", "", ID)))
 
-common_cols <- intersect(names(gen_poster_session_posters),  names(org_poster_session_posters))
-
-gen_poster_session_posters <- 
-  gen_poster_session_posters %>% 
-  select(common_cols)
-
-org_poster_session_posters <- 
-  org_poster_session_posters %>% 
-  select(common_cols)
-
-gen_and_org_poster_session_posters <- 
-  bind_rows(gen_poster_session_posters, 
-            org_poster_session_posters)
-
-# check the number of gen and org poster sessions:
-gen_and_org_poster_sessions_only <- 
-  gen_and_org_poster_session_posters %>% 
-  group_by(session) %>% 
-  slice(1)
-
-after_hours_poster_sessions_all_posters <- 
-  org_and_after_hours_posters_individual %>% 
+# how many ah posters?
+after_hours_poster_sessions_posters_individual <- 
+  all_posters_individual %>%
   filter(`Session Id` %in% after_hours_poster_sessions_only$`Session Id`) %>% 
@@
 
+# remove these from all posters
+all_posters_individual_no_ah <- 
+  all_posters_individual %>% 
+  filter(!`Session Id` %in% after_hours_poster_sessions_only$`Session Id`)
+
+# put out the after hours sessions and posters in a CSV
+write_csv(after_hours_poster_sessions_posters_individual,
+          "../data/derived_data/after_hours_poster_sessions_all_posters.csv")
+
+# Make sure that we don't have the Thu-evening posters in the remaining posterings
+all_posters_individual_no_ah$`Session Id`[all_posters_individual_no_ah$`Session Id` %in% after_hours_poster_sessions_posters_individual$`Session Id`] 
 ```
 
-We have `r nrow(n_ah_poster_sessions)` sessions of after-hours posters, for a total of `r nrow(after_hours_poster_sessions_all_posters)` posters.
+We have `r length(unique(after_hours_poster_sessions_only$'Session Id'))` sessions of after-hours posters, with a total of `r nrow(after_hours_poster_sessions_posters_individual)` posters.
+This leaves `r nrow(all_posters_individual_no_ah)` remaning to allocate in the day-parts.
 
-We have `r nrow(gen_and_org_poster_sessions_only)` sessions of general and organised session posters, for a total of `r nrow(gen_and_org_poster_session_posters)` posters.
+100 posters are permitted in the after-hours session.
+We have `r nrow(after_hours_poster_sessions_posters_individual)` posters in the after hours session.
+That's `r ifelse(nrow(after_hours_poster_sessions_posters_individual) <= 100, "great", "not ideal")`.
 
@@
   )) %>% 
+  mutate(start = substr(start, 11, nchar(start))) %>% 
+  mutate(end = substr(end, 11, nchar(end))) %>% 
   mutate(day_part = glue('{stri_trans_totitle(substr(day, 1, 2))}-{day_part}'))
@@
 poster_session_durations
+
+diff_between_need_and_have <- nrow(all_posters_individual_no_ah) - sum(poster_session_durations$'N max')
 ```
@@
 There are spaces for a total of `r sum(poster_session_durations$'N max')` posters.
-We have received `r nrow(gen_and_org_poster_session_posters)` poster submissions.
-So we have `r nrow(gen_and_org_poster_session_posters) - sum(poster_session_durations$'N max')` more posters than available spaces.
+We have received `r nrow(all_posters_individual_no_ah)` poster submissions.
+So we have `r nrow(all_posters_individual_no_ah) - sum(poster_session_durations$'N max')` `r ifelse(diff_between_need_and_have < 0, glue('{abs(diff_between_need_and_have)} free spaces'), glue('{abs(diff_between_need_and_have)} more posters than available spaces'))`.
 
 Thu-evening is the 'after hours' session, so that's already got the after-hours poster sessions in there.
-We'll just omit it from the following, and we don't need to worry about clashes because it's after all the papers.
-100 posters are permitted in the after-hours session.
+We'll just omit it from the following available schedules, and we don't need to worry about clashes because it's after all the papers.
+
+And we'll combine day-parts that are in the same paper day-part, which will make it simper for avoiding clashes.
 
@@
 # exclude Thurs evening from now on
-poster_session_durations <- 
+poster_session_durations_no_th_ev <- 
   poster_session_durations %>% 
-  filter(day_part != "Th-evening")
+  filter(day_part != "Th-evening") %>% 
+  group_by(day_part) %>% 
+  summarise(n_max = sum(`N max`) )
 
-# put out the after hours sessions and posters in a CSV
-write_csv(after_hours_poster_sessions_all_posters,
-          "../data/derived_data/after_hours_poster_sessions_all_posters.csv")
-
-# Make sure that we don't have the Thu-evening posters in the remaining posterings
-gen_and_org_poster_sessions_only$session[gen_and_org_poster_sessions_only$session %in% after_hours_poster_sessions_all_posters$`Session Id`] 
+poster_session_durations_no_th_ev
 ```
-
-We have `r nrow(after_hours_poster_sessions_all_posters)` posters in the after hours session. That's `r ifelse(nrow(after_hours_poster_sessions_all_posters) <= 100, "great", "not ideal")`.
 
@@
 ```{r}
-# prepare sessions we have
-  poster_session_durations_summary <- 
-    poster_session_durations %>% 
-    group_by(day_part) %>% 
-    summarise(`N max` = sum(`N max`))
+# to avoid overfilling day-parts
+buffer <-7 # is the largest to avoid overfill
 
-  # make a unique value for names
-  gen_and_org_poster_sessions_only <- 
-  gen_and_org_poster_sessions_only %>% 
+# day-parts to fill
+ poster_session_durations_no_th_ev <- 
+   poster_session_durations_no_th_ev %>% 
+   arrange(desc(n_max))
+
+# make a unique value for names for all posters
+all_posters_individual_no_ah <- 
+  all_posters_individual_no_ah %>% 
     mutate(uid = glue('{`First Name`}_{`Last Name`}'))
 
-  gen_and_org_poster_session_posters <- 
-    gen_and_org_poster_session_posters %>% 
-    mutate(uid = glue('{`First Name`}_{`Last Name`}'))
+# get table of session id's only 
+poster_session_ids <- 
+unique(all_posters_individual_no_ah$`Session Id`)
 
 # we're going to change these
-  gen_and_org_poster_sessions_only_tmp <- gen_and_org_poster_sessions_only
+  all_posters_individual_no_ah_tmp <- all_posters_individual_no_ah
   container_for_this_day_part <- 0
-  poster_session_durations_summary$numbers_of_posters <- 0
-  poster_session_durations_summary$n_posters <- 0
-  poster_session_durations_summary$n_sessions <- 0
-  poster_session_durations_summary$c_sessions <- 0
-  poster_session_durations_summary$c_posters <- 0
+  poster_session_durations_no_th_ev$numbers_of_posters <- 0
+  poster_session_durations_no_th_ev$n_posters <- 0
+  poster_session_durations_no_th_ev$n_sessions <- 0
+  poster_session_durations_no_th_ev$c_sessions <- 0
+  poster_session_durations_no_th_ev$c_posters <- 0
   all_of_the_day_parts <- 0
@@
 # loop over each day-slot available for poster sessions
-for(i in seq_len(nrow(poster_session_durations_summary))) {
+# i is each day-part
+for(i in seq_len(nrow(poster_session_durations_no_th_ev))) {
 
     # stop loop gracefully when we have assigned all people to day-parts
-  if (nrow(gen_and_org_poster_sessions_only_tmp) == 0){ break }
+  if (nrow(all_posters_individual_no_ah_tmp) == 0){ break }
 
     # how many rows (posters) to we have remaning to allocate to a day-part?
-  print(glue('number of poster sessions remaining to classify is {nrow(gen_and_org_poster_sessions_only_tmp)}'))
+  print(glue('number of poster sessions remaining to classify is {nrow(all_posters_individual_no_ah_tmp)}'))
 
-  # need a buffer
-  buffer <- 3
-  target_capacity <- poster_session_durations_summary$`N max`[i] - buffer
+  # need a buffer to avoid overfilling day-parts
+  target_capacity <- poster_session_durations_no_th_ev$n_max[i] - buffer
 
@@
   this_session_posters <- 0
+  # j is each poster session
   j <- 1
 
-  while(cumulative_number_of_posters <= target_capacity & j <= nrow(gen_and_org_poster_sessions_only_tmp)){
+  while(cumulative_number_of_posters <= target_capacity & 
+        j <= length(poster_session_ids)) {
 
   # get a session 
-  this_session <- gen_and_org_poster_sessions_only[j, ]
+  this_session <- poster_session_ids[j]
 
@@
   this_session_posters <- 
-    gen_and_org_poster_session_posters %>% 
-      filter(session == this_session$session)
+    all_posters_individual_no_ah_tmp %>% 
+      filter(`Session Id` == this_session)
 
@@
         sessions_in_day_parts_data_with_all_names_df %>% 
-        filter(day_part_from_while.y == poster_session_durations_summary[i,]$day_part) %>% 
+        filter(day_part_from_while.y == poster_session_durations_no_th_ev[i,]$day_part) %>% 
         pull(uid.y)
@@
     cumulative_number_of_posters <- 
-      length(!is.na(container_for_this_day_part$`Abstract Id`))
+    nrow(container_for_this_day_part)
 
    # store this result
-    poster_session_durations_summary$numbers_of_posters[i] <- cumulative_number_of_posters
+    poster_session_durations_no_th_ev$n_posters[i] <- cumulative_number_of_posters
 
@@
    container_for_this_day_part %>% 
-   mutate(day_part_from_while  = as.character(poster_session_durations_summary$day_part[i]))
+   mutate(day_part_from_while  = as.character(poster_session_durations_no_th_ev$day_part[i]))
 
@@
 
-  poster_session_durations_summary$n_posters[i] <- length(unique(container_for_this_day_part_with_session_from_while$`Abstract Id`))
+ poster_session_durations_no_th_ev$n_sessions[i] <- length(unique(container_for_this_day_part_with_session_from_while$`Session Id`))
 
- poster_session_durations_summary$n_sessions[i] <- length(unique(container_for_this_day_part_with_session_from_while$session))
+ poster_session_durations_no_th_ev$c_sessions[i] <-  sum(poster_session_durations_no_th_ev$n_sessions)
+ poster_session_durations_no_th_ev$c_posters[i] <-  sum(poster_session_durations_no_th_ev$n_posters)
 
- poster_session_durations_summary$c_sessions[i] <-  sum(poster_session_durations_summary$n_sessions)
- poster_session_durations_summary$c_posters[i] <-  sum(poster_session_durations_summary$n_posters)
-
- print(poster_session_durations_summary %>% select(`N max`, 
+ print(poster_session_durations_no_th_ev %>% select(n_max, 
                                                    day_part,
@@
 
-gen_and_org_poster_sessions_only_tmp <-  
-   gen_and_org_poster_sessions_only_tmp %>% 
-   filter(! session %in% unique(container_for_this_day_part$session))
+all_posters_individual_no_ah_tmp <-  
+   all_posters_individual_no_ah_tmp %>% 
+   filter(! `Session Id` %in% unique(container_for_this_day_part$`Session Id`))
 
@@
 
-write.csv(poster_session_durations_summary, "../data/derived_data/poster_session_durations.csv")
+write.csv(poster_session_durations_no_th_ev, "../data/derived_data/poster_session_durations_no_th_ev.csv")
 ```
 
-Seems that we have too many posters for the available space. Let's double-check:
+We have from the SAA spaces for `r sum(poster_session_durations_no_th_ev$n_max)` posters. And we have `r sum(poster_session_durations_no_th_ev$c_posters)`
+
+It's tricky to allocate the posters so they don't exceed the number allowed in each day part.
 
 ```{r}
+# we have 532 posters not in ah
+x1 <- 
+all_posters_individual_no_ah
+
+x2 <- 
+all_of_the_day_parts
+
+# what sessions are not in there?
+session_output <- 
+x2 %>% 
+  group_by(`Session Id`) %>% 
+  pull(`Session Id`) %>% 
+  unique
+
+session_input <- 
+  x1 %>% 
+  group_by(`Session Id`) %>% 
+  pull(`Session Id`) %>% 
+  unique
+
+setdiff(session_input, session_output)
+
+x1 %>% 
+  filter(is.na(Title))
+
+# what are in the output of this loop that is not in the input?
 # duplicates?

### 2017-11-05 analysis/paper/saa2018_nonposter_sessions_while_loop.Rmd
why did I miss 3115, 3147, 3175, 3116, 3152, 3162? I don't know. Perhapsthere is no perfect solution

@@
 
+    # randomly reorder the rows to see if we can solve a problem
+    # z7 <- z7[sample(nrow(z7)),]
     get_a_session <- z7[j, ]
@@
 
+Did we get all sessions allocated to a day-part?
+
+```{r}
+sessions_we_missed <- 
+  setdiff(unique(all_nonposter_sessions$`Session Id`),
+          unique(all_of_the_day_parts$`Session Id`))
+```
+
+We missed `r length(sessions_we_missed)` sessions: `r sessions_we_missed`.
+
 What did we get from this process?
@@
 
-She's ok, great!
+Sa-afternoon and Sa-morning, she's ok, great!
 
@@
 
-She's ok, oon Fr-morning
+She's ok, on Fr-morning
 
@@
 
-We have `r Su_morning_spare_hours` hours unused on Sunday morning. Heaps of space. So let's move Session 2929 to Sunday morning:
-
-```{r}
+We have `r Su_morning_spare_hours` hours unused on Sunday morning.
+Heaps of space.
+So let's move Session 2929 to Sunday morning.
+Don't forget to update the computed slot value so it gets it's own slot.
+
+```{r}
+
+# what is the max computed_slot for Su-morning?
+max_computed_slot_Su_morning <- 
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(day_part_from_while.y == "Su-morning") %>% 
+  summarise(max_computed_slot_Su_morning = max(computed_slot)) %>% 
+  pull(max_computed_slot_Su_morning)
+
 sessions_in_day_parts_data_with_all_names_df <- 
@@
                                          "Su-morning", 
-                                         day_part_from_while.y))
+                                         day_part_from_while.y)) %>% 
+  mutate(computed_slot = ifelse(`Session Id` == 2929, 
+                                         max_computed_slot_Su_morning + 1, 
+                                         computed_slot))
 
@@
   unique()
+
+# check to see if there are conflicts with names, should be 0
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(day_part_from_while.y == "Su-morning") %>%
+  group_by(`Session Id`, uid.y)  %>% 
+  add_tally() %>% 
+  distinct(`Session Id`) %>% 
+  group_by(uid.y) %>% 
+  filter(n() > 1 )
+
+# how many slots in Sun?
+
+# available?
+total_cap #28
+
+# used?
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(day_part_from_while.y == "Su-morning") %>% 
+  distinct(computed_slot) %>% 
+  max()
+# 21
+
+```
+
+What do we stil need to do?
+
+- Many sessions of the same topic are put in the same day-part, which is undesirable.
+We may want to consider changing the while-loop to fill each day-part with one session per day-part, rather than filling a day-part, then moving to the next day-part, etc.
+
+- AB says
+
+>-  row 10 has session 2929 sharing with session 3165. But session 2929 needs 16 slots, which is 4 hours.
+
+Problem was that I forgot to update the computed slot value. Fixed now.
+
+>- Missing 3115, 3147, 3175, 3116, 3152, 3162
+
+```{r}
+
+all_nonposter_sessions %>% 
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162")) 
+# yes, there
+
+# how many sessions""
+length(unique(all_nonposter_sessions$`Session Id`)) #  271
+
+all_nonposter_sessions_with_session_durations  %>% 
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162")) 
+# yes
+
+org_and_gen_session_durations  %>% 
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162")) 
+# yes
+
+all_of_the_day_parts  %>% 
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162")) 
+# not there
+
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162"))
+# not there
+
+# how many sessions?
+length(unique(sessions_in_day_parts_data_with_all_names_df$`Session Id`)) # 265
+
+# are these sessions very long?
+hours_of_missing_sessions <- 
+all_nonposter_sessions %>% 
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162"))  %>% 
+  group_by(`Session Id`) %>% 
+  tally() %>% 
+  mutate(minutes = n * 15) %>% 
+  mutate(hours = minutes / 60)
+# nope
+total_hours_missing <-  sum(hours_of_missing_sessions$hours)
+
+# what are they about?
+all_nonposter_sessions %>% 
+  filter(`Session Id` %in% missing_sessions) %>% 
+  select(Abstract, Title)
+# no title or abstracts!
+
+```
+
+```{r}
+
+# So, no idea what's going on with these. Let's see if there are any problems with putting them on Sunday
+missing_sessions_details <- 
+  all_nonposter_sessions %>% 
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162")) %>% 
+  mutate(uid.y = uid)
+
+# are any people in more than one session Sunday? yes
+# what about all the day-parts with spare hours
+clashes_on_day_x <- function(day) {
+  sessions_in_day_parts_data_with_all_names_df %>% 
+ filter(day_part_from_while.y == day) %>% 
+  bind_rows(missing_sessions_details) %>% 
+  group_by(`Session Id`, uid.y)  %>% 
+  add_tally() %>% 
+  distinct(`Session Id`) %>% 
+  group_by(uid.y) %>% 
+  filter(n() > 1 )
+}
+
+total_cap_remaining <- 
+  total_cap %>% 
+  mutate(hours_unused =  total_capacity - hours_of_sessions) %>% 
+  select(day_part, hours_unused)
+
+day_with_spare_hours <- c("Fr-morning",
+                          "Fr-afternoon",
+                          "Sa-morning",
+                          "Sa-afternoon",
+                          "Su-morning")
+
+total_cap_remaining %>% 
+  mutate(clashes = map(day_part, clashes_on_day_x)) %>% 
+  mutate(n_clashes = map_int(.$clashes, ~nrow(.x)))
+
+# Looks like Th-morning has no clashes, double-check, should be 0
+
+ sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(day_part_from_while.y ==  "Th-morning") %>% 
+  bind_rows(missing_sessions_details) %>% 
+  group_by(`Session Id`, uid.y)  %>% 
+  add_tally() %>% 
+  distinct(`Session Id`) %>% 
+  group_by(uid.y) %>% 
+  filter(n() > 1 )
+```
+
+Seems that Th-morning is a good place to put these missing sessions. So let's do that. We'll have to double-up in some slots
+
+```{r}
+# find out how many hours we have for each slot on th-morn
+slot_on_thu_morn_with_less_than_4_h <- 
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(day_part_from_while.y == "Th-morning") %>% 
+  distinct(`Session Id`, .keep_all = TRUE) %>% 
+  group_by(computed_slot) %>% 
+  summarise(total_duration = sum(session_time_allowed.y)) %>% 
+  filter(total_duration < 4) %>% 
+  mutate(minutes = total_duration * 60) %>% 
+  mutate(minutes_free = (4*60) - minutes)
+
+# how many hours do we need to put in there?
+minutes_of_missing_sessions <- 
+  hours_of_missing_sessions$minutes
+
+# # what is the max computed_slot for Su-morning?
+# max_computed_slot_Th_morning <- 
+# sessions_in_day_parts_data_with_all_names_df %>% 
+#   filter(day_part_from_while.y == "Th-morning") %>% 
+#   summarise(max_computed_slot_Th_morning = max(computed_slot)) %>% 
+#   pull(max_computed_slot_Th_morning)
+# 
+# # assign missing sessions to Th morn, update slots
+# missing_sessions_details <- 
+# missing_sessions_details %>% 
+#   mutate(day_part_from_while.y = "Th-morning")
+# 
+# missing_sessions_details_indices  <- 
+# missing_sessions_details %>% 
+#   group_by(`Session Id`) %>% 
+#   group_indices()
+# 
+# # update computed slot
+# missing_sessions_details <- 
+#   missing_sessions_details %>% 
+#     mutate(computed_slot = missing_sessions_details_indices + max_computed_slot_Th_morning) 
+```
+
+We can't put them all on Th morn, there's not slot with enough free space to fit more than one of the missing sessions in
+
+```{r}
+
+# # bind to rest of the papers
+# sessions_in_day_parts_data_with_all_names_df <- 
+#   bind_rows(sessions_in_day_parts_data_with_all_names_df ,
+#             missing_sessions_details)
+#             
+# 
+# # check it, should be Th-morning
+# check_missing_sessions_to_thu_morning <- 
+# sessions_in_day_parts_data_with_all_names_df %>% 
+#   filter(`Session Id` %in% missing_sessions) %>% 
+#   select(day_part_from_while.y) %>% 
+#   unique()
+# 
+# # check for conflicts with names
+#  sessions_in_day_parts_data_with_all_names_df %>% 
+#   filter(day_part_from_while.y ==  "Th-morning") %>% 
+#   group_by(`Session Id`, uid.y)  %>% 
+#   add_tally() %>% 
+#   distinct(`Session Id`) %>% 
+#   group_by(uid.y) %>% 
+#   filter(n() > 1 )
 
@@
 
-What do we stil need to do?
-
-- Many sessions of the same topic are put in the same day-part, which is undesirable.
-We may want to consider changing the while-loop to fill each day-part with one session per day-part, rather than filling a day-part, then moving to the next day-part, etc.
-
-- Posters
-
 ### Colophon

### 2017-11-06 analysis/paper/saa2018_posters_knapsack.Rmd
divide up poster sessions with >8 papers and schedule

--- 2017-11-04 analysis/paper/saa2018_posters.Rmd
@@
 
-We can combine general and organised sessions, for the purpose of scheduling. BUt we have to get the after hours sessions and put them aside.
+We can combine general and organised sessions, for the purpose of scheduling. But we have to get the after hours sessions and put them aside.
 
@@
 
-And we'll combine day-parts that are in the same paper day-part, which will make it simper for avoiding clashes.
+```{r}
+# see if we can do the poster day-parts individually
 
-```{r}
-# exclude Thurs evening from now on
 poster_session_durations_no_th_ev <- 
@@
   filter(day_part != "Th-evening") %>% 
-  group_by(day_part) %>% 
-  summarise(n_max = sum(`N max`) )
-
-poster_session_durations_no_th_ev
+  mutate(poster_day_part = glue('poster-{day_part}')) %>% 
+  group_by(poster_day_part) %>% 
+  mutate(label = 1:n()) %>% 
+  ungroup() %>% 
+  mutate(poster_day_part = glue('{poster_day_part}_{label}')) %>% 
+  select(-label) %>% 
+  mutate(n_max = `N max`)
 ```
 
-We want to avoid people giving a poster and a talk at the same time, so we need to check with the paper schedule:
+We're not going to bother with clashes between people giving posters and people giving papers, because they are both 'presenting' roles, and a person can only have one of those.
 
 ```{r}
-sessions_in_day_parts_data_with_all_names_df <- 
-  read_csv("../data/derived_data/sessions_in_day_parts_data_with_all_names_df.csv")
+# check 
+people_with_2_non_poster_roles <- 
+all_abstracts %>% 
+  group_by(`First Name`, `Last Name`) %>% 
+  add_tally() %>% 
+  filter(n >= 2) %>% 
+  filter(!grepl("Poster", `Session Type`)) %>% 
+  mutate(uid = glue('{`First Name`} {`Last Name`}'))
+
+people_in_posters_with_2_roles <- 
+  all_abstracts %>% 
+  group_by(`First Name`, `Last Name`) %>% 
+  add_tally() %>% 
+  filter(n >= 2) %>% 
+  filter(grepl("Poster", `Session Type`)) %>% 
+  mutate(uid = glue('{`First Name`} {`Last Name`}'))
+
+# any overlap?
+people_with_posters_and_another_role <- 
+people_in_posters_with_2_roles$uid[people_in_posters_with_2_roles$uid %in% people_with_2_non_poster_roles$uid]
 ```
 
-Now we need to put the poster sessions into the remaining slots while checking to see if one of the poster authors is already scheduled to give a paper.
+So let's go right onto the knapsack to see if we can fit them all in.
 
 ```{r}
-# to avoid overfilling day-parts
-buffer <-7 # is the largest to avoid overfill
 
-# day-parts to fill
- poster_session_durations_no_th_ev <- 
-   poster_session_durations_no_th_ev %>% 
-   arrange(desc(n_max))
+# number of posters in each session
+posters_to_allocate <- 
+  all_posters_individual_no_ah %>% 
+  group_by(`Session Id`) %>% 
+  tally() %>% 
+  pull(n)
+sum(posters_to_allocate)
 
-# make a unique value for names for all posters
-all_posters_individual_no_ah <- 
+# size of each slot to fill with posters
+poster_sessions_available <- poster_session_durations_no_th_ev$n_max
+sum(poster_session_durations_no_th_ev$n_max)
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(posters_to_allocate, 
+                      posters_to_allocate, 
+                      poster_sessions_available)
+
+solution$ksack
+
+# poster sessions only
+all_posters_individual_no_ah_sessions <- 
   all_posters_individual_no_ah %>% 
-    mutate(uid = glue('{`First Name`}_{`Last Name`}'))
+  group_by(`Session Id`) %>% 
+  select(`Session Id`) %>% 
+  slice(1) %>% 
+  ungroup() %>% 
+  mutate(ksack = solution$ksack)
 
-# get table of session id's only 
-poster_session_ids <- 
-unique(all_posters_individual_no_ah$`Session Id`)
+# put the results back on indiv posters
+all_posters_individual_no_ah_ksack <- 
+  all_posters_individual_no_ah %>% 
+  left_join(all_posters_individual_no_ah_sessions)
 
-# we're going to change these
-  all_posters_individual_no_ah_tmp <- all_posters_individual_no_ah
-  container_for_this_day_part <- 0
-  poster_session_durations_no_th_ev$numbers_of_posters <- 0
-  poster_session_durations_no_th_ev$n_posters <- 0
-  poster_session_durations_no_th_ev$n_sessions <- 0
-  poster_session_durations_no_th_ev$c_sessions <- 0
-  poster_session_durations_no_th_ev$c_posters <- 0
-  all_of_the_day_parts <- 0
-  i <- 1
-
-# loop over each day-slot available for poster sessions
-# i is each day-part
-for(i in seq_len(nrow(poster_session_durations_no_th_ev))) {
-
-    # stop loop gracefully when we have assigned all people to day-parts
-  if (nrow(all_posters_individual_no_ah_tmp) == 0){ break }
-
-    # how many rows (posters) to we have remaning to allocate to a day-part?
-  print(glue('number of poster sessions remaining to classify is {nrow(all_posters_individual_no_ah_tmp)}'))
-
-  # need a buffer to avoid overfilling day-parts
-  target_capacity <- poster_session_durations_no_th_ev$n_max[i] - buffer
-
-  # prepare variables for the next round, just to be sure
-  cumulative_number_of_posters <- 0
-  container_for_this_day_part <- 0
-  this_session_posters <- 0
-  # j is each poster session
-  j <- 1
-
-  while(cumulative_number_of_posters <= target_capacity & 
-        j <= length(poster_session_ids)) {
-
-  # get a session 
-  this_session <- poster_session_ids[j]
-
-  # get names for all the people in this poster session
-  this_session_posters <- 
-    all_posters_individual_no_ah_tmp %>% 
-      filter(`Session Id` == this_session)
-
-#-------------------------------------------------------------------  
-  # check to see if these names are already in this day-part
-  # or in the paper sessions
-  if (j == 1){
-
-    # do nothing is this is the first incoming session 
-    names_already_in <-  0
-    names_already_in2 <- 0
-
-  } else {
-      # get distinct names for each session previously added
-      # to this day-part
-      # (don't care about same name multiple times in one session)
-      distinct_names_for_each_session_so_far <- 
-      container_for_this_day_part %>% 
-        group_by(uid) %>% 
-        distinct(uid, .keep_all = TRUE) %>% 
-        pull(uid)
-
-      # check to see if incoming session has 
-      # names that already exist in this day_part
-      names_already_in <- sum(this_session_posters$uid %in% distinct_names_for_each_session_so_far)
-
-      # check to see if papers for this day-part already have the name in it
-      names_papers_in_this_day_part <- 
-        sessions_in_day_parts_data_with_all_names_df %>% 
-        filter(day_part_from_while.y == poster_session_durations_no_th_ev[i,]$day_part) %>% 
-        pull(uid.y)
-
-      names_already_in2 <- sum(this_session_posters$uid %in% names_papers_in_this_day_part)
-
-  }
-
-#-------------------------------------------------------------------
-    # if there are no name conflicts, then
-    # accumulate the incoming session onto the sessions
-    # we've already added to this day-part
-
-    if (j == 1) {
-
-       container_for_this_day_part <- 
-        this_session_posters
-
-      } else {
-      if (names_already_in == 0 & names_already_in2 == 0) {
-
-      # if there are no name clashes, 
-      # add these names to previous names in this day-part
-      # bind rows of session that has no clashing  names with 
-      # sessions we already added to this day-part
-      container_for_this_day_part <- 
-        bind_rows(this_session_posters,
-                  container_for_this_day_part)
-
-    }  else {
-
-      # do not add this session to the day-part 
-      # because at least one of the names is already in there
-      }
-      }
-
-#-------------------------------------------------------------------  
-  # compute the number of posters we've added to this day-part
-  # by counting the abstract Id, because this excludes the session chairs
-    cumulative_number_of_posters <- 
-    nrow(container_for_this_day_part)
-
-   # store this result
-    poster_session_durations_no_th_ev$n_posters[i] <- cumulative_number_of_posters
-
-    print(glue('cumulative number of posters: {cumulative_number_of_posters}, j = {j}, i = {i}'))
-
-  # go to next session
-  j <- j + 1
-    }
-  # end of while loop
-
-   # add an identifier so we can see where we have assigned each session so far
- container_for_this_day_part_with_session_from_while <- 
-   container_for_this_day_part %>% 
-   mutate(day_part_from_while  = as.character(poster_session_durations_no_th_ev$day_part[i]))
-
- # bind rows to previous day-parts we might have done
-  if (i == 1) {
-
-  all_of_the_day_parts <- container_for_this_day_part_with_session_from_while
-
-  } else {
-
-    # check to see
-
-  all_of_the_day_parts <- bind_rows(container_for_this_day_part_with_session_from_while, 
-                                      all_of_the_day_parts)
-  }
-
- poster_session_durations_no_th_ev$n_sessions[i] <- length(unique(container_for_this_day_part_with_session_from_while$`Session Id`))
-
- poster_session_durations_no_th_ev$c_sessions[i] <-  sum(poster_session_durations_no_th_ev$n_sessions)
- poster_session_durations_no_th_ev$c_posters[i] <-  sum(poster_session_durations_no_th_ev$n_posters)
-
- print(poster_session_durations_no_th_ev %>% select(n_max, 
-                                                   day_part,
-                                                   n_posters,
-                                                   n_sessions, 
-                                                   c_sessions,
-                                                   c_posters))
-
- # for the next day-part, we need to make sure
- # we don't add the same session twice to different day-parts
- # so let's update our big table of posters, and drop the session that 
- # we've already got in the day-part
-
-all_posters_individual_no_ah_tmp <-  
-   all_posters_individual_no_ah_tmp %>% 
-   filter(! `Session Id` %in% unique(container_for_this_day_part$`Session Id`))
-
-}
-# end of for loop
+all_posters_individual_no_ah_ksack %>%
 
@@
 
-```{r}
-# we have 532 posters not in ah
-x1 <- 
-all_posters_individual_no_ah
+We may need to split up some of the general poster sessions to make it easier to pack them in.
+What is the distribution of session sizes?
+What if we chopped in half all the sessions bigger than 12 posters?
 
-x2 <- 
-all_of_the_day_parts
-
-# what sessions are not in there?
-session_output <- 
-x2 %>% 
-  group_by(`Session Id`) %>% 
-  pull(`Session Id`) %>% 
-  unique
-
-session_input <- 
-  x1 %>% 
-  group_by(`Session Id`) %>% 
-  pull(`Session Id`) %>% 
-  unique
-
-setdiff(session_input, session_output)
-
-x1 %>% 
-  filter(is.na(Title))
-
-# what are in the output of this loop that is not in the input?
-# duplicates?
-all_of_the_day_parts %>% 
-  distinct()
-
-# why is each poster in each day-part?
-all_of_the_day_parts %>% 
- arrange(uid) %>% View
-
-# how many abstracts
-posters_with_abstract_ids <- 
-all_of_the_day_parts %>% 
-  filter(!is.na(`Abstract Id`)) %>% 
-  group_by(day_part_from_while) %>% 
-  tally() %>% 
-  summarise(sum(n)) %>% 
-  pull()
-```
-
-We have `r posters_with_abstract_ids` posters with Abstract IDs (and so not chair/co-chairs).
-Compare to `r nrow(gen_and_org_poster_session_posters)` posters that we started with, and space available for `r sum(poster_session_durations$'N max')` posters.
+We have no clash with paper presenters.
 
 What do we stil need to do?
-
-- Posters
 

### 2017-11-06 analysis/paper/saa2018_posters_while_loop.Rmd
divide up poster sessions with >8 papers and schedule

--- 2017-11-04 analysis/paper/saa2018_posters.Rmd
@@
 gen_poster_session_posters_individual <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx")
-gen_poster_session_posters_individual$`Session Id` <- gen_poster_session_posters_individual$cuts
+gen_poster_session_posters_individual$`Session Id` <- as.character(gen_poster_session_posters_individual$cuts)
 # 422 posters in the general session
+
+# In an attempt to solve the scheduling problem...
+# Let's make each gen session poster a unique session to see if we can allocat them ok into the slots
+# gen_poster_session_posters_individual$`Session Id` <-
+#   as.character(glue('9999_{1:nrow(gen_poster_session_posters_individual)}_{gen_poster_session_posters_individual$`Session Id`}'))
 
@@
 ah_poster_session_posters_individual <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", sheet = "AFTER_HOURS", skip = 11)
-ah_poster_session_posters_individual$`Session Id` <- ah_poster_session_posters_individual$cuts
+ah_poster_session_posters_individual$`Session Id` <- as.character(ah_poster_session_posters_individual$cuts)
 
@@
   filter(`Session Type` == "Poster Symposium") %>% 
-  filter(Role == "Presenter")
+  filter(Role == "Presenter") %>% 
+  mutate(`Session Id`  = as.character(`Session Id`))
 # 196 posters in org + ah sessions
@@
   filter(!is.na(Title))
+```
+
+There are a small number of people with posters who are also discussants, etc. So we need to be sure we don't schedule posters at the same time as those activities.
+
+```{r}
+# Do people giving posters have other roles?
+# check 
+people_with_2_non_poster_roles <- 
+all_abstracts %>% 
+  group_by(`First Name`, `Last Name`) %>% 
+  add_tally() %>% 
+  filter(n >= 2) %>% 
+  filter(!grepl("Poster", `Session Type`)) %>% 
+  mutate(uid = glue('{`First Name`} {`Last Name`}'))
+
+people_in_posters_with_2_roles <- 
+  all_abstracts %>% 
+  group_by(`First Name`, `Last Name`) %>% 
+  add_tally() %>% 
+  filter(n >= 2) %>% 
+  filter(grepl("Poster", `Session Type`)) %>% 
+  mutate(uid = glue('{`First Name`} {`Last Name`}'))
+
+# any overlap?
+people_with_posters_and_another_role <- 
+people_in_posters_with_2_roles$uid[people_in_posters_with_2_roles$uid %in% people_with_2_non_poster_roles$uid]
+```
+
+Here's an experiment to break up the general sessions to see if we can improve the fit into the SAA session limits. We split the general sessions with more than so many papers.
+
+```{r}
+gen_poster_session_posters_individual %>%
+  group_by(`Session Id`) %>%
+  tally() %>%
+  ggplot(aes(n)) +
+  geom_histogram()
+
+n_allowed <- 8 # almost each poster is it's own session
+gen_poster_sessions_n_more_than_n <-
+gen_poster_session_posters_individual %>%
+  group_by(`Session Id`) %>%
+  tally()  %>%
+  filter(n >= n_allowed) %>%
+  pull(`Session Id`)
+
+gen_posters_in_sessions_n_more_than_n <-
+gen_poster_session_posters_individual %>%
+  filter(`Session Id` %in% gen_poster_sessions_n_more_than_n) %>%
+  group_by(`Session Id`) %>%
+  mutate(label = 1:n()) %>%
+  mutate(new_session_id1 =  label %% n_allowed) %>%
+  group_by(new_session_id1, idx = cumsum(new_session_id1 == 1L)) %>% 
+  ungroup() %>%
+  mutate(`Session Id` = glue('{`Session Id`}_{idx}'))  %>%
+  select(-label, -new_session_id1, -idx )
+
+gen_posters_in_sessions_n_more_than_n %>%
+  group_by(`Session Id`) %>%
+  tally() %>%
+  ggplot(aes(n)) +
+  geom_histogram()
+
+# now update the all_posters_individual data frame
+all_posters_individual <-
+all_posters_individual %>%
+  mutate(`Session Id` = as.character(`Session Id`)) %>%
+  # first remove the old gen session with more than 12 papers each
+  filter(!`Session Id` %in% gen_poster_sessions_n_more_than_n) %>%
+  # then add in those same posters, but with their new session ids
+  bind_rows(gen_posters_in_sessions_n_more_than_n)
 ```
@@
 
-poster_session_durations_no_th_ev
+# see if we can do the poster day-parts individually
+
+poster_session_durations_no_th_ev <- 
+  poster_session_durations %>% 
+  filter(day_part != "Th-evening") %>% 
+  mutate(poster_day_part = glue('poster-{day_part}')) %>% 
+  group_by(poster_day_part) %>% 
+  mutate(label = 1:n()) %>% 
+  ungroup() %>% 
+  mutate(poster_day_part = glue('{poster_day_part}_{label}')) %>% 
+  select(-label) %>% 
+  mutate(n_max = `N max`)
 ```
@@
 # to avoid overfilling day-parts
-buffer <-7 # is the largest to avoid overfill
+buffer <- 5 # is the largest to avoid overfill
 
@@
   container_for_this_day_part <- 0
-  poster_session_durations_no_th_ev$numbers_of_posters <- 0
   poster_session_durations_no_th_ev$n_posters <- 0
@@
     # how many rows (posters) to we have remaning to allocate to a day-part?
-  print(glue('number of poster sessions remaining to classify is {nrow(all_posters_individual_no_ah_tmp)}'))
+    # print(glue('number of poster sessions remaining to classify is {nrow(all_posters_individual_no_ah_tmp)}'))
 
@@
 
-    print(glue('cumulative number of posters: {cumulative_number_of_posters}, j = {j}, i = {i}'))
+    # print(glue('cumulative number of posters: {cumulative_number_of_posters}, j = {j}, i = {i}'))
 
@@
    container_for_this_day_part %>% 
-   mutate(day_part_from_while  = as.character(poster_session_durations_no_th_ev$day_part[i]))
+   mutate(day_part_from_while  = as.character(poster_session_durations_no_th_ev$poster_day_part[i]))
 
@@
  print(poster_session_durations_no_th_ev %>% select(n_max, 
-                                                   day_part,
+                                                   poster_day_part,
                                                    n_posters,
@@
 ```{r}
-# we have 532 posters not in ah
-x1 <- 
-all_posters_individual_no_ah
 
-x2 <- 
-all_of_the_day_parts
-
-# what sessions are not in there?
-session_output <- 
-x2 %>% 
-  group_by(`Session Id`) %>% 
-  pull(`Session Id`) %>% 
-  unique
-
-session_input <- 
-  x1 %>% 
-  group_by(`Session Id`) %>% 
-  pull(`Session Id`) %>% 
-  unique
-
-setdiff(session_input, session_output)
-
-x1 %>% 
-  filter(is.na(Title))
-
-# what are in the output of this loop that is not in the input?
-# duplicates?
-all_of_the_day_parts %>% 
-  distinct()
-
-# why is each poster in each day-part?
-all_of_the_day_parts %>% 
- arrange(uid) %>% View
-
-# how many abstracts
-posters_with_abstract_ids <- 
-all_of_the_day_parts %>% 
-  filter(!is.na(`Abstract Id`)) %>% 
-  group_by(day_part_from_while) %>% 
-  tally() %>% 
-  summarise(sum(n)) %>% 
-  pull()
 ```
 
-We have `r posters_with_abstract_ids` posters with Abstract IDs (and so not chair/co-chairs).
-Compare to `r nrow(gen_and_org_poster_session_posters)` posters that we started with, and space available for `r sum(poster_session_durations$'N max')` posters.
+We have no clash with paper presenters.
 
-What do we stil need to do?
-
-- Posters
+What do we still need to do?
 

### 2017-11-08 README.Rmd
AB says DONE

@@
 
-The goal of saaabstracts is to ...
+The goal of saaabstracts is to help organise the program for the 2018 Society of American Archaeology meeting.
+
+The specific tasks were to:
+
+- Classify the general paper and poster submissions into sessions of common topics    
+- Schedule the organised and general sessions into the slots made available by the SAA
+
+The task of classification was done by using hierarchical cluster analysis on the keywords submitted by authors with their abstracts.
+We tried topic modelling the abstracts, but this did not give much discriminating power.
+A little manual handling was needed after the `hclust` results because we don't have an easy way to set the upper limit of a cluster size.
+Once the general submissions were grouped into sessions, then we could send them out to reviwers
+
+The task of scheduling was done in several steps:
+
+1.
+Get the details of the available slots in which to put the sessions.
+We have 3.5 days, and each day has a number of 4 hour slots.
+Sessions vary in duration from 2 to 4 hours.
+Some sessions have durations assigned by the SAA, others we have to compute a sensible duration.
+We round to the nearest 15 min, and add 15 min of buffer time to each session to make the transitions between sessions is smooth.
+
+2.
+We use `for` and `while` loops to allocate sessions into day-parts (e.g.
+Thursday morning) to prevent any first author from being scheduled in two simultaneous day-parts.
+This looping process in necessary because we have to check the names of the people alrady allocated to a session to see if there will be a clash as we add each new session to the schedule.
+This process takes a minute or two.
+
+Currently, we take the papers and sessions in sequence from the spreadsheet we recieved them in, and serially assign them to day parts.
+This has the problem of putting many sessions on the same topic in the sam day-part (i.e.
+at the same time).
+In future, we should change the allocation so that session 1 goes into Thu-morning, session 2 goes into Thu-afternoon, session 3 into Thu-evening.
+This means that if sessions 1-3 are on similar topics, they wont be scheduled at the same time.
+This should not be a difficult modification for next time.
+We may also want to check for topic clashes using keywords.
+The posters are particularly challenging, and require some of the general session posters to be divided into very small sessions so we can fill the 60-poster slots most efficiently.
+
+3.
+Once papers and sessions are assigned to day-parts, we then use the multiple knapsack algorithm to sort sessions into 4-hour slots.
+This is an automatic method to organise sessions so that we can fill the slots most efficiently, for example, two 2-hour sessions can be found and placed together in one 4-hour slot.
+This process took up to 30 minutes.
+We did not do this to the poster sessions because the poster day-parts are not divided into slots based on duration, only the total number of posters.
+
+4.
+After the slots are filled, we do some checking to make sure we didn't lose any sessions or papers.
+This is important because the multiple knapsack algorithm may fail to sort all sessions if it cannot find an optimum solution.
+
+5. Then we check for conflicts with special requests from the SAA, and relocate sessions if necessary. And then check again to ensure we haven't introduced any clashes.
+
+6. Then do some more manual checkig, assign 4-hour slots to rooms, and upload to SAA website.
 

### 2018-01-10 analysis/paper/saa2018_nonposter_sessions_while_loop.Rmd
move to SAA folder

@@
   all_nonposter_sessions_with_session_durations$TIME
-z7 <- all_nonposter_sessions_with_session_durations
+z7 <- # just the sessions IDs
+  all_nonposter_sessions_with_session_durations %>% 
+  group_by(`Session Id`) %>% 
+  distinct(`Session Id`)
+
 all_of_the_day_parts <- NULL
@@
 
-       # print(progress_check)
+       print(progress_check)
 
@@
  # 
- # print(
- # all_of_the_day_parts %>% 
- #   group_by(day_part_from_while) %>% 
- #   summarise(papers = n(),
- #             sessions = length(unique(`Session Id`))))
+ print(
+ all_of_the_day_parts %>%
+   group_by(day_part_from_while) %>%
+   summarise(papers = n(),
+             sessions = length(unique(`Session Id`))))
 

### 2018-01-10 R/functions.R

@@ -1,0 +1,328 @@
+
+
+#' fill_day_parts_with_sessions
+#'
+#' @description The purpose of this function is to assign papers into time blocks of a conference schedule in a way that avoids scheduling the same person in more than one time block at the same time.
+#'
+#' @param total_capacity Data frame with one columns of day-parts (day_part, e.g. Thursday morning) and another column of total amount of hours to fill in each day-part (total_hours)
+#' @param all_papers Data frame of submitted papers, one row per paper. Each paper must already be allocated to a session, and have a value in a session_id column
+#' @param all_sessions Data frame of sessions, one row per session. Each session must have a value in a session_id column, and each session must have a value for the session_duraction' column
+#' @param buffer integer A number of papers to avoid overfilling each day-part. Vary this number to experiment with optimising how efficiently each block of time is filled.
+#' @param verbose_output logical, Show detailed output while the function is running? May be useful for trouble-shooting.
+#' @import dplyr
+#'
+#' @return A list with two data frames
+#' @export
+#'
+#' @examples
+#'
+#' output <-
+#' fill_day_parts_with_sessions(total_capacity,
+#'                              all_papers,
+#'                              all_sessions)
+#'
+#'
+fill_day_parts_with_sessions <- function(total_capacity,
+                                         all_papers,
+                                         all_sessions,
+                                         buffer = 30,
+                                         verbose_output = TRUE){
+
+  # check if we have a session_id column
+  if (length(all_papers$session_id) == 0) stop("There is no column 'session_id' in the data frame of conference papers. Please add this column and try again.")
+
+  # check to see if every paper has been assigned to a session
+  number_of_papers_with_session_id <- sum(!is.na(all_papers$session_id))
+  total_number_of_papers <-  nrow(all_papers)
+  number_of_papers_without_session_id <-
+    total_number_of_papers - number_of_papers_with_session_id
+  if (number_of_papers_with_session_id == 0) stop(glue('{number_of_papers_without_session_id} papers do not have a session ID. Please ensure that all papers are assigned to a session and try again.'))
+
+  # check to see if we have session durations
+  if (length(all_sessions$session_duration) == 0) stop("There is no column 'session_duration' in the data frame of conference sessions. Please add this column and try again.")
+
[…]

### 2018-01-10 analysis/paper/scratch.R

@@ -1,0 +1,217 @@
+create <- function(pkg = ".", auth_token, private = FALSE, copyright_holder){
+  devtools::use_github(pkg, auth_token = auth_token, protocol = "ssh", private = private)
+  devtools::use_mit_license(copyright_holder)
+  rrtools::use_analysis()
+  rrtools::use_readme_rmd()
+  rrtools::use_dockerfile()
+  rrtools::use_travis()
+  devtools::use_testthat()
+}
+
+create(auth_token = "12faaec9d29ab5cce52a27ca51d1112fc83a2e93")
+
+
+mtcars %>%
+  filter(mpg > 10) %>%
+  select(cyl)
+
+
+x <- 1:18
+u <- c(3,6,12)
+v <- c(5,10,15)
+data1 <- data.frame(x, loc_lower = findInterval(x, u),
+                    loc_upper =    findInterval(x, v + 1))
+data1
+
+write.csv(iris,
+          "../data/derived_data/iris.csv")
+
+all_nonposter_sessions %>%
+  filter(`Session Id` %in% c(3115, 3147, 3175, 3116, 3152, 3162)) %>%
+  distinct(`Session Id`)
+  View()
+
+  all_abstracts %>%
+    group_by(`Session Type`, `First Name`, `Last Name`) %>%
+    tally(sort = TRUE) %>%
+    filter(grepl("Poster", `Session Type`))
+
+# want no more than n in each group
+  n <-  3
+  x <- 1:13
+
+data.frame(x = x,
+           y = x %% 2)
+
+x2 <- (split(x, ceiling(seq_along(x)/n)))
+y <- unlist(map(names(split(x, ceiling(seq_along(x)/n))), ~rep(.x, n)))
+
+data.frame(x = x,
+           y = y)
+
+
+y <- as.character(cut_number(x, 3))
+data.frame(x = x,
+           y = y)
+
+chunk <- function(x,n)
+{
+  f <- sort(rep(1:(trunc(length(x)/n)+1),n))[1:length(x)]
+  return(split(x,f))
+}
+
+y <- chunk(x, 3)
+names(y)
+length(y[[1]])
+map(y, ~rep(x, length(.x)))
+
+sapply(y, )
+
+
+
+z7  <-
+  all_nonposter_sessions_with_session_durations  %>%
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162"))
+
+xx <- c(1:5, 1:5, 1:5)
+rle(xx)$lengths
+
+
[…]

### 2018-05-21 README.Rmd
more detailed readme

@@
 
-# saaabstracts
+# confschedlr
 
-The goal of saaabstracts is to help organise the program for the 2018 Society of American Archaeology meeting.
+The goal of confschedlr is to help organise the program for the 2018 Society of American Archaeology meeting.
 
@@
 A little manual handling was needed after the `hclust` results because we don't have an easy way to set the upper limit of a cluster size.
-Once the general submissions were grouped into sessions, then we could send them out to reviwers
+Once the general submissions were grouped into sessions, then we could send them out to reviwers.
+In the future this could be imporved by using a controlled vocabulary for keywords so we don't have so many synonyms to manually handle.
 
@@
 This is important because the multiple knapsack algorithm may fail to sort all sessions if it cannot find an optimum solution.
-
 5. Then we check for conflicts with special requests from the SAA, and relocate sessions if necessary. And then check again to ensure we haven't introduced any clashes.
@@
 
-> Authors, (`r format(Sys.Date(), "%Y")`). _Title of compendium_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
+> Authors, (`r format(Sys.Date(), "%Y")`). _confschedlr: An R package for scheduling conference sessions_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
 
@@
 
-You can install saaabstracts from github with:
+You can install confschedlr from github with:
 
@@
 # install.packages("devtools")
-devtools::install_github("benmarwick/saaabstracts")
+devtools::install_github("benmarwick/confschedlr")
 ```

### 2018-05-21 analysis/paper/paper.Rmd
more detailed readme

--- 2017-09-19 analysis/paper/paper.Rmd
@@
 
-library(saaabstracts)
-library(tm)
-library(topicmodels)
+library(confschedlr)
 library(readxl)
-library(Rmpfr)
-library(ggplot2)
+library(tidyverse)
+library(knitr)
+library(glue)
+
+```
+## Introduction
+
+## How many sessions do we have?
+
+```{r}
+library(readxl)
+org_sessions <- read_excel("../data/raw_data/Organized Session Abstracts_2018_AB.xlsx")
+
+gen_sessions <- read_excel("../data/raw_data/01 - gen_paper_sessions_FINAL(oct5)_DG_including_moves v2.xlsx")
+
+gen_sessions <- 
+gen_sessions %>% 
+  mutate(`Session Id` = SESSION) %>% 
+  mutate(`Session Id` =  ifelse(`Session Id` == "35B",
+          max(as.numeric(`Session Id`), na.rm = TRUE) + 1, 
+          as.numeric(`Session Id`) ))
+
+poster_sessions <- 
+  org_sessions %>% 
+  filter(`Session Type` == "Poster Symposium")
+
+# organised sessions without posters
+org_sessions <- 
+   org_sessions %>% 
+  filter(`Session Type` != "Poster Symposium")
+
+# how many unique organised sessions? 
+n_org_sessions <- length(unique(org_sessions$`Session Id`))
+
+# how many unique general sessions? 
+n_gen_sessions <- length(unique(gen_sessions$`Session Id`))
+
+# how many unique poster sessions? 
+n_pos_sessions <- length(unique(poster_sessions$`Session Id`))
+
+# total number of sessions
+total_number_of_sessions <-  n_org_sessions + n_gen_sessions + n_pos_sessions
+
+# combine organised and general sessions
+org_sessions$`Session Id` <- as.numeric(org_sessions$`Session Id`)
+gen_sessions$`Session Type` <- "Symposium"
+
+# get only columns that are also in the organised sessions data frame
+gen_sessions_to_join <- 
+gen_sessions %>% 
+  select(names(gen_sessions)[names(gen_sessions) %in% names(org_sessions)])
+
+all_nonposter_sessions <- 
+  bind_rows(gen_sessions_to_join, org_sessions)
 ```
 
-# Introduction
+We have this many sessions in each major category, according to my tally of the individual papers:
 
-```{r eval=FALSE}
-# read in the abstracts
-general <- read_excel("../data/raw_data/General Abstracts.xlsx")
+```{r}
+# table
+tribble(            ~'Session type',             ~number,
+                      'Organised sessions',        n_org_sessions,
+                      'Poster sessions',           n_pos_sessions,
+                      'General sessions',          n_gen_sessions)
 ```
 
-```{r eval=FALSE}
-# prepare a mapping structure, first item is column with text to model, others are metadata
-m <- list(content = "Abstract",  # content
-          id = "Abstract Id",      # metadata
-          title = "Title", 
-          author_first = "First Name",
-          author_last = "Last Name",
-          geo = "Geographic Focus",
-          keyword1 = "Keyword1", 
-          keyword2 = "Keyword2", 
-          keyword3 = "Keyword3", 
-          org = "Affiliation")
+With a total of `r total_number_of_sessions` sessions altogether, and a total number of non-poster sessions `r total_number_of_sessions - n_pos_sessions`.
+
+For scheduling into rooms and time slots, we wont include the posters, because they go in their own schedule in the ballroom.
+
+To help with preventing schedule clashes, we make a unique ID out of the first name and last name to track specific individuals.
+
+```{r}
+all_nonposter_sessions$uid <- 
+  with(all_nonposter_sessions, paste0(`First Name`, 
+                                                 "_", 
+                                                 `Last Name`))
 ```
 
-```{r eval=FALSE}
-# Create a corpus object from the abstracts
-general_corpus <- tm::Corpus(tm::DataframeSource(general), 
-                             readerControl = list(reader = tm::readTabular(mapping = m)))
+## How many 4 hour slots do we have available?
 
-# create a document term matrix
-general_tm <- tm::DocumentTermMatrix(general_corpus, 
-                                     control = list(stemming = TRUE, 
-                                                    stopwords = TRUE,
-                                                    minWordLength = 2, 
-                      s                              removeNumbers = TRUE, 
-                                                    removePunctuation = TRUE))
-```
+To find out how many 4 h slots we have available, I looked on the grid on 'Rooom&Time_DG.xlsx' and counted up the 4  slots.
+I did not count poster sessions here, because they are less than 4 h and they are outside of the room schedule.
+I rounded down a little:
 
-```{r eval=FALSE}
-# Use TF-IDF to wieght terms and remove rare ones
-term_tfidf <- 
-  tapply(general_tm$v/slam::row_sums(general_tm)[general_tm$i], 
-         general_tm$j, mean) * log2(tm::nDocs(general_tm)/slam::col_sums(general_tm > 0))
+- Thur afternoon has the ethics bowl from 1-3 pm, so I removed one whole 4 h slot from that part   
+- Fri morning has some time reserved from 9am-12 pm with the note "Reserved for Business Meeting after 12:00pm", so removed one whole 4 h slot from that part also
 
-summary(term_tfidf)
-# Median =  0.08989
-```
+Here's what I got for session slots in each day and part of day:
 
-```{r eval=FALSE }
-## Keeping the rows with tfidf >= median
-general_reduced_dtm <- general_tm[,term_tfidf >= summary(term_tfidf)[3]]
-summary(slam::col_sums(general_reduced_dtm))
-# save
-saveRDS(general_reduced_dtm, "../data/derived_data/general_reduced_dtm.rds")
+```{r}
+library(tidyverse)
+four_h_slots <- 
+  tribble(~part,       ~Th, ~Fr, ~Sa, ~Su,
+          "morning",    36,  31,  30,  28,
+          "afternoon",  35,  32,  30,  0,
+          "evening",    36,  0,    0,  0)
+
+(four_h_slots)
 ```
@@
 ```{r}
-general_reduced_dtm <- readRDS("../data/derived_data/general_reduced_dtm.rds")
+four_h_slots_long <- 
+  four_h_slots %>% 
+  gather(times, slot, -part)
+
+# how many 4 h slots
+n_four_h_slots <-  sum(four_h_slots_long$slot)
+```
+
+So that's a total of `r n_four_h_slots` 4 hour slots, in which we have to fit all the organised and general sessions. This does not include the posters.
+
+## What are the durations of the sessions?
+
+We've received information from the SAA about the length of specific sessions, so let's use that:
+
+```{r}
+saa_info_on_organised_session_duration <- 
+  read_excel("../data/raw_data/saa all organizedsession_timeslotABv3.xls", sheet = 2) %>% 
+  mutate(`Session Id` = `SESSION ID`)
+
+computed_general_session_duration <- 
+  gen_sessions %>% 
+  group_by(`Session Id`) %>% 
+  summarise(n_papers = n()) %>% 
+  mutate(TIME = n_papers * 0.25) %>%  # 15 mins per paper
+  mutate(`Session Id` =  ifelse(`Session Id` == "35B",
+          max(as.numeric(`Session Id`), na.rm = TRUE) + 1, 
+          as.numeric(`Session Id`) ))
+
+# combine general and organised session durations
+org_and_gen_session_durations <- 
+  saa_info_on_organised_session_duration %>% 
+  filter(!is.na(`Session Id`)) %>% 
+  mutate(n_papers = `N papers`,
+         TIME = as.numeric(TIME)) %>% 
+  select(`Session Id`, TIME, n_papers) %>% 
+  rbind(computed_general_session_duration) %>% 
+  filter(!is.na(TIME)) %>% 
+  # add 15 m min  to each session for a changeover time
+  mutate(TIME = if_else(TIME == 4.0, 
+                        TIME,
+                        TIME + 0.25))
+
+head(org_and_gen_session_durations)
+
+# join the session duration onto each paper
+
+all_nonposter_sessions_with_session_durations <- 
+  all_nonposter_sessions %>% 
+  mutate(`Session Id` = as.numeric(`Session Id`)) %>% 
+  left_join(org_and_gen_session_durations)
+
+head(all_nonposter_sessions_with_session_durations)
+```
+
+## How can we prevent one person scheduled in two simultaneous sessions?
+
+Because we will have many parallel sessions, we need to take care that we don't put one person in two simultaneous sessions.
+
+We start by taking a day-part, assign a session to it, and then looping over that day-part to add another session, and another, and so on.
+Each time we add a session to that day-part, we check to see if the names in the incoming session already exist in the sessions we previously added to this day-part.
+If a name has already been added to that day-part, then we skip adding the session, and move to the next session.
+We keep doing this until the tally of hours of all the sessions in the day-part matches the total number of hours we have for that day-part.
+
+We can compute the total capacity in hours of each day-part like this:
+
+```{r}
+four_h_slots_long$cumsum <- cumsum(four_h_slots_long$slot)
+
+xx <- 
+four_h_slots_long %>% 
+  mutate(day_part = paste0(times, "-", part)) %>% 
+  select(-part, -times) %>% 
+  rename(end = cumsum) %>% 
+  filter(slot != 0)
+
+# get the number of the start slot to go with the number of the end slot
+ xx$start <- 0
+for(i in 1:(nrow(xx))) {
+  xx$start[i] <- 
+    if_else(i == 1, 
+           1,
+            as.numeric(xx[ i-1, 2]) + 1)
+}
+# each slot is 4 hours long
+total_cap <- 
+  xx %>% 
+  mutate(total_capacity = slot * 4) %>% 
+  select(day_part, slot, total_capacity)
+
+# View
+total_cap
+```
+
+Now we will take session from the big table of sessions and add it into, for example, Th-morning. We'll keep adding more sessions until we reach the capacity of that day-part.
+
+But, we also need to loop over the sessions as we add more to the day-part to check to see if a person's name is already there or not.
+If the name is already there, we'll skip adding that session, and try the next session.
+
+```{r while-loop, echo = TRUE, eval = TRUE cache = TRUE}
+
+# prepare some variables... get all the sessions, 
+# but we're going to modify this, so make a copy
+org_and_gen_session_durations$session_time_allowed <- 
+  as.numeric(org_and_gen_session_durations$TIME)
+all_nonposter_sessions_with_session_durations$session_time_allowed <- 
+  all_nonposter_sessions_with_session_durations$TIME
+z7 <- # just the sessions IDs
+  all_nonposter_sessions_with_session_durations %>% 
+  group_by(`Session Id`) %>% 
+  distinct(`Session Id`)
+
+all_of_the_day_parts <- NULL
+total_cap$hours_of_sessions <- 0
+total_cap$n_sessions <- 0
+total_cap$c_sessions <- 0
+container_for_this_day_part_with_day_from_while <- 0
+i <- 1
+
+# start with each day-part...
+for(i in seq_len(nrow(total_cap))) {
+
+  # stop loop gracefully when we have assigned all people to day-parts
+  if (nrow(z7) == 0){ break }
+
+  # how many rows (people) to we have remaning to allocate to a day-part?
+  print(glue('number of people remaining to classify is {nrow(z7)}'))
+
+  # add some buffer time to nudge it back so we don't overfill a day-part
+  # maybe if this buffer is bigger, the knapsack will be faster?
+  # A big enough buffer should avoid the 0 group in the knapsack results
+  # which indicates that the knapsack fn can't fit all the sessions in the
+  # day-part
+  buffer <- 30
+  target_capacity <- total_cap$total_capacity[i] - buffer
+
+  # prepare a while statement that will take a session, 
+  # check to see if there is a name clash with sessions,
+  # skip adding it if there is clash, then check to see
+  # the cumulative duration of the sessions in the day-part 
+  # has exceeded the total duration of the day-part,
+  # and stop adding sessions if it has.
+
+# make sure some variables are empty
+cumulative_duration_of_sessions   <-  0
+container_for_this_day_part <- 0
+all_the_people_in_the_session <- 0
+j <- 1
+
+while(cumulative_duration_of_sessions <= target_capacity & 
+      j <= nrow(z7)) {
+
+    # randomly reorder the rows to see if we can solve a problem
+    # z7 <- z7[sample(nrow(z7)),]
+    get_a_session <- z7[j, ]
+
+    # get the names of all the people in this session...
+    all_the_people_in_this_session <- 
+      all_nonposter_sessions_with_session_durations %>% 
+      filter(`Session Id` == get_a_session$`Session Id`)
+
+    # get names from what we already have added to this day-part
+     if (j == 1) {
+      # do nothing is this is the first incoming session 
+       distinct_names_for_each_session_so_far <- 0
+    } else {
+      # get distinct names for each session previously added
+      # to this day-part
+      # (don't care about same name multiple times in one session)
+      distinct_names_for_each_session_so_far <- 
+      container_for_this_day_part %>% 
+        group_by(uid) %>% 
+        distinct(uid, .keep_all = TRUE) %>% 
+        pull(uid)
+    }
+
+    # compare names already in here to names we are about to add
+    if (j == 1) {
+      # do nothing 
+      names_already_in <- 0
+      } else {
+      # check to see if incoming session has names that already exist in this day_part
+      names_already_in <- sum(all_the_people_in_this_session$uid %in% distinct_names_for_each_session_so_far)
+      }
+
+    # accumulate the incoming session onto the sessions
+    # we've already added to this day-part, if there are no
+    # name conflicts
+    if (j == 1) {
+       container_for_this_day_part <- 
+        all_the_people_in_this_session
+      } else {
+      if (names_already_in == 0) {
+      # if there are no name clashes, 
+      # add these names to previous names in this day-part
+
+      # bind rows of session that has no clashing  names with 
+      # sessions we already added to this day-part
+      container_for_this_day_part <- 
+        bind_rows(all_the_people_in_this_session,
+                  container_for_this_day_part)
+    }  else {
+
+        # do not add this session to the day-part
+      }
+    }
+
+    # compute the duration of session that we've added so far to 
+    # this day-part
+    cumulative_duration_of_sessions <- 
+      container_for_this_day_part %>% 
+      distinct(`Session Id`, .keep_all = TRUE) %>% 
+      pull(session_time_allowed) %>% 
+      sum
+
+    # store this result
+    total_cap$hours_of_sessions[i] <- cumulative_duration_of_sessions
+
+    # show a check of what's happening
+    if (j == 1){
+    progress_check <-    
+      data_frame(i = i, 
+                 j = j,
+               'cumulative duration' = cumulative_duration_of_sessions,
+               'sessions added' =  length(unique(container_for_this_day_part$`Session Id`)),
+               people =  nrow(container_for_this_day_part)
+                )
+    } else {
+      progress_check <- bind_rows(progress_check, 
+                                  data_frame(i = i,
+                                             j = j,
+               'cumulative duration' = cumulative_duration_of_sessions,
+               'sessions added' =  length(unique(container_for_this_day_part$`Session Id`)),
+               people =  nrow(container_for_this_day_part)
+                ))
+    }
+
+       print(progress_check)
+
+       # move to next session
+       j <- j + 1
+    } 
+# end of the while loop
+
+ # add an identifier so we can see where we have assigned each session so far
+ container_for_this_day_part_with_day_from_while <- 
+   container_for_this_day_part %>% 
+   mutate(day_part_from_while  = as.character(total_cap$day_part[i]))
+
+ # bind rows to previous day-parts we might have done
+  if (i == 1) {
+
+  all_of_the_day_parts <- container_for_this_day_part_with_day_from_while
+
+  } else {
+
+    # check to see
+
+  all_of_the_day_parts <- bind_rows(container_for_this_day_part_with_day_from_while, 
+                                      all_of_the_day_parts)
+  }
+
+ # # progress check
+ print(glue('i = {i}'))
+ # 
+ print(
+ all_of_the_day_parts %>%
+   group_by(day_part_from_while) %>%
+   summarise(papers = n(),
+             sessions = length(unique(`Session Id`))))
+
+ total_cap$n_sessions[i] <- length(unique(container_for_this_day_part_with_day_from_while$`Session Id`))
+
+ total_cap$c_sessions[i] <-  sum(total_cap$n_sessions)
+
+ print(total_cap)
+
+ # for the next day-part, we need to make sure
+ # we don't add the same session twice to different day-parts
+ # so let's update our big table of abstracts, and drop the session that 
+ # we've already got in the day-part
+
+ z7 <-  
+   z7 %>% 
+   filter(! `Session Id` %in% unique(container_for_this_day_part$`Session Id`))
+
+}
+# end of for loop
+
+# out put the results
+write.csv(all_of_the_day_parts, 
+          "../data/derived_data/output_of_day_part_classification.csv")
+
+write.csv(total_cap, "../data/derived_data/total_cap.csv")
 ```
@@
 ```{r}
-library(ldatuning)
-result <- FindTopicsNumber(
-  general_reduced_dtm,
-  topics = seq(from = 2, to = 100, by = 2),
-  metrics = c("Griffiths2004", "CaoJuan2009", "Arun2010", "Deveaud2014"),
-  method = "Gibbs",
-  control = list(seed = 77),
-  mc.cores = 2L,
-  verbose = TRUE
-)
+# read them in, if we need to
+all_of_the_day_parts <- readr::read_csv("../data/derived_data/output_of_day_part_classification.csv")
 
-knitr::kable(result)
-FindTopicsNumber_plot(result)
+total_cap <- readr::read_csv("../data/derived_data/total_cap.csv")
+```
+
+Did we get all sessions allocated to a day-part?
+
+```{r}
+sessions_we_missed <- 
+  setdiff(unique(all_nonposter_sessions$`Session Id`),
+          unique(all_of_the_day_parts$`Session Id`))
+```
+
+We missed `r length(sessions_we_missed)` sessions: `r sessions_we_missed`.
+
+What did we get from this process?
+The table below shows each day-part, the number of hours that we have available to fill (`total_capacity`) and the number of hours that we have filled so far (`hours_of_sessions`):
+
+```{r}
+total_cap %>% 
+  rename(`Number of sessions` = n_sessions,
+         `Cumulative number of sessions` = c_sessions)
+```
+
+Have we succeeded in avoiding placing one person in two sessions scheduled in the same day-part?
+
+Let's see if there are any names that occur in two or more distinct sessions in the same day-part.
+
+```{r while-loop-testing, results = 'hide', echo = TRUE}
+# check to see what we got from the loop...
+sessions_per_day <- 
+all_of_the_day_parts %>% 
+  group_by(day_part_from_while) %>% 
+  summarise(n_sessions = length(unique(`Session Id`)))
+
+hours_per_day <- 
+  all_of_the_day_parts %>% 
+  group_by(day_part_from_while, `Session Id`) %>% 
+  slice(1) %>% 
+  summarise(n_hours = sum(session_time_allowed)) %>% 
+  ungroup() %>% 
+  group_by(day_part_from_while) %>% 
+  summarise(n_hours = sum(n_hours))
+
+n_missing_papers <- 
+  nrow(all_nonposter_sessions_with_session_durations) - nrow(all_of_the_day_parts)
+
+n_missing_sessions <- 
+  length(unique(all_nonposter_sessions_with_session_durations$`Session Id`)) - length(unique(all_of_the_day_parts$`Session Id`))
+
+missing_sessions <- 
+  unique(all_nonposter_sessions_with_session_durations$`Session Id`)[!(unique(all_nonposter_sessions_with_session_durations$`Session Id`) %in% unique(all_of_the_day_parts$`Session Id`)) ]
+
+# have we succeeded with no name clashes?
+# in each day-part, do any names occur more than once?
+
+# how to get it? testing...
+ex <- 
+tribble(~'day_part', ~'sess', ~'auth',
+        "th-m",       1,        'al', 
+        "th-a",       2,        'be',
+        "th-a",       3,        'be',
+        "th-e",       4,        'ce',
+        "fr-a",       5,        'de',
+        "fr-a",       5,        'de')
+
+out <- 
+ex %>% 
+  group_by(day_part, auth) %>% 
+  filter(n() > 1) %>% 
+  group_by(sess) %>% 
+  filter(n() == 1)
+
+# deploy...
+clash <- 
+all_of_the_day_parts %>% 
+  # get authors that have multiple rows (roles or sessions) day-part
+  group_by(day_part_from_while, uid) %>% 
+  filter(n() > 1  ) %>% 
+  # get rows where 
+  group_by(`Session Id`) %>% 
+  filter(n() == 1) # get rows where there are names that occur in more than one session in each day-part
+
+```
+
+Currently we have `r nrow(clash)` schedule clashes, that is, people who are scheduled to be in two or more sessions that are scheduled in the same day-part.
+That's `r ifelse(nrow(clash) == 0, "great", "not ideal")`.
+
+## Organise the sessions in each day-part into 4 hour slots
+
+Now that we have allocated all the organised sessions into day-parts, we need to group them into 4 hour slots so we can assign them to specific rooms.
+
+We can use the mulitple knapsack algorithm for this. For example, Thurs morning has 36 slots of 4 hours each.
+
+We have a series of functions, one for each day-part, then we run them, and write the output to CSV.
+
+```{r}
+# thu-morning
+
+thu_mo <- function(){
+
+# get the slots available for this day-part
+thu_morning_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Th-morning') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+thu_morning_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Th-morning') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, thu_morning_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- thu_morning_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+thu_morning_sessions$computed_slot <- solution$ksack
+
+write.csv(thu_morning_sessions,
+          "../data/derived_data/thu_morning_sessions.csv")
+}
+
+# Here's thursday afternoon:
+
+# thu-afternoon
+thu_af <- function(){
+
+# get the slots available for this day-part
+thu_afternoon_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Th-afternoon') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+thu_afternoon_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Th-afternoon') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, thu_afternoon_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- thu_afternoon_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+thu_afternoon_sessions$computed_slot <- solution$ksack
+
+write.csv(thu_afternoon_sessions, 
+          "../data/derived_data/thu_afternoon_sessions.csv")
+}
+
+# Here's thursday evening:
+
+# thu-evening
+
+thu_ev <- function(){
+
+# get the slots available for this day-part
+thu_evening_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Th-evening') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+thu_evening_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Th-evening') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, thu_evening_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- thu_evening_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+thu_evening_sessions$computed_slot <- solution$ksack
+
+write.csv(thu_evening_sessions, 
+          "../data/derived_data/thu_evening_sessions.csv")
+}
+
+# Here's friday morning:
+
+# fri-morning
+
+fri_mo <- function(){
+
+# get the slots available for this day-part
+fri_morning_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Fr-morning') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+fri_morning_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Fr-morning') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, fri_morning_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- fri_morning_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+fri_morning_sessions$computed_slot <- solution$ksack
+
+write.csv(fri_morning_sessions, 
+          "../data/derived_data/fri_morning_sessions.csv")
+}
+
+# Here's friday afternoon:
+
+# fri-afternoon
+
+fri_af <- function(){
+
+# get the slots available for this day-part
+fri_afternoon_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Fr-afternoon') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+fri_afternoon_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Fr-afternoon') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, fri_afternoon_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- fri_afternoon_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+fri_afternoon_sessions$computed_slot <- solution$ksack
+
+write.csv(fri_afternoon_sessions, 
+          "../data/derived_data/fri_afternoon_sessions.csv")
+}
+
+# Here's Saturday morning:
+
+# sat-morning
+sat_mo <- function(){
+
+# get the slots available for this day-part
+sat_morning_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Sa-morning') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+sat_morning_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Sa-morning') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, sat_morning_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- sat_morning_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+sat_morning_sessions$computed_slot <- solution$ksack
+
+write.csv(sat_morning_sessions, 
+          "../data/derived_data/sat_morning_sessions.csv")
+}
+
+# Here's saturday afternoon:
+
+# sat-afternoon
+
+sat_af <- function(){
+
+# get the slots available for this day-part
+sat_afternoon_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Sa-afternoon') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+sat_afternoon_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Sa-afternoon') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, sat_afternoon_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- sat_afternoon_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+sat_afternoon_sessions$computed_slot <- solution$ksack
+
+write.csv(sat_afternoon_sessions, 
+          "../data/derived_data/sat_afternoon_sessions.csv")
+}
+
+# Here's sunday morning:
+
+# sun-morning
+sun_mo <- function(){
+
+# get the slots available for this day-part
+sun_morning_slots <- 
+  total_cap %>% 
+  filter(day_part == 'Su-morning') %>% 
+  pull(slot)
+
+# get the session id's assigned to this day-part
+sun_morning_sessions <- 
+  all_of_the_day_parts %>% 
+  group_by(`Session Id`) %>% 
+  filter(day_part_from_while == 'Su-morning') %>% 
+  slice(1)
+
+slots_to_fill <- rep(4, sun_morning_slots) * 60 # so many 4 hour slots that we need to fill
+
+sessions_we_have <- sun_morning_sessions$session_time_allowed * 60
+
+library(adagio)
+# mknapsack calling signature is: mknapsack(values, weights, capacities)
+solution <- mknapsack(sessions_we_have, 
+                      sessions_we_have,
+					  slots_to_fill)
+
+sun_morning_sessions$computed_slot <- solution$ksack
+
+write.csv(sun_morning_sessions, 
+          "../data/derived_data/sun_morning_sessions.csv")
+}
+
 ```
@@
 ```{r}
-# Determine k number of topics
-harmonicMean <- function(logLikelihoods, precision = 2000L) {
-  llMed <- median(logLikelihoods)
-  as.double(llMed - log(Rmpfr::mean(exp(-Rmpfr::mpfr(logLikelihoods,
-                                       prec = precision) + llMed))))
-}
+# run all of these
+thu_mo()
+thu_af()
+thu_ev()
+fri_mo()
+fri_af()
+sat_mo()
+sat_af()
+sun_mo()
+
+# why do these sometime have a 0 group? Because the fn can't fit it all in! So we need to add more buffer to each day-part in the while loop
+```
+
+## Resolving conflicts with requested times
+
+Now let's get all the names of the people in each day-part to check for the special requests that a few people have made:
+
+```{r}
+
+# read in the day-part data from the while-loop
+sessions_in_day_parts <- list.files("../data/derived_data", pattern = "_sessions.csv", full.names = TRUE)
+sessions_in_day_parts_data <- 
+  map(sessions_in_day_parts, 
+       ~read_csv(.x))
+
+# for each day-part, join with the all names of the people in each session
+sessions_in_day_parts_data_with_all_names <- 
+  map(sessions_in_day_parts_data,
+      ~.x %>% 
+        left_join(all_of_the_day_parts, by = 'Session Id')
+      )
+
+# get day-parts to use in file names
+day_parts_of_list <- 
+  map_chr(sessions_in_day_parts_data_with_all_names, 
+    ~unique(.x$day_part_from_while.y))
+
+# write these out to CSV
+walk2(sessions_in_day_parts_data_with_all_names, 
+     day_parts_of_list,
+     ~write_csv(.x,  path = glue("../data/derived_data/{.y}-all-names.csv")))
+
+# any computer slots with 0?
+sessions_in_day_parts_data_with_all_names_df <- 
+bind_rows(sessions_in_day_parts_data_with_all_names)
+
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(computed_slot == 0)
+```
+
+Sat morning is a focus of many of the special requests, let's have a look at that
+
+```{r}
+Sa_morning_names <- sessions_in_day_parts_data_with_all_names[[which(day_parts_of_list == "Sa-morning")]]
+
+# read in the sheet on conflicts
+conflicts <- read_excel("../data/raw_data/Conflicts.xlsx")
+# make a unique name variable
+conflicts$uid <- glue('{conflicts$PRENOM}_{conflicts$NOM}')
+# get the ones for Sa-morning
+Sa_morning_conflicts <- 
+  conflicts %>% 
+  filter(grepl("Sa", CONFLICTS))
+
+# check to see if we have a problem
+Sa_morning_conflict_result <- Sa_morning_names$uid.y[Sa_morning_names$uid.y %in% Sa_morning_conflicts$uid]
+
+# double-check
+double_check <- 
+Sa_morning_names[Sa_morning_names$`Last Name.y` %in% Sa_morning_conflicts$NOM, ]
+# it's ok
+```
+
+We have `r length(Sa_morning_conflict_result)` conflicts on Saturday morning, that's `r ifelse(length(Sa_morning_conflict_result) == 0, "great", "not ideal")`
+
+What about conflicts in other day-parts?
+
+```{r}
+other_conflicts <- 
+  conflicts %>% 
+  filter(!grepl("Sa", CONFLICTS))
+```
+
+There are just `r nrow(other_conflicts)`, so let's do it one by one.
+
+1. Deborah Nichols doesn't want to be in Thurs AM, Fri AM. When do we have her?
+
+```{r}
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(uid.y == "Deborah_Nichols") %>% 
+  select(`Session Id`, uid.y, day_part_from_while.y )
+```
+
+Sa-afternoon and Sa-morning, she's ok, great!
+
+2. Leah McCurdy and session 3109, but what is the conflict exactly? Not on Sat morning from 9-11 am.
+
+```{r}
+Sa_morning_names %>% 
+  filter(uid.y == "Leah_McCurdy")
+
+# when is she?
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(uid.y == "Leah_McCurdy") %>% 
+  select(day_part_from_while.y)
+```
+
+She's ok, on Fr-morning
+
+3. Session 2929 should not be on "Thurs PM, Thurs EVE, Fri PM, SAT AM, SAT PM". When do we have it?
+
+```{r}
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(`Session Id` == 2929) %>% 
+  select(`Session Id`, uid.y, day_part_from_while.y )
+```
+
+This is a problem, we put this session in Th-afternoon.
+
+Can we tack it onto Su-morning, right at the end? According to the results of our while-loop, we have many hours spare on Su-morning:
+
+```{r}
+total_cap
+```
+
+Let's double-check what we've got after the multiple knapsack work:
+
+```{r}
+day_parts_filled_hours <- 
+sessions_in_day_parts_data_with_all_names_df %>% 
+  group_by(`Session Id`) %>% 
+  slice(1) %>% 
+  group_by(day_part_from_while.y) %>% 
+  summarise(hours_in_day_part = sum(session_time_allowed.y ))
+```
+
+It's the same. So how many spare hours do we have on Su-morning?
+
+```{r}
+Su_morning_hours_available <- 
+total_cap %>% 
+  filter(day_part == "Su-morning") %>% 
+  select(total_capacity) %>% 
+  pull()
+
+Su_morning_hours_filled <- 
+  day_parts_filled_hours %>% 
+  filter(day_part_from_while.y == "Su-morning") %>% 
+  select(hours_in_day_part) %>% 
+  pull()
+
+Su_morning_spare_hours <- 
+  Su_morning_hours_available - Su_morning_hours_filled
+
+```
+
+We have `r Su_morning_spare_hours` hours unused on Sunday morning.
+Heaps of space.
+So let's move Session 2929 to Sunday morning.
+Don't forget to update the computed slot value so it gets it's own slot.
+
+```{r}
+
+# what is the max computed_slot for Su-morning?
+max_computed_slot_Su_morning <- 
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(day_part_from_while.y == "Su-morning") %>% 
+  summarise(max_computed_slot_Su_morning = max(computed_slot)) %>% 
+  pull(max_computed_slot_Su_morning)
+
+sessions_in_day_parts_data_with_all_names_df <- 
+sessions_in_day_parts_data_with_all_names_df %>% 
+  mutate(day_part_from_while.y = if_else(`Session Id` == 2929, 
+                                         "Su-morning", 
+                                         day_part_from_while.y)) %>% 
+  mutate(computed_slot = ifelse(`Session Id` == 2929, 
+                                         max_computed_slot_Su_morning + 1, 
+                                         computed_slot))
+
+# check it
+check_session_2929_to_sunday_morning <- 
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(`Session Id` == 2929) %>% 
+  select(day_part_from_while.y) %>% 
+  unique()
+
+# check to see if there are conflicts with names, should be 0
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(day_part_from_while.y == "Su-morning") %>%
+  group_by(`Session Id`, uid.y)  %>% 
+  add_tally() %>% 
+  distinct(`Session Id`) %>% 
+  group_by(uid.y) %>% 
+  filter(n() > 1 )
+
+# how many slots in Sun?
+
+# available?
+total_cap #28
+
+# used?
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(day_part_from_while.y == "Su-morning") %>% 
+  distinct(computed_slot) %>% 
+  max()
+# 21
+
+```
+
+What do we stil need to do?
+
+- Many sessions of the same topic are put in the same day-part, which is undesirable.
+We may want to consider changing the while-loop to fill each day-part with one session per day-part, rather than filling a day-part, then moving to the next day-part, etc.
+
+- AB says
+
+>-  row 10 has session 2929 sharing with session 3165. But session 2929 needs 16 slots, which is 4 hours.
+
+Problem was that I forgot to update the computed slot value. Fixed now.
+
+>- Missing 3115, 3147, 3175, 3116, 3152, 3162
+
+```{r}
+
+all_nonposter_sessions %>% 
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162")) 
+# yes, there
+
+# how many sessions""
+length(unique(all_nonposter_sessions$`Session Id`)) #  271
+
+all_nonposter_sessions_with_session_durations  %>% 
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162")) 
+# yes
+
+org_and_gen_session_durations  %>% 
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162")) 
+# yes
+
+all_of_the_day_parts  %>% 
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162")) 
+# not there
+
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162"))
+# not there
+
+# how many sessions?
+length(unique(sessions_in_day_parts_data_with_all_names_df$`Session Id`)) # 265
+
+# are these sessions very long?
+hours_of_missing_sessions <- 
+all_nonposter_sessions %>% 
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162"))  %>% 
+  group_by(`Session Id`) %>% 
+  tally() %>% 
+  mutate(minutes = n * 15) %>% 
+  mutate(hours = minutes / 60)
+# nope
+total_hours_missing <-  sum(hours_of_missing_sessions$hours)
+
+# what are they about?
+all_nonposter_sessions %>% 
+  filter(`Session Id` %in% missing_sessions) %>% 
+  select(Abstract, Title)
+# no title or abstracts!
+
 ```
@@
 ```{r}
-# run a loop over the abstract with different numbers of topics to find which number of topcis is the best
-seqk <- seq(2, 100, 1)
-burnin <- 1000
-iter <- 1000
-keep <- 50
+
+# So, no idea what's going on with these. Let's see if there are any problems with putting them on Sunday
+missing_sessions_details <- 
+  all_nonposter_sessions %>% 
+  filter(`Session Id` %in% c("3115", "3147", "3175", "3116", "3152", "3162")) %>% 
+  mutate(uid.y = uid)
+
+# are any people in more than one session Sunday? yes
+# what about all the day-parts with spare hours
+clashes_on_day_x <- function(day) {
+  sessions_in_day_parts_data_with_all_names_df %>% 
+ filter(day_part_from_while.y == day) %>% 
+  bind_rows(missing_sessions_details) %>% 
+  group_by(`Session Id`, uid.y)  %>% 
+  add_tally() %>% 
+  distinct(`Session Id`) %>% 
+  group_by(uid.y) %>% 
+  filter(n() > 1 )
+}
+
+total_cap_remaining <- 
+  total_cap %>% 
+  mutate(hours_unused =  total_capacity - hours_of_sessions) %>% 
+  select(day_part, hours_unused)
+
+day_with_spare_hours <- c("Fr-morning",
+                          "Fr-afternoon",
+                          "Sa-morning",
+                          "Sa-afternoon",
+                          "Su-morning")
+
+total_cap_remaining %>% 
+  mutate(clashes = map(day_part, clashes_on_day_x)) %>% 
+  mutate(n_clashes = map_int(.$clashes, ~nrow(.x)))
+
+# Looks like Th-morning has no clashes, double-check, should be 0
+
+ sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(day_part_from_while.y ==  "Th-morning") %>% 
+  bind_rows(missing_sessions_details) %>% 
+  group_by(`Session Id`, uid.y)  %>% 
+  add_tally() %>% 
+  distinct(`Session Id`) %>% 
+  group_by(uid.y) %>% 
+  filter(n() > 1 )
 ```
 
-```{r eval=FALSE }
-system.time(fitted_many <- 
-              lapply(seqk, function(k) topicmodels::LDA(general_reduced_dtm, 
-                                                        k = k,
-                                                        method = "Gibbs",
-                                                        control = list(burnin = burnin,
-                                                                       iter = iter, 
-                                                                       keep = keep) )))
+Seems that Th-morning is a good place to put these missing sessions. So let's do that. We'll have to double-up in some slots
 
-saveRDS(fitted_many, "../data/derived_data/fitted_many.rds")
+```{r}
+# find out how many hours we have for each slot on th-morn
+slot_on_thu_morn_with_less_than_4_h <- 
+sessions_in_day_parts_data_with_all_names_df %>% 
+  filter(day_part_from_while.y == "Th-morning") %>% 
+  distinct(`Session Id`, .keep_all = TRUE) %>% 
+  group_by(computed_slot) %>% 
+  summarise(total_duration = sum(session_time_allowed.y)) %>% 
+  filter(total_duration < 4) %>% 
+  mutate(minutes = total_duration * 60) %>% 
+  mutate(minutes_free = (4*60) - minutes)
+
+# how many hours do we need to put in there?
+minutes_of_missing_sessions <- 
+  hours_of_missing_sessions$minutes
+
+# # what is the max computed_slot for Su-morning?
+# max_computed_slot_Th_morning <- 
+# sessions_in_day_parts_data_with_all_names_df %>% 
+#   filter(day_part_from_while.y == "Th-morning") %>% 
+#   summarise(max_computed_slot_Th_morning = max(computed_slot)) %>% 
+#   pull(max_computed_slot_Th_morning)
+# 
+# # assign missing sessions to Th morn, update slots
+# missing_sessions_details <- 
+# missing_sessions_details %>% 
+#   mutate(day_part_from_while.y = "Th-morning")
+# 
+# missing_sessions_details_indices  <- 
+# missing_sessions_details %>% 
+#   group_by(`Session Id`) %>% 
+#   group_indices()
+# 
+# # update computed slot
+# missing_sessions_details <- 
+#   missing_sessions_details %>% 
+#     mutate(computed_slot = missing_sessions_details_indices + max_computed_slot_Th_morning) 
 ```
 
+We can't put them all on Th morn, there's not slot with enough free space to fit more than one of the missing sessions in
+
 ```{r}
-fitted_many <- readRDS("../data/derived_data/fitted_many.rds")
+
+# # bind to rest of the papers
+# sessions_in_day_parts_data_with_all_names_df <- 
+#   bind_rows(sessions_in_day_parts_data_with_all_names_df ,
+#             missing_sessions_details)
+#             
+# 
+# # check it, should be Th-morning
+# check_missing_sessions_to_thu_morning <- 
+# sessions_in_day_parts_data_with_all_names_df %>% 
+#   filter(`Session Id` %in% missing_sessions) %>% 
+#   select(day_part_from_while.y) %>% 
+#   unique()
+# 
+# # check for conflicts with names
+#  sessions_in_day_parts_data_with_all_names_df %>% 
+#   filter(day_part_from_while.y ==  "Th-morning") %>% 
+#   group_by(`Session Id`, uid.y)  %>% 
+#   add_tally() %>% 
+#   distinct(`Session Id`) %>% 
+#   group_by(uid.y) %>% 
+#   filter(n() > 1 )
+
 ```
 
-```{r}
-# extract logliks from each topic
-logLiks_many <- lapply(fitted_many, function(L)  L@logLiks[-c(1:(burnin/keep))])
+## Summary
 
-# compute harmonic means
-hm_many <- sapply(logLiks_many, function(h) harmonicMean(h))
-
-# to know the optimal number of topics:
-optimal_number_of_topics <- seqk[which.max(hm_many)]
-
-# plot it
-library(ggplot2)
-lda_plot <- ggplot(data.frame(seqk, 
-                             hm_many), 
-                  aes(x=seqk, 
-                      y=hm_many)) + 
-  geom_path(lwd=1.5)  +
-  xlab('Number of Topics') +
-  ylab('Harmonic Mean') +
-  geom_vline(xintercept = seqk[which.max(hm_many)],
-             colour = "red") +
-  annotate("text",
-           x = 55, 
-           y = -150000, 
-           label = paste("The optimal number of topics is", 
-                         seqk[which.max(hm_many)])) +
-  theme_bw() +
-  ggtitle(expression(atop("Latent Dirichlet Allocation Analysis of SAA General Abstracts", atop(italic("How many distinct topics in the abstracts?"), ""))))
-```
+Now we've got all the organised sessions scheduled into slots and day-parts with no clashes, and no conflicts with requested times.
 
 ```{r}
-# now run the model with the optimum number of topics
-system.time(general_model <- topicmodels::LDA(general_reduced_dtm, 
-                                              optimal_number_of_topics,
-                                              method = "Gibbs", 
-                                              control = list(iter=2000, 
-                                                             seed = 0622)))
+# this is our final result
+write_csv(sessions_in_day_parts_data_with_all_names_df,
+          "../data/derived_data/sessions_in_day_parts_data_with_all_names_df.csv")
 
-# explore the model
-general_topics <- topicmodels::topics(general_model, 1)
+# can we make a visual schedule?
+x <- 
+sessions_in_day_parts_data_with_all_names_df %>% 
+  group_by(day_part_from_while.y, computed_slot) %>% 
+  summarise(sessions_for_plot = toString(unique(`Session Id`)))
 
-## In this case I am returning the top 30 terms.
-general_terms <- as.data.frame(topicmodels::terms(general_model, 30), 
-                               stringsAsFactors = FALSE)
-```
+# put day-slots in order
+x$day_part_from_while.y <- factor(x$day_part_from_while.y, 
+                                  levels = total_cap$day_part)
 
-```{r}
-library(tidytext)
-tidy_topics <- tidy(general_model, matrix = "beta")
+ggplot(x,
+       aes(day_part_from_while.y,
+           computed_slot)) +
+ geom_tile(colour = "black",
+           fill = "white") +
+  geom_text(aes(label = sessions_for_plot), 
+                  size = 3) +
+    scale_y_continuous(breaks = 0:max(x$computed_slot)) +
+  theme_bw() +
+  xlab("") +
+  ylab("Room-ish") +
+  ggtitle("Visual schedule for SAA2019 conference non-poster sessions",
+          subtitle = "The numbers on the plot are the Session Ids. \n We can think of  each row as equivalent to a room at the conference venue. ")
 
-library(ggplot2)
-library(dplyr)
+ggsave("../figures/visual_schedule_for_saa2019_conference_non_poster_sessions.png")
 
-tidy_topics_top_terms <- tidy_topics %>%
-  group_by(topic) %>%
-  top_n(10, beta) %>%
-  ungroup() %>%
-  arrange(topic, -beta)
-```
-
-```{r}
-tidy_topics_top_terms %>%
-  mutate(term = reorder(term, beta)) %>%
-  ggplot(aes(term, beta, fill = factor(topic))) +
-  geom_col(show.legend = FALSE) +
-  facet_wrap(~ topic, scales = "free") +
-  coord_flip()
-```
-
-```{r}
-tidy_documents <- tidy(general_model, matrix = "gamma")
+# make a spreadsheet that looks like this plot
 
 library(tidyr)
-tidy_documents_topics_gamma <-
-  tidy_documents %>% 
-    spread(topic, gamma)
+visual_schedule_plot_as_spreadsheet <- 
+  x %>% 
+  select(day_part_from_while.y, 
+         sessions_for_plot, 
+         computed_slot) %>% 
+  distinct(sessions_for_plot, 
+           .keep_all = TRUE) %>% 
+  spread(day_part_from_while.y, 
+         sessions_for_plot, 
+         fill = "")
 
-abstract_classifications <- 
-  tidy_documents %>%
-  group_by(document) %>%
-  top_n(1, gamma) %>%
-  ungroup() %>% 
-  arrange(desc(gamma))
-```
-
-```{r eval=FALSE}
-mydata <- tidy_documents_topics_gamma[,-1]
-
-# Determine number of clusters
-library(NbClust)
-nclust <- NbClust(mydata, method = "kmeans")
-
-library(fpc)
-pamk.best <- pamk(mydata)
-
-# Determine number of clusters
-wss <- (nrow(mydata)-1)*sum(apply(mydata,2,var))
-for (i in 2:15) wss[i] <- sum(kmeans(mydata, 
-  	centers=i)$withinss)
-plot(1:15, wss, type="b", xlab="Number of Clusters",
-  ylab="Within groups sum of squares")
-
-# K-Means Cluster Analysis
-fit <- kmeans(mydata, 5) # 5 cluster solution
-# get cluster means 
-aggregate(mydata,by=list(fit$cluster),FUN=mean)
-# append cluster assignment
-mydata <- data.frame(mydata, fit$cluster)
-```
-
-## Wokring in a local Docker container
-
-To run this project in a local Docker container, I start a bash shell in the project directory:
-
-Do this one time only, build my Docker container from Dockerfile at top-level in my project:
+write_csv(visual_schedule_plot_as_spreadsheet,
+          "../data/derived_data/visual_schedule_plot_as_spreadsheet.csv")
 
 ```
-docker build - < Dockerfile
-docker ps
-```
-
-Take note of the container ID, which you get after `docker ps`
-
-Then run my Docker container, access it via the web browser, link the project dir to the Docker drive. Make sure you're in a local directory that is sharable (C:\Users\yourname is usually good):
-
-```
-docker run -dp 8787:8787 -v ${pwd}:/home/rstudio/ -e ROOT=TRUE  <container ID>
-```
-
-Then go to `http://192.168.99.100:8787/` or `localhost:8787` in youor browser, log in (rstudio/rstudio), and start your RStudio project by double-clickin on the .Rproj file
-
-When you're done, in the shell, stop all docker containers with:
-
-```
-docker stop $(docker ps -a -q)
-```
-
-Or do `docker ps` and `docker stop <container ID>` just to stop one container.
 

### 2018-05-21 analysis/paper/scratch.R

@@ -8,7 +8,7 @@ create <- function(pkg = ".", auth_token, private = FALSE, copyright_holder){
   devtools::use_testthat()
 }
 
-create(auth_token = "12faaec9d29ab5cce52a27ca51d1112fc83a2e93")
+create(auth_token = "xx")
 
 
 mtcars %>%
